### 2022-05-17 GSE158433/GSE158422.Rmd
added GSE158422 to enable comparison of gene expression and methylation in cancer

---
title: "GSE158422: DNA methylation in Lung squamous cell carcinoma"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
  toc: true
theme: cosmo
---

## Introduction

The purpose of this analysis is to apply different enrichment analysis approaches,
and see which approach is most parsimonious with gene expressoin differences.

We will specifically be testing the GMEA approach versus existing approaches,
namely GOseq and mitch.

This study (GSE158433) is a super-series contains matched RNA-seq (GSE158420) and 
Infinium EPIC array (GSE158422) data for matched control and cancer tissues.

In this script, we will be processing the EPIC DNA methylation data and performing the
statistical analysis in limma and saving the results for downstream enrichment analysis.

```{r,packages}

suppressPackageStartupMessages({
  library("plyr")
  library("R.utils")
  library("missMethyl")
  library("limma")
  library("DMRcate")
  library("DMRcatedata")
  library("topconfects")
  library("minfi")
  library("IlluminaHumanMethylation450kmanifest")
  library("RColorBrewer")
  library("IlluminaHumanMethylation450kanno.ilmn12.hg19")
  library("GEOquery")
  library("eulerr")
  library("plyr")
  library("gplots")
  library("reshape2")
  library("forestplot")
  library("beeswarm")
  library("RCircos")
  library("qqman")
  library("ENmix")
})

```

## Obtaining array annotations

```{r,annotation}

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","Regulatory_Feature_Group","Islands_Name","Relation_to_Island")])
promoters <- grep("Prom",myann$Regulatory_Feature_Group)

```

## Obtain array data

GSE158422 is the accession number for the array data.

```{r,arraydata}

ARRAY_DATA="GSE158422_RAW.tar"
if(!file.exists(ARRAY_DATA)){
  download.file("ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE158nnn/GSE158422/suppl/GSE158422_RAW.tar",
    destfile=DEST) 
  untar(tarfile = ARRAY_DATA)
}
  SERIES_MATRIX="GSE158422_series_matrix.txt.gz"
download.file("ftp://ftp.ncbi.nlm.nih.gov/geo/series/GSE158nnn/GSE158422/matrix/GSE158422_series_matrix.txt.gz", 
  destfile=SERIES_MATRIX)
gse <- getGEO(filename=SERIES_MATRIX)
baseDir <- "."
sample_metadata <- pData(phenoData(gse))
targets <- sample_metadata
mybase <- unique(gsub("_Red.idat.gz" ,"",  gsub("_Grn.idat.gz", "" ,list.files(".",pattern = "GSM",recursive = TRUE))))
targets$Basename <- mybase
head(targets)
rgSet <- read.metharray.exp(targets = targets)
rgSet

```

## Filtering and normalisation

```{r,norm,fig.cap="Figure 1. Normalisation of bead-array data with SWAN."}

mSet <- preprocessRaw(rgSet)
mSetSw <- SWAN(mSet,verbose=FALSE)
par(mfrow=c(1,2), cex=0.8)
densityByProbeType(mSet[,1], main = "Raw")
densityByProbeType(mSetSw[,1], main = "SWAN")

```

```{r,filterprobes}

# include sex chromosomes
detP <- detectionP(rgSet)
keep <- rowSums(detP < 0.01) == ncol(rgSet)
mSetSw <- mSetSw[keep,]

# exclude SNP probes
mSetSw <- mapToGenome(mSetSw)
mSetSw_nosnp <- dropLociWithSnps(mSetSw)

# exclude sex chromosomes
xyprobes <- anno$Name[anno$chr %in% c("chrX","chrY")]
mSetFlt <- mSetSw[which(!rownames(mSetSw) %in% xyprobes),]

```

```{r,beta_m_vals}

# include sex chromosomes
meth <- getMeth(mSetSw)
unmeth <- getUnmeth(mSetSw)
Mval <- log2((meth + 100)/(unmeth + 100))
beta <- getBeta(mSetSw)
dim(Mval)

# exclude sex chromosomes
meth <- getMeth(mSetFlt)
unmeth <- getUnmeth(mSetFlt)
Mval_flt <- log2((meth + 100)/(unmeth + 100))
beta_flt <- getBeta(mSetFlt)
dim(Mval_flt)

```

# MDS analysis

```{r,scree,fig.cap="Figure 2. Screeplot shows contribution of top principal components to the overall variation in the experiment."}

# scree plot shows the amount of variation in a dataset that is accounted
# for by the first N principal components
myscree <- function(mx,n=10,main="") {
  pc <- princomp(mx)$sdev
  pcp <- pc/sum(pc)*100
  pcp <- pcp[1:10]
  barplot(pcp,cex.names = 1,las=2,ylim=c(0,60),
      ylab="percent (%) variance explained", main=main)
  text((0.5:length(pcp)*1.2),pcp,label=signif(pcp,3),pos=3,cex=0.8)
}

par(mfrow=c(2,1))  
myscree(Mval,main="incl sex chr")
myscree(Mval_flt,main="excl sex chr")

```

#MDS plot 1

```{r,mds1,fig.width = 8 ,fig.height = 8,fig.cap="Figure 3. MDS plot coloured by ART classification."}

colnames(targets) <- gsub(" ","_",colnames(targets)) 
colnames(targets) <- gsub(":","_",colnames(targets))
targets$sex<-factor(targets$`gender_ch1`)
targets$art<-factor(targets$`art_subtype`)
sample_groups <- factor(targets$art)
colour_palette=brewer.pal(n = length(levels(sample_groups)), name = "Paired")
colours <- colour_palette[as.integer(factor(targets$art))]
plot(1,axes = FALSE,xlab="",ylab="",main="MDS by ART type")
legend("center",legend=levels(sample_groups),pch=16,cex=1.2,col=colour_palette)
mydist <- plotMDS(Mval, labels=targets$Sample_Name,col=colours,main="sex chromosomes included")
mydist_flt <- plotMDS(Mval_flt, labels=targets$Sample_Name,col=colours,main="sex chromosomes excluded")

```

## save data object

```{r,save}

save.image("GSE158422.Rdata")

```