### 2022-05-24 GSE158433/GSE158420.Rmd
integrative analysis begun

---
title: "GSE158420: RNA expression in Lung squamous cell carcinoma"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
  toc: true
theme: cosmo
---

## Introduction

The purpose of this analysis is to apply different enrichment analysis approaches,
and see which approach is most parsimonious with gene expression differences.

This study (GSE158433) is a super-series contains matched RNA-seq (GSE158420) and 
Infinium EPIC array (GSE158422) data for matched control and cancer tissues.

In this script, we will be analysing the gene expression profiles, followed by mitch
enrichment analysis.

```{r,packages}

suppressPackageStartupMessages({
  library("DESeq2")
  library("mitch")
  library("gplots")
  library("kableExtra")
})

```

## Download expression data

The data is available from the following address.
There are matched cancer and normal tissues.

https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE158420

The count matrix contains excel errors which need to be fixed manually or with sed.

```{r,datadl1}

URL="https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE158420&format=file&file=GSE158420%5Fcounts%2Etxt%2Egz"
FILE="GSE158420_counts.txt.gz"
FILE2="GSE158420_counts.txt"
download.file(URL,destfile=FILE)

x <- read.table(FILE2)

head(x)

```

## Samplesheet

The sample names will be converted to samplesheet.

```{r,samplesheet1}

ss <- data.frame(colnames(x))
names(ss) <- "label"
ss$tumor <- factor(grepl("T",ss$label))
ss$patient <- factor(gsub("T","",gsub("N","",ss$label)))
rownames(ss) <- ss$label

```

## Filtering and normalisation

```{r,filt}

dim(x)
xx <- x[which(rowMeans(x)>=10),]
dim(xx)

rpm <- xx/colSums(xx) * 1000000

```

## MDS plot

```{r,mds}

colours <- gsub("FALSE","lightblue",gsub("TRUE","pink",as.character(ss$tumor)))

labels <- gsub("N","",gsub("T","",gsub("X","",colnames(xx))))

plot(cmdscale(dist(t(xx))), xlab="Coordinate 1", ylab="Coordinate 2", type = "p",pch=19,col=colours,cex=2)
text(cmdscale(dist(t(xx))), labels=labels )

```

## DESEQ2

```{r,deseq2}

dds <- DESeqDataSetFromMatrix(countData = xx , colData = ss, design = ~ patient + tumor )
res <- DESeq(dds)
z<- results(res)
vsd <- vst(dds, blind=FALSE)
zz<-cbind(as.data.frame(z),assay(vsd))
dge<-as.data.frame(zz[order(zz$pvalue),])
dge[1:10,1:6]

sig <- subset(dge,padj<0.05)
dge_up <- rownames(subset(sig,log2FoldChange>0))
dge_dn <- rownames(subset(sig,log2FoldChange<0))
length(dge_up)
length(dge_dn)

write.table(dge,file="GSE158420_dge.tsv",sep="\t",quote=FALSE)

```

## Mitch

```{r,mitch}

download.file("https://reactome.org/download/current/ReactomePathways.gmt.zip", destfile="ReactomePathways.gmt.zip")
unzip("ReactomePathways.gmt.zip")
genesets <- gmt_import("ReactomePathways.gmt")

m <- mitch_import(dge,DEtype="deseq2")

res <- mitch_calc(m, genesets, priority="effect")

head(res$enrichment_result,20) %>% 
  kbl(caption = "top differentially regulated pathways") %>%
  kable_paper("hover", full_width = F)

```

## Session information

```{r,sessioninfo}

sessionInfo()

```