### 2022-10-11 GSE158433/GSE158422_fry.Rmd

---
title: "GSE158422: DNA methylation in Lung squamous cell carcinoma - GMEA using fry()"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
  toc: true
theme: cosmo
---

## Introduction

Here I will use fry to perform GMEA analysis.

```{r,packages}

suppressPackageStartupMessages({
  library("plyr")
  library("R.utils")
  library("missMethyl")
  library("limma")
  library("DMRcate")
  library("DMRcatedata")
  library("topconfects")
  library("minfi")
  library("IlluminaHumanMethylation450kmanifest")
  library("RColorBrewer")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("GEOquery")
  library("eulerr")
  library("plyr")
  library("gplots")
  library("reshape2")
  library("forestplot")
  library("beeswarm")
  library("RCircos")
  library("qqman")
  library("ENmix")

  library("tictoc")
  library("kableExtra")

})

source("../meth_functions.R")

```

## Load data

```{r,annotation}

load("GSE158422.Rdata")

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","Regulatory_Feature_Group","Islands_Name","Relation_to_Island")])
promoters <- grep("Prom",myann$Regulatory_Feature_Group)
promoters <- myann[promoters,]

```

Curate probe-sets.

Firstly for whole gene.

```{r,probesets_wholegene}

gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
sets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))

summary(unlist(lapply(sets,length)))

```

Now for promoters only.

```{r,probesets_promoters}

gp <- promoters[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
psets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))

summary(unlist(lapply(psets,length)))

```

## Design matrix

```{r,design}

samplesheet<-targets
groups <- factor(samplesheet$case,levels=c("tumor: normal","tumor: tumor"))
sex <- factor(samplesheet$sex,levels=c("0","1"))
design <- model.matrix(~ sex + groups)
mxs <- Mval

```

## set up mroast

Roast is a self contained test involving sample permulations.

This is beneficial for two reasons.

1. Sample permutations will prevent against false positives caused by strong
probe-probe correlation structures.

2. Competitive tests are not suited to experiments where there are unequal up and down
regulaiton, which is common for epigenetics studies.

Parallel analysis of GMEA with whole gene and promoters only.

mroast with 1000 whole gene probe sets, 2000 rotations: 42 s

mroast with 1000 gene promoter probe sets, 2000 rotations: 41 s

```{r,mroast1}

tic()
mro_res1 <- mroast(y=mxs, index = head(sets,1000), design = design, contrast = ncol(design), nrot = 2000 )
toc()

head(mro_res1,20) %>%
  kbl(caption="mroast test with 1000 whole gene probe sets") %>%
  kable_styling(full_width=FALSE)

tic()
mro_res2 <- mroast(y=mxs, index = head(psets,1000), design = design, contrast = ncol(design), nrot = 2000 )
toc()

head(mro_res2,20) %>%
  kbl(caption="mroast test with 1000 promoter only probe sets") %>%
  kable_styling(full_width=FALSE)

```

## Set up fry test

Fry is good because the p-values are analytically determined, and will have high precision.

Small test whole gene: 28 secs.
Small test gene promoter: 27 secs.

```{r,smallfry1}

tic()
fry_res1 <- fry(y=mxs, index = head(sets,1000), design = design, contrast = ncol(design) )
toc()

head(fry_res1,20) %>%
  kbl(caption="fry test with 1000 whole gene probe sets") %>%
  kable_styling(full_width=FALSE)

tic()
fry_res2 <- fry(y=mxs, index = head(psets,1000), design = design, contrast = ncol(design) )
toc()

head(fry_res2,20) %>%
  kbl(caption="fry test with 1000 gene promoter probe sets") %>%
  kable_styling(full_width=FALSE)

```

Now we can run the test with all probe sets.

Whole gene: 639 sec (10.5 mins)

Promoter: 261 sec (4.35 mins)

```{r,bigfry1}

tic()
fry_res1 <- fry(y=mxs, index = sets, design = design, contrast = ncol(design) )
toc()

head(fry_res1,20) %>%
  kbl(caption="fry test with all whole gene probe sets") %>%
  kable_styling(full_width=FALSE)

tic()
fry_res2 <- fry(y=mxs, index = psets, design = design, contrast = ncol(design) )
toc()

head(fry_res2,20) %>%
  kbl(caption="fry test with all gene promoter probe sets") %>%
  kable_styling(full_width=FALSE)

```

## save data object

```{r,save}

save.image("GSE158422_fry.Rdata")

```