### 2022-11-14 GSE158433/GSE158422_genesetlevel.Rmd
reorganising analysis: gene and geneset Rmd reports

---
title: "GSE158422: DNA methylation in Lung squamous cell carcinoma - GMEA using fry()"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
  toc: true
theme: cosmo
---

## Introduction

There are many ways to perform enrichment analysis of Infinium methylation data.
The purpose of this work is to determine which one of these methods perform "best".
Best performance is a function of specificity, sensitivity and biological consistency.

* Specificity can be defined as the ability to control false discovery rate appropriately.
This will be evaluated by analyzing a randomised data set.

* Sensitivity is defined as the ability to identify genesets undergoing differentially methylation.
This can be evaluated if we begin with a large study with many samples, analyse it and then repeat
the process with fewer samples.
The more sensitive methods will be able to identify differentially methylated pathways with
a smaller sample size.

* As methylation is known to be a characteristic of silenced genes, we can use this feature to examine
whether the methylation pathway enrichment analysis is congruent with gene expression data, as this
dataset has matched methylation and gene expression profiles for all samples.

In addition I will be measuring at the execution time for these approaches.

The approaches that I will be examining:

* GOmeth: an approach to filter differential methylated probes followed by over-representation testing.

* ebGSEA: global test to rank genes followed by Wilcoxon Test or Known Population Median Test

* methylGSA: mRRA robust rank aggregation, mGLM

* Fry test on probes belonging to a gene, followed by Camera PR based on significance/effectsize?

* Median of probe intensity for each gene in each sample followed by Fry test on pathways.

```{r,packages}

suppressPackageStartupMessages({
  library("plyr")
  library("R.utils")
  library("missMethyl")
  library("limma")
  library("DMRcate")
  library("DMRcatedata")
  library("topconfects")
  library("minfi")
  library("IlluminaHumanMethylation450kmanifest")
  library("RColorBrewer")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("GEOquery")
  library("eulerr")
  library("plyr")
  library("gplots")
  library("reshape2")
  library("forestplot")
  library("beeswarm")
  library("RCircos")
  library("qqman")
  library("ENmix")

  library("tictoc")
  library("kableExtra")
  library("mitch")
  library("fgsea")

})

source("../meth_functions.R")

```

## Load data

```{r,annotation}

load("GSE158422.Rdata")

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","Regulatory_Feature_Group","Islands_Name","Relation_to_Island")])
promoters <- grep("Prom",myann$Regulatory_Feature_Group)
promoters <- myann[promoters,]

```

Curate probe-sets.

Firstly for whole gene.

```{r,probesets_wholegene}

gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
sets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))

summary(unlist(lapply(sets,length)))

```

Now for promoters only.

```{r,probesets_promoters}

gp <- promoters[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
psets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))

summary(unlist(lapply(psets,length)))

```

## Design matrix

```{r,design}

samplesheet<-targets
groups <- factor(samplesheet$case,levels=c("tumor: normal","tumor: tumor"))
sex <- factor(samplesheet$sex,levels=c("0","1"))
design <- model.matrix(~ sex + groups)
mxs <- Mval

```

## set up mroast

Roast is a self contained test involving sample permulations.

This is beneficial for two reasons.

1. Sample permutations will prevent against false positives caused by strong
probe-probe correlation structures.

2. Competitive tests are not suited to experiments where there are unequal up and down
regulaiton, which is common for epigenetics studies.

Parallel analysis of GMEA with whole gene and promoters only.

mroast with 1000 whole gene probe sets, 2000 rotations: 42 s

mroast with 1000 gene promoter probe sets, 2000 rotations: 41 s

```{r,mroast1}

tic()
mro_res1 <- mroast(y=mxs, index = head(sets,1000), design = design, contrast = ncol(design), nrot = 2000 )
toc()

head(mro_res1,20) %>%
  kbl(caption="mroast test with 1000 whole gene probe sets") %>%
  kable_styling(full_width=FALSE)

tic()
mro_res2 <- mroast(y=mxs, index = head(psets,1000), design = design, contrast = ncol(design), nrot = 2000 )
toc()

head(mro_res2,20) %>%
  kbl(caption="mroast test with 1000 promoter only probe sets") %>%
  kable_styling(full_width=FALSE)

```

## Set up fry test

Fry is good because the p-values are analytically determined, and will have high precision.

Small test whole gene: 28 secs.
Small test gene promoter: 27 secs.

```{r,smallfry1}

tic()
fry_res1 <- fry(y=mxs, index = head(sets,1000), design = design, contrast = ncol(design) )
toc()

head(fry_res1,20) %>%
  kbl(caption="fry test with 1000 whole gene probe sets") %>%
  kable_styling(full_width=FALSE)

tic()
fry_res2 <- fry(y=mxs, index = head(psets,1000), design = design, contrast = ncol(design) )
toc()

head(fry_res2,20) %>%
  kbl(caption="fry test with 1000 gene promoter probe sets") %>%
  kable_styling(full_width=FALSE)

```

Now we can run the test with all probe sets.

Whole gene: 639 sec (10.5 mins)

Promoter: 261 sec (4.35 mins)

```{r,bigfry1}

tic()
fry_res1 <- fry(y=mxs, index = sets, design = design, contrast = ncol(design) )
toc()

head(fry_res1,20) %>%
  kbl(caption="fry test with all whole gene probe sets") %>%
  kable_styling(full_width=FALSE)

tic()
fry_res2 <- fry(y=mxs, index = psets, design = design, contrast = ncol(design) )
toc()

head(fry_res2,20) %>%
  kbl(caption="fry test with all gene promoter probe sets") %>%
  kable_styling(full_width=FALSE)

```

## Pathway analysis

Simple GSEA using FGSEA (gene permutation) and mitch (rank-ANOVA).

```{r,fgsea}

#download.file("https://reactome.org/download/current/ReactomePathways.gmt.zip", destfile="ReactomePathways.gmt.zip")
#unzip("ReactomePathways.gmt.zip")
genesets <- gmt_import("ReactomePathways.gmt")

stat1 <- sign(as.numeric(factor(fry_res1$Direction))-1.5)  * -log10(fry_res1$PValue)
stat1[is.na(stat1)] <- 0
names(stat1) <- rownames(fry_res1)
fres1 <- fgseaMultilevel(pathways=genesets,stats=stat1)
fres1 <- fres1[order(fres1$pval),]

head(fres1,20) %>%
  kbl(caption="WG fry + fgsea Reactome results") %>%
  kable_styling(full_width=FALSE)

stat2 <- sign(as.numeric(factor(fry_res2$Direction))-1.5)  * -log10(fry_res2$PValue)
stat2[is.na(stat2)] <- 0
names(stat2) <- rownames(fry_res2)
fres2 <- fgseaMultilevel(pathways=genesets,stats=stat2)
fres2 <- fres2[order(fres2$pval),]
head(fres2,20)  %>%
  kbl(caption="Promoter fry + fgsea Reactome results") %>%
  kable_styling(full_width=FALSE)

```

```{r,mitch}

stat1 <- data.frame(stat1)
mres1 <- mitch_calc(stat1, genesets, priority="effect")
head(mres1$enrichment_result,20) %>%
  kbl(caption="WG fry + mitch Reactome results") %>%
  kable_styling(full_width=FALSE)

stat2 <- data.frame(stat2)
mres2 <- mitch_calc(stat2, genesets, priority="effect")
head(mres2$enrichment_result,20) %>%
  kbl(caption="Promoter fry + mitch Reactome results") %>%
  kable_styling(full_width=FALSE)

```

## save data object

```{r,save}

save.image("GSE158422_fry.Rdata")

```