### 2022-11-24 gmea2.Rmd
some work on the b-proof project

---
title: "GSE74548: Gene Methylation Enrichment Analysis"
author: "hcy_meth meta-analysis group"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

##  Introduction

Previously, Mandhri and I analysed the B-PROOF 450K data, trying to understand
whether vitamin supplementation caused changes in gene methylation.
We used Limma and some basic analyses, which showed no specific probes with FDR<0.05,
nor any DMRs.

Here we will attempt another approach.
First aggregate median signal for probes belonging to a gene, followed by limma and enrichment analysis.

To validate this approach it is possible to cut the dataset in half and see what % of results are consistent.

```{r,libs}

suppressPackageStartupMessages({
  library("limma")
  library("parallel")
  library("dplyr")
  library("kableExtra")
  library("eulerr")
  library("mitch")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylation450kanno.ilmn12.hg19")
  library("tictoc")
})

source("meth_functions.R")

```

## Get array annotation data

```{r,anno}

ann450k <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
myann <- data.frame(ann450k[,c("UCSC_RefGene_Name","Regulatory_Feature_Group")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
sets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))
promoters <- grep("Prom",myann$Regulatory_Feature_Group)

```

# DM3 - association of Hcy with gene methylation

It is thought that high plasma homocysteine can inhibit DNA methylation.
Let's see whether that is the case, and which genes are most affected.
This analysis is conducted at the whole gene level as well as on the level of promoters.

This approach involves performing limma, and then for each gene the median t-score is
calculated.
This value is then used by mitch.

```{r,dma3_mitch}

design <- readRDS(file="hcy_design.rds")
mval <- readRDS(file="hcy_mval.rds")

tic()
res <- main(mval=mval,design=design,sets=sets,cores=4)
res_df <- res[[1]] # enrichment results
res_top <- res[[2]] # limma results
toc() ## 502-522 sec elapsed on 8 cores (8.3-8.7 mins)

hist(res_top$t)

hist(res_df$median)

head(res_df,30) %>%
  kbl(caption="Top GMEA genes") %>%
  kable_styling("hover", full_width = F)

genesets <- gmt_import("ReactomePathways.gmt")

y <- res_df[,"median",drop=FALSE]

colnames(y) <- "meth"

res <- mitch_calc(y, genesets, priority="effect")

head(res$enrichment_result,30) %>%
  kbl(caption="Top GMEA pathways", row.names = F) %>%
  kable_styling("hover", full_width = F)

```

A slight variation is to use cameraPR instead of mitch to better control for correlation.

```{r,dma3_cameraPR}

stat <- res_df$median

names(stat) <- rownames(res_df)
stat[is.na(stat)] <- 0

tic()
cres <- cameraPR(statistic=stat, index=genesets, use.ranks = FALSE, inter.gene.cor=0.01, sort = TRUE)
cres <- subset(cres,NGenes>4)
cres$FDR <- p.adjust(cres$PValue,method="fdr")
mymedians <- sapply(1:nrow(cres),function(i) {
  myset <- genesets[[which(names(genesets) == rownames(cres)[i])]]
  mystats <- stat[names(stat) %in% myset]
  median(mystats)
})
cres$median <- mymedians
toc() # 1.0 sec

sig <- subset(cres,FDR<0.05)

head(sig[order(-abs(sig$median)),],20) %>%
  kbl(caption = "Top GMEA pathways found with CAMERA after 5% FDR filtering") %>%
  kable_styling("hover", full_width = F)

```

Now try another approach by first aggregating and then performing limma, then pathway enrichment.

Aggregate mval to genes and examine pathway enrichment in blood.

```{r,gsmea_hcy}

gsmea <- function(mval,design,probesets,genesets,cores=2) {
  ag <- mclapply(probesets,function(ps) {
    mval <- mval[rownames(mval) %in% ps,,drop=FALSE]
    if (nrow(mval>1)) {
      o <- apply(mval,2,median)
    } else {
      o <- mval
    }
   o
  },mc.cores=cores)
  ag <- do.call(rbind,ag)
  res <- main(mval=ag,design=design,sets=genesets,cores=cores )
}

tic()
hcy_gsmea <- gsmea(mval=mval,design=design,probesets=sets,genesets=genesets,cores=8)
toc()

hcy_gsmea$res <- subset(hcy_gsmea$res,nprobes>=5)
hcy_gsmea$res$FDR <- p.adjust(hcy_gsmea$res$PValue,method="fdr")

head(hcy_gsmea$res,20) %>%
  kbl(caption = "Pathways found using the agg,lim,fry method") %>%
  kable_styling("hover", full_width = F)

```

## Whole gene analysis

```{r,wg}
# set cores to used for parallel execution
CORES= detectCores()

dm <- read.table("dma3a.tsv.gz")
dm <- dm[,4:9]
dm <- merge(myann,dm,by=0)
rownames(dm) <- dm[,1]
dm[,1] = NULL

calc_sc <- function(dm) {
  gn <- unique(unlist(strsplit( dm$UCSC_RefGene_Name ,";")))
  gnl <- strsplit( dm$UCSC_RefGene_Name ,";")
  gnl <- mclapply(gnl,unique,mc.cores=CORES)
  dm$UCSC_RefGene_Name <- gnl
  l <- mclapply(1:nrow(dm), function(i) {
    a <- dm[i,]
    len <- length(a[[1]][[1]])
    tvals <- as.numeric(rep(a["t"],len))
    genes <- a[[1]][[1]]
    data.frame(genes,tvals)
  },mc.cores=CORES)
  df <- do.call(rbind,l)
  gme_res <- mclapply( 1:length(gn), function(i) {
    g <- gn[i]
    tstats <- df[which(df$genes==g),"tvals"]
    myn <- length(tstats)
    mymean <- mean(tstats)
    mymedian <- median(tstats)
    wtselfcont <- wilcox.test(tstats)
    res <- c("gene"=g,"nprobes"=myn,"mean"=mymean,"median"=mymedian,
      "p-value(sc)"=wtselfcont$p.value)
  } , mc.cores=CORES )
  gme_res_df <- do.call(rbind, gme_res)
  rownames(gme_res_df) <- gme_res_df[,1]
  gme_res_df <- gme_res_df[,-1]
  tmp <- apply(gme_res_df,2,as.numeric)
  rownames(tmp) <- rownames(gme_res_df)
  gme_res_df <- as.data.frame(tmp)
  gme_res_df$sig <- -log10(gme_res_df[,4])
  gme_res_df <- gme_res_df[order(-gme_res_df$sig),]
  gme_res_df$`fdr(sc)` <- p.adjust(gme_res_df$`p-value(sc)`)
  out <- list("df"=df,"gme_res_df"=gme_res_df)
  return(out)
}

tic()
gme_res_wholegene <- calc_sc(dm)
time2 <- toc() #38 44 41 40 44 41 40
time2
df <- gme_res_wholegene[[1]]
res <- gme_res_wholegene[[2]]
write.table(res,file="gmea_wholegene.tsv")

```

## Table of top results.

```{r, toptable}

head(res,50) %>%
  kbl(caption = "Top significant genes with GMEA") %>%
  kable_styling("hover", full_width = F)

```

## Volcano plots

```{r,volcano1}

# volcano selfcont
sig <- subset(res,`fdr(sc)` < 0.05)
plot(res$median , -log10(res$`p-value(sc)`) ,
  xlab="effect size (mean t-stat)", ylab="-log10(p-value)",
  pch=19, cex=0.5, col="gray",main="self contained test")
grid()
points(sig$median , -log10(sig$`p-value(sc)`) ,
  pch=19, cex=0.5, col="red")

```

## Probe bias

```{r,probe_bias1}

plot(res$nprobes,res$sig,log="x",ylim=c(0,50),pch=19,cex=0.6)
points(sig$nprobes,sig$sig,col="red",pch=19,cex=0.62)
MIN = min(sig$nprobes)
LEFT = nrow(subset(res,nprobes<MIN))
RIGHT = nrow(subset(res,nprobes>MIN))
SIG = nrow(sig)
TOT = nrow(res)
HEADER <- paste(TOT, "genes in total.", SIG, "with FDR<0.05.",
  RIGHT, "well covered and", LEFT, "poorly covered")
mtext(HEADER)
abline(v=MIN,lty=2)

```

## Boxplots

Boxplots smallest pvalue.

```{r,box1}

par(mfrow=c(1,2))
n=50
# self contained

gs <- head(rownames(res),50)
tstats <- lapply(gs, function(g) {
  df[which(df$genes==g),"tvals"]
})
names(tstats) <- gs

tstats <- tstats[order(unlist(lapply(tstats,median)))]
boxplot(tstats,horizontal=TRUE,las=1,
  main="smallest p-val(selfcont)",cex.axis=0.6,
  xlab="t-statistic")
grid()

n=50
# effect size median
sig <- subset(res,`fdr(sc)` < 0.05)
gs <- head(rownames(sig[order(-abs(sig$median)),]),n)

if ( length(gs) >2 ) {
  tstats <- lapply(gs, function(g) {
    df[which(df$genes==g),"tvals"]
  })
  names(tstats) <- gs
  tstats <- tstats[order(unlist(lapply(tstats,median)))]
  boxplot(tstats,horizontal=TRUE,las=1,
    main="biggest effect size(median)",cex.axis=0.6,
    xlab="t-statistic")
  grid()
} else {
  plot(1)
  mtext("too few significant genes found")
}

```

## Mitch

```{r,dm1_mitch}

dmscore <- data.frame( res$median * res$sig)

rownames(dmscore) <- rownames(res)

colnames(dmscore) <- "metric"

if ( ! file.exists("ReactomePathways.gmt") ) {
  download.file("https://reactome.org/download/current/ReactomePathways.gmt.zip",
    destfile="ReactomePathways.gmt.zip")
  unzip("ReactomePathways.gmt.zip")
}
file.info("ReactomePathways.gmt")
genesets <- gmt_import("ReactomePathways.gmt")
length(genesets)

mres <- mitch_calc(x=dmscore, genesets=genesets, priority="effect")

head(mres$enrichment_result,20) %>%
  kbl(caption = "Top enriched gene sets with GMEA-Mitch", row.names = FALSE) %>%
  kable_styling("hover", full_width = F)

mitch_report(mres,outfile="dma3a_mitch.html",overwrite=TRUE)

```

## Session Information

For reproducibility.

```{r,sessioninfo}

sessionInfo()

```