### 2023-09-14 figs/step_01b_test_gometh.Rmd

---
title: "GOmeth test"
author: "hcy_meth meta-analysis group"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Here we will run test enrichment analysis approaches for undderstanding the effect of elevated
homocysteine on blood methylomes.
Homocysteine is thought to be an inhibitor of methyltransferases and hyperhomocysteine is
associated with overall reduced genomic methylation.
Previously we analysed the B-PROOF study data which enabled us to identify probes whose methylation correlatesor anti-correlated with homocysteine level.

The two enrichment analysis approaches we will use are:

* `gsameth`: an ORA approach that addressed the biases inherent with infinium array analysis.

* `gmea`: an experimental application of functional class scoring for infinium analysis.

```{r,libs}

suppressPackageStartupMessages({
  library("missMethyl")
  library("kableExtra")
  library("mitch")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylation450kanno.ilmn12.hg19")
  library("tictoc")
})

```

Load the data.

```{r,load}

dm <- read.table("dma3a.tsv")
rownames(dm) <- dm$Row.names
dm$Row.names=NULL

head(dm, 10) %>% kbl() %>% kable_paper("hover", full_width = F)

```

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,gsdl}

gs_entrez <- gmt_import("c2.cp.reactome.v2023.1.Hs.entrez.gmt")

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

## ORA with GSAMETH - parth of the MissMethyl package

Now perform ORA analysis using gsameth with the top 500 probes.

```{r,gsameth1}

sigup <- rownames(head(subset(dm,logFC>0),1000))
gsaup <- gsameth(sig.cpg=head(sigup,500), all.cpg=rownames(dm), collection=gs_entrez)
gsaup <- gsaup[order(gsaup$P.DE),]
head(gsaup,10) %>% kbl(caption="Top hypermethylated pathways") %>% kable_paper("hover", full_width = F)

sigdn <- rownames(head(subset(dm,logFC<0),1000))
gsadn <- gsameth(sig.cpg=head(sigdn,500), all.cpg=rownames(dm), collection=gs_entrez)
gsadn <- gsadn[order(gsadn$P.DE),]
head(gsadn,10) %>% kbl(caption="Top hypomethylated pathways") %>% kable_paper("hover", full_width = F)

```

## GMEA

Now with GMEA.

Start with gathering the annotation set.

```{r,anno}

ann450k <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
myann <- data.frame(ann450k[,c("UCSC_RefGene_Name","Regulatory_Feature_Group")])
promoters <- grep("Prom",myann$Regulatory_Feature_Group)

```

Histogram of limma t-statistics.

```{r,hist}

dm <- dm[,c("UCSC_RefGene_Name","t")]
hist(dm$t,breaks=seq(from=-6,to=6,by=1)) ; grid()

```

Aggregate probe t-scores to gene level scores.

The object generated has two parts:

1. `df`: a simple table of the gene name and the median tvalue

2. `gme_res_df`: results of 1-sample Wilcox test.

```{r,agg_wt}

CORES= detectCores()

agg <- function(dm) {
  gn <- unique(unlist(strsplit( dm$UCSC_RefGene_Name ,";")))
  gnl <- strsplit( dm$UCSC_RefGene_Name ,";")
  gnl <- mclapply(gnl,unique,mc.cores=CORES)
  dm$UCSC_RefGene_Name <- gnl
  l <- mclapply(1:nrow(dm), function(i) {
    a <- dm[i,]
    len <- length(a[[1]][[1]])
    tvals <- as.numeric(rep(a[2],len))
    genes <- a[[1]][[1]]
    data.frame(genes,tvals)
  },mc.cores=CORES)
  df <- do.call(rbind,l)
  gme_res <- mclapply( 1:length(gn), function(i) {
    g <- gn[i]
    tstats <- df[which(df$genes==g),"tvals"]
    myn <- length(tstats)
    mymean <- mean(tstats)
    mymedian <- median(tstats)
    wtselfcont <- wilcox.test(tstats)
    res <- c("gene"=g,"nprobes"=myn,"mean"=mymean,"median"=mymedian,
      "p-value(sc)"=wtselfcont$p.value)
  } , mc.cores=CORES )
  gme_res_df <- do.call(rbind, gme_res)
  rownames(gme_res_df) <- gme_res_df[,1]
  gme_res_df <- gme_res_df[,-1]
  tmp <- apply(gme_res_df,2,as.numeric)
  rownames(tmp) <- rownames(gme_res_df)
  gme_res_df <- as.data.frame(tmp)
  gme_res_df$sig <- -log10(gme_res_df[,4])
  gme_res_df <- gme_res_df[order(-gme_res_df$sig),]
  gme_res_df$`fdr(sc)` <- p.adjust(gme_res_df$`p-value(sc)`)
  out <- list("df"=df,"gme_res_df"=gme_res_df)
  return(out)
}

tic()
dmagg <- agg(dm)
toc()
head(dmagg$gme_res_df) %>% kbl(caption="top gmea genes (Wilcox)") %>% kable_paper("hover", full_width = F)

```

Does 1-sample t-test work better?

```{r,agg_tt}

agg2 <- function(dm) {
  gn <- unique(unlist(strsplit( dm$UCSC_RefGene_Name ,";")))
  gnl <- strsplit( dm$UCSC_RefGene_Name ,";")
  gnl <- mclapply(gnl,unique,mc.cores=CORES)
  dm$UCSC_RefGene_Name <- gnl
  l <- mclapply(1:nrow(dm), function(i) {
    a <- dm[i,]
    len <- length(a[[1]][[1]])
    tvals <- as.numeric(rep(a["t"],len))
    genes <- a[[1]][[1]]
    data.frame(genes,tvals)
  },mc.cores=CORES)
  df <- do.call(rbind,l)
  keep <- names(which(table(df$genes)>1))
  df <- df[df$genes %in% keep,]
  gn <- unique(df$genes)
  gme_res <- lapply( 1:length(gn), function(i) {
    g <- gn[i]
    tstats <- df[which(df$genes==g),"tvals"]
    myn <- length(tstats)
    mymean <- mean(tstats)
    mymedian <- median(tstats)
    tselfcont <- t.test(tstats)
    res <- c("gene"=g,"nprobes"=myn,"mean"=mymean,"median"=mymedian,
      "p-value(sc)"=tselfcont$p.value)
  } )
  gme_res_df <- do.call(rbind, gme_res)
  rownames(gme_res_df) <- gme_res_df[,1]
  gme_res_df <- gme_res_df[,-1]
  tmp <- apply(gme_res_df,2,as.numeric)
  rownames(tmp) <- rownames(gme_res_df)
  gme_res_df <- as.data.frame(tmp)
  gme_res_df$sig <- -log10(gme_res_df[,4])
  gme_res_df <- gme_res_df[order(-gme_res_df$sig),]
  gme_res_df$`fdr(sc)` <- p.adjust(gme_res_df$`p-value(sc)`)
  out <- list("df"=df,"gme_res_df"=gme_res_df)
  return(out)
}

tic()
dmagg2 <- agg2(dm)
toc()
head(dmagg2$gme_res_df) %>% kbl(caption="top gmea genes (t-test)") %>% kable_paper("hover", full_width = F)

```

This results shows that wilcox test executes faster but 1-sample t-test gives more significant findings.

Check that the distribution is even.

```{r,tmeddist}

hist(dmagg2$gme_res_df$median,breaks=30) ; grid()

```

If the distribution is even, then established enrichment analysis can be performed.
Options include FGSEA and mitch.

Here I will try downstream analysis with mitch.

```{r,mitch1}

m <- dmagg2$gme_res_df

m <- m[,"median",drop=FALSE]

mres <- mitch_calc(x=m, genesets=gs_symbols, priority="effect")

head(mres$enrichment_result,20) %>% kbl(caption="top mitch pathways") %>% kable_paper("hover", full_width = F)
hist(mres$enrichment_result$s.dist) ; grid()

tbl <- mres$enrichment_result

sig <- subset(tbl,p.adjustANOVA<0.05)
nrow(sig)

plot(tbl$s.dist,-log10(tbl$pANOVA),pch=19,col="gray")
grid()
points(sig$s.dist,-log10(sig$pANOVA),pch=19,col="red")

msig <- sig

```

Some interesting pathways found.

There was no overlap with `gsameth()`.

Interestingly many gene sets exhibit increased methylation.

There could be a directional bias, so I will also conduct a 1-sample t-test.

```{r,t_enrichment_test1}

ttenrich <- function(m,genesets) {
  res <- mclapply( 1:length(genesets), function(i) {
    gs <- genesets[i]
    name <- names(gs)
    n_members <- length(which(rownames(m) %in% gs[[1]]))
    if ( n_members > 4 ) {
      tstats <- m[which(rownames(m) %in% gs[[1]]),]
      myn <- length(tstats)
      mymean <- mean(tstats)
      mymedian <- median(tstats)
      wt <- t.test(tstats)
      res <- c(name,myn,mymean,mymedian,wt$p.value)
    }
  } , mc.cores = CORES)
  res_df <- do.call(rbind, res)
  rownames(res_df) <- res_df[,1]
  res_df <- res_df[,-1]
  colnames(res_df) <- c("n_genes","t_mean","t_median","p.value")
  tmp <- apply(res_df,2,as.numeric)
  rownames(tmp) <- rownames(res_df)
  res_df <- tmp
  res_df <- as.data.frame(res_df)
  res_df <- res_df[order(res_df$p.value),]
  res_df$logp <- -log10(res_df$p.value )
  res_df$fdr <- p.adjust(res_df$p.value,method="fdr")
  res_df[order(abs(res_df$p.value)),]
  return(res_df)
}

gse_res_df <- ttenrich(m=m,genesets=gs_symbols)

head(gse_res_df,30) %>%
  kbl(caption = "Top significant genesets") %>%
  kable_paper("hover", full_width = F)

head(gse_res_df[order(-abs(gse_res_df$t_median)),],30) %>%
  kbl(caption = "Top effect size genesets") %>%
  kable_paper("hover", full_width = F)

sig <- subset(gse_res_df,fdr<0.05)
nrow(sig)

tsig <- sig

```

## Session information

For reproducibility.

```{r,session}

sessionInfo()

```