### 2023-09-14 old/GSE158433/GSE158422_gmea.Rmd

---
title: "GSE158433: GMEA of DNA methylation differences in lung squamous cell carcinoma"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

The purpose of this analysis is to explore different enrichment approaches to the
cancer methylation data: Infinium EPIC array (GSE158422) data for matched control and cancer tissues.

We will specifically be testing the GMEA approach versus existing approaches:

* 1-way Wilcox text

* 1-way t-test

* FGSEA test

* roast test

After each of these tests, downstream GSEA can be conducted.

```{r,packages}

suppressPackageStartupMessages({
  library("plyr")
  library("R.utils")
  library("missMethyl")
  library("limma")
  library("DMRcate")
  library("DMRcatedata")
  library("topconfects")
  library("minfi")
  library("IlluminaHumanMethylation450kmanifest")
  library("RColorBrewer")
  library("IlluminaHumanMethylation450kanno.ilmn12.hg19")
  library("GEOquery")
  library("eulerr")
  library("plyr")
  library("gplots")
  library("reshape2")
  library("forestplot")
  library("beeswarm")
  library("RCircos")
  library("qqman")
  library("ENmix")
  library("tictoc")
  library("mitch")
  library("kableExtra")
  library("fgsea")
})

source("../meth_functions.R")
CORES=8

```

## Obtaining array annotations

```{r,annotation}

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name",
  "Regulatory_Feature_Group","Islands_Name","Relation_to_Island")])
promoters <- grep("Prom",myann$Regulatory_Feature_Group)

```

Gene probe sets

```{r,sets}

gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
sets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))

hist(unlist(lapply(sets,length)),xlim=c(0,200),breaks=500,xlab="set size",main="probes per gene")

```

## Load array result

GSE158422 is the accession number for the array data.

```{r,arraydataload}

#GSE158422 <- readRDS("GSE158422.rds")
dm <- read.table("GSE158422_limma.tsv",header=TRUE,sep="\t")

head(dm,50) %>%
  kbl(caption = "Top significant probes with limma") %>%
  kable_paper("hover", full_width = F)

rownames(dm) <- dm[,1]
dm <- dm[,c("UCSC_RefGene_Name","t")]

hist(dm$t)

```

## GMEA whole gene - wilcox test

```{r,gmea_wg}

tic()
resw <- calc_sc(dm)
toc() #16 cores 129 sec
res2 <- resw[[2]]

head(res2,20) %>%
  kbl(caption = "Top significant genes with GMEA") %>%
  kable_paper("hover", full_width = F)

sig <- subset(res2,`fdr(sc)`<0.05)

es <- sig[order(-abs(sig$median)),]

head(es,20) %>%
  kbl(caption = "Top effect size probes with GMEA") %>%
  kable_paper("hover", full_width = F)

gmea_volc(res2)
gmea_boxplot(res)
res2_wg <- res2

```

Investigate how probe number influences significance.

```{r,probebias}

plot(res2$nprobes,res2$sig,log="x",ylim=c(0,50),pch=19,cex=0.6)
points(sig$nprobes,sig$sig,col="red",pch=19,cex=0.62)
MIN = min(sig$nprobes)
LEFT = nrow(subset(res2,nprobes<MIN))
RIGHT = nrow(subset(res2,nprobes>MIN))
SIG = nrow(sig)
TOT = nrow(res2)
HEADER <- paste(TOT, "genes in total.", SIG, "with FDR<0.05.",
  RIGHT, "well covered and", LEFT, "poorly covered")
mtext(HEADER)
abline(v=MIN,lty=2)

```

## GMEA whole gene - t.test

```{r,gmea_t}

dm <- read.table("GSE158422_limma.tsv",header=TRUE,sep="\t")
rownames(dm) <- dm[,1]
dm <- dm[,c("UCSC_RefGene_Name","t")]

tic()
rest <- calc_sc2(dm)
toc() #16 cores 252 sec

res2 <- rest[[2]]

head(res2,20) %>%
  kbl(caption = "Top significant genes with GMEA-t") %>%
  kable_paper("hover", full_width = F)

sig <- subset(res2,`fdr(sc)`<0.05)

es <- sig[order(-abs(sig$median)),]

head(es,20) %>%
  kbl(caption = "Top effect size probes with GMEA") %>%
  kable_paper("hover", full_width = F)

gmea_volc(res2)
gmea_boxplot(res)
res2_t <- res2

plot(res2$nprobes,res2$sig,log="x",ylim=c(0,50),pch=19,cex=0.6)
points(sig$nprobes,sig$sig,col="red",pch=19,cex=0.62)
MIN = min(sig$nprobes)
LEFT = nrow(subset(res2,nprobes<MIN))
RIGHT = nrow(subset(res2,nprobes>MIN))
SIG = nrow(sig)
TOT = nrow(res2)
HEADER <- paste(TOT, "genes in total.", SIG, "with FDR<0.05.",
  RIGHT, "well covered and", LEFT, "poorly covered")
mtext(HEADER)
abline(v=MIN,lty=2)

```

## GMEA using FGSEA test

```{r,gmea_fgsea}

head(dm)
tstats <- dm$t
names(tstats) <- rownames(dm)

tic()
fgseaRes <- fgsea(pathways = sets,
                  stats    = tstats,
                  minSize  = 5,
                  nproc = 16)
toc()

fgseaRes <- as.data.frame(fgseaRes)[,1:7]
fgseaRes <- fgseaRes[order(fgseaRes$pval),]

head(fgseaRes,20) %>%
  kbl(caption = "Top significant genes with GMEA-FGSEA") %>%
  kable_paper("hover", full_width = F)

sig <- subset(fgseaRes,padj<0.05)
SIG = nrow(sig)
UP = nrow(subset(sig,ES>0))
DN = nrow(subset(sig,ES<0))

plot(fgseaRes$ES,-log10(fgseaRes$pval))
points(sig$ES,-log10(sig$pval),col="red")
mtext(paste(SIG,"signifiant genes,",UP,"up,",DN,"down"))

plot(fgseaRes$size,-log10(fgseaRes$pval),log="x")
points(sig$size,-log10(sig$pval),col="red")
mtext(paste(SIG,"signifiant genes,",UP,"up,",DN,"down"))

```

## Using CAMERA

Need to add volcano and nprobe charts here.

```{r,cam_func}

stat <- dm$t
names(stat) <- rownames(dm)

# parallel approach
CORES=detectCores()
#sets <- head(sets,100000)
sets_split <- split(sets,1:CORES)

tic()

cres <- mclapply(X=sets_split, function(mysets) {
  cameraPR(statistic=stat, index=mysets, use.ranks = FALSE, inter.gene.cor=0.01, sort = TRUE)
}, mc.cores=CORES)

cres <- do.call(rbind,cres)
rownames(cres) <- sub("\\.","@", rownames(cres))
rownames(cres) <- sapply(strsplit(rownames(cres),"@"),"[[",2)
cres <- cres[order(cres$PValue),]
g <- rownames(cres)[1]

m <- mclapply(rownames(cres) , function(g) {
  probes <- sets[[g]]
  scores <- stat[which(names(stat) %in% probes)]
  mymedian <- median(scores)
  mymean <- mean(scores)
  c(mymean,mymedian)
},mc.cores=CORES)

mdf <- do.call(rbind,m)
colnames(mdf) <- c("mean","median")
cres <- cbind(mdf,cres)
cres$FDR <- p.adjust(cres$PValue,method="fdr")
toc()

head(subset(cres,Direction=="Up"),25)
head(subset(cres,Direction=="Down"),25)

head(cres,50) %>%
  kbl(caption = "Top significant genes with CAMERA") %>%
  kable_paper("hover", full_width = F)

sig <- subset(cres,FDR<0.05)
SIG <- nrow(sig)

plot(cres$median,-log10(cres$PValue))
points(sig$median,-log10(sig$PValue),col="red")

plot(cres$NGenes,-log10(cres$PValue),log="x")
points(sig$NGenes,-log10(sig$PValue),col="red")

```

## GMEA using ROAST test

ROAST is self contained test using rotations.

## TODO

The goal is to use a self contained test that can handle extensive correlation.

Globaltest, ANCOVA and SAM-GS

GSVA

Q2 (Tian et al. [14])
14. Tian L, Greenberg SA, Kong SW, et al. Discovering statistically significant pathways in expression profiling studies. Proc Natl Acad Sci USA. 2005;102(38):13544–9. [PMC free article] [PubMed] [Google Scholar]

FCS.SC—self-contained version of FCS (Pavlidis et al. [24])
24. Pavlidis P, Qin J, Arango V, et al. Using the gene ontology for microarray data mining: a comparison of methods and application to age effect in human prefrontal cortex. Neurochem Res. 2004;29(6):1213–22. [PubMed] [Google Scholar]

## Mitch whole gene

Methylation only.

```{r,mitch_wg}

genesets <- gmt_import("../ReactomePathways.gmt")

meth <- res2$sig * res2$median
meth <- as.data.frame(meth)
rownames(meth) <- rownames(res2)

mres <- mitch_calc(x=meth, genesets=genesets,priority="effect")

head(mres$enrichment_result,20) %>%
  kbl(caption = "Top differential pathways with GMEA+mitch") %>%
  kable_paper("hover", full_width = F)

```

## Session Information

```{r,session}

sessionInfo()

```