### 2023-09-20 figs/step_01x_gmea2.Rmd
different approaches for fcs

---
title: "Test different aggregation approaches"
author: "hcy_meth meta-analysis group"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: yes
    toc_float: true
    code_folding: hide
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

In this analysis I will try out an approach that preserved per-sample information.
It aggregates methylaiton values for all CpGs into a single median value.

```{r,libs}

suppressPackageStartupMessages({
  library("missMethyl")
  library("kableExtra")
  library("mitch")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylation450kanno.ilmn12.hg19")
  library("tictoc")
  library("limma")
})

CORES=12

```

## Annotation data

Start with gathering the annotation set.

```{r,anno}

ann450k <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
myann <- data.frame(ann450k[,c("UCSC_RefGene_Name","Regulatory_Feature_Group")])
promoters <- grep("Prom",myann$Regulatory_Feature_Group)

```

## Load methylation data

```{r,load}

mx <- readRDS("dm3_mx.Rds")
design <- readRDS("dm3_design.Rds")

```

## Load pathways

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,gsdl}

gs_entrez <- gmt_import("c2.cp.reactome.v2023.1.Hs.entrez.gmt")

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

## Join

Determine the median of probe values

```{r,med1}

gn <- unique(unlist(strsplit( myann$UCSC_RefGene_Name ,";")))
gnl <- strsplit( myann$UCSC_RefGene_Name ,";")
gnl <- mclapply(gnl,unique,mc.cores=CORES)
myann$gnl <- gnl

mymed <- function(g) {
  probes <- rownames(myann[grep(g,myann$gnl),])
  rows <- which(rownames(mx) %in% probes)
  if ( length(rows) > 1 ) {
    b <- mx[rows,]
    med <- apply(b,2,median)
    med <- matrix(med,nrow=1)
    colnames(med) <- colnames(b)
    rownames(med) <- g
    return(med)
  }
}

med <- mclapply(gn,mymed,mc.cores=CORES)
med <- med[lapply(med,length)>0]
medf <- do.call(rbind,med)

```

Now aggregate by median for gene sets.

```{r,med2}

i=1
mymed2 <- function(mx,gsets,i) {
  nm <- names(gsets)[i]
  gs <- gsets[[i]]
  med <- matrix(apply(mx[which(rownames(mx) %in% gs),],2,median),nrow=1)
  colnames(med) <- colnames(mx)
  rownames(med) <- nm
  return(med)
}

med2 <- mclapply(1:length(gs_symbols), function(i) {
  mymed2(mx=medf,gsets=gs_symbols,i=i)
},mc.cores=CORES)

medf2 <- do.call(rbind,med2)

```

## Limma differential methylation

Then do a limma test.
See if it is more robust than other approaches.

```{r,limma1}

fit.reduced <- lmFit(medf,design)
fit.reduced <- eBayes(fit.reduced)
summary(decideTests(fit.reduced))
dm <- topTable(fit.reduced,coef=4, number = Inf)

#write.table(dma3a,file="dma3a.tsv")

head(dm, 30) %>% kbl() %>% kable_paper("hover", full_width = F)

```

## Mitch

```{r,mitchtest}

y <- mitch_import(dm, DEtype="limma")

res <- mitch_calc(y, genesets=gs_symbols, priority="effect",cores=CORES)

sig <- subset(res$enrichment_result,p.adjustANOVA<0.05)

nrow(subset(sig,s.dist>0))

nrow(subset(sig,s.dist<0))

head(subset(sig,s.dist>0),20) %>%
  kbl(caption="hypermethylated pathways") %>%
  kable_paper("hover", full_width = F)

head(subset(sig,s.dist<0),20) %>%
  kbl(caption="hypomethylated pathways") %>%
  kable_paper("hover", full_width = F)

```

## 1-sample t-test

```{r,ttest}

ttenrich <- function(m,genesets,cores=1) {
  res <- mclapply( 1:length(genesets), function(i) {
    gs <- genesets[i]
    name <- names(gs)
    n_members <- length(which(rownames(m) %in% gs[[1]]))
    if ( n_members > 4 ) {
      tstats <- m[which(rownames(m) %in% gs[[1]]),]
      myn <- length(tstats)
      mymean <- mean(tstats)
      mymedian <- median(tstats)
      wt <- t.test(tstats)
      res <- c(name,myn,mymean,mymedian,wt$p.value)
    }
  } , mc.cores = cores)
  res_df <- do.call(rbind, res)
  rownames(res_df) <- res_df[,1]
  res_df <- res_df[,-1]
  colnames(res_df) <- c("n_genes","t_mean","t_median","p.value")
  tmp <- apply(res_df,2,as.numeric)
  rownames(tmp) <- rownames(res_df)
  res_df <- tmp
  res_df <- as.data.frame(res_df)
  res_df <- res_df[order(res_df$p.value),]
  res_df$logp <- -log10(res_df$p.value )
  res_df$fdr <- p.adjust(res_df$p.value,method="fdr")
  res_df[order(abs(res_df$p.value)),]
  return(res_df)
}

m <- as.data.frame(dm$t)
rownames(m) <- rownames(dm)
colnames(m) <- "t"

tres <- ttenrich(m=m,genesets=gs_symbols,cores=CORES)

sig <- subset(tres,fdr<0.05)

nrow(sig)

nrow(subset(sig,t_median>0))

head(sig[order(-sig$t_median),],20) %>%
  kbl(caption="hypermethylated pathways") %>%
  kable_paper("hover", full_width = F)

nrow(subset(sig,t_median<0))

head(sig[order(sig$t_median),],20) %>%
  kbl(caption="hypomethylated pathways") %>%
  kable_paper("hover", full_width = F)

```

## Todo

Compare mitch and t-test approaches.

## Session information

For reproducibility.

```{r,session}

sessionInfo()

```