### 2023-09-27 figs/GSE158422_rna_int.Rmd

---
title: "GSE158422: DNA methylation in Lung squamous cell carcinoma - RNA expression and integration"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

```{r,packages}

suppressPackageStartupMessages({
  library("DESeq2")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("HGNChelper")
  library("gplots")
})

CORES=12

```

## Load annotations

```{r,annotation}

load("GSE158422_gmea.Rdata")

```

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,gsdl}

gs_entrez <- gmt_import("c2.cp.reactome.v2023.1.Hs.entrez.gmt")

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

## Load RNA expression

From GEO https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE158420
GSE158420_counts.txt.gz

## DESEQ2

```{r,deseq1}

x <- read.table("GSE158420_counts.txt.gz",row.names=NULL)
dim(x)
x <- aggregate(. ~ row.names,x,sum)
dim(x)

#RUFKM
head(x)

```

There are some gene name errors, so I will use HGNChelper to repair them.

```{r,gene_name_fix}

human=x$row.names
fix <- checkGeneSymbols(human)
x$gene <- fix$Suggested.Symbol
x$row.names=NULL
x <- aggregate(. ~ gene,x,sum)
dim(x)
row.names(x)=x$gene
x$gene=NULL
head(x)

```

Now do DESeq2

```{r,deseq2}

patient <- head(as.character(unlist(lapply(1:ncol(x),function(i) {c(i,i)}))),ncol(x))
tumor <- rep(c(0,1),ncol(x)/2)
ss <- data.frame(patient,tumor)

dds <- DESeqDataSetFromMatrix(countData = x , colData=ss, design = ~ patient + tumor)
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <-cbind(as.data.frame(z),assay(vsd))
dge <- as.data.frame(zz[order(zz$pvalue),])

sig <- subset(dge,padj<0.05)
SIG = nrow(sig)
DN = nrow(subset(sig,log2FoldChange<0))
UP = nrow(subset(sig,log2FoldChange>0))
HEADER = paste("mRNA", SIG , "DEGs,", UP ,"upregulated,", DN, "downregulated")
# smear
plot(log2(dge$baseMean),dge$log2FoldChange,cex=0.6,cex.axis=1.2,cex.lab=1.3,
  xlab="log2 base mean",
  ylab="log2 fold change" ,pch=19,col="#838383")
points(log2(sig$baseMean),sig$log2FoldChange,cex=0.6,pch=19,col="red")
mtext((HEADER),cex=1.2)
top <- head(sig,20)
# volcano
plot(dge$log2FoldChange, -log2(dge$pvalue) ,cex=0.6, cex.lab=1.3,cex.axis=1.2,
 ,xlab="log2 fold change", ylab="-log2 p-value" ,pch=19,col="#838383")
points(sig$log2FoldChange, -log2(sig$pvalue),cex=0.6,pch=19,col="red")
mtext((HEADER),cex=1.2)
# top N gene heatmap
colCols <- rep(c(0,1),ncol(x)/2)
colCols <- gsub("0","gray",colCols)
colCols <- gsub("1","black",colCols)

colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2(  as.matrix(dge[1:50,c(7:ncol(dge))]), col=colfunc(25),scale="row",
  ColSideColors=colCols ,
  trace="none",margins = c(6,10), cexRow=.6, cexCol=.5, main="Top 50 genes")

```

Pathway enrichment with mitch.

```{r,mitch}

y <- mitch_import(dge, DEtype="deseq2")

res <- mitch_calc(y, genesets=gs_symbols, priority="effect",cores=CORES)

mres <- res$enrichment_result
msig <- subset(mres,p.adjustANOVA<0.05)

head(subset(mres,p.adjustANOVA<0.05 & s.dist>0),20) %>%
  kbl(caption="mRNA upregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

head(subset(mres,p.adjustANOVA<0.05 & s.dist<0),20) %>%
  kbl(caption="mRNA downregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

```

## Load methylation enrichment data and calculate overlap

Use Euler diagrams to find the similarity.
As medians are normally distributed about the zero, one would expect equal numbers of up and down-
methylated gene sets.

```{r,overlap1}

load("GSE158422_gmea.Rdata")

head(mres)

head(alres)

head(lares)

msig_up <- subset(msig,s.dist>0)$set
msig_dn <- subset(msig,s.dist<0)$set

lasig_up <- rownames(subset(lasig,t_median>0))
lasig_dn <- rownames(subset(lasig,t_median<0))

alsig_up <- rownames(subset(alsig,t_median>0))
alsig_dn <- rownames(subset(alsig,t_median<0))

v1 <- list("LA up"=lasig_up, "LA dn"=lasig_dn,
  "RNA up"=msig_up,"RNA dn"=msig_dn)
plot(euler(v1),quantities = TRUE)

v1 <- list("AL up"=alsig_up,"AL dn"=alsig_dn,
  "RNA up"=msig_up,"RNA dn"=msig_dn)
plot(euler(v1),quantities = TRUE)

```

## Load promoter methylation enrichment data and calculate overlap

```{r,overlap2}

head(alpres)

head(lapres)

lapsig_up <- rownames(subset(lapsig,t_median>0))
lapsig_dn <- rownames(subset(lapsig,t_median<0))

alpsig_up <- rownames(subset(alpsig,t_median>0))
alpsig_dn <- rownames(subset(alpsig,t_median<0))

v1 <- list("LAP up"=lapsig_up, "LAP dn"=lapsig_dn,
  "RNA up"=msig_up,"RNA dn"=msig_dn)
plot(euler(v1),quantities = TRUE)

v1 <- list("ALP up"=alpsig_up,"ALP dn"=alpsig_dn,
  "RNA up"=msig_up,"RNA dn"=msig_dn)
plot(euler(v1),quantities = TRUE)

```

## Conclusion

Although there were some differentially methylated pathways analysed with different methods,
they were not concordant with differential expression.

## Session information

```{r,save}

sessionInfo()

```