### 2023-09-30 figs/GSE158422_split.Rmd
collect results

---
title: "GSE158422: DNA methylation in Lung squamous cell carcinoma - gene methylation enrichment analysis"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Here we are splitting the dataset and testing concordance between LA and AL approaches.

```{r,packages}

suppressPackageStartupMessages({
  library("limma")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
})

CORES=12

```

## Load data

* annotations

* probe sets

* gene sets

* design matrix

* mval matrix

```{r,annotation}

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","Regulatory_Feature_Group","Islands_Name","Relation_to_Island")])

gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
sets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))

summary(unlist(lapply(sets,length)))

genesets <- gmt_import("https://ziemann-lab.net/public/gmea_prototype/ReactomePathways.gmt")

if (!file.exists("GSE158422_design.rds")) {
  download.file("https://ziemann-lab.net/public/gmea_prototype/GSE158422_design.rds", "GSE158422_design.rds")
}
design <- readRDS("GSE158422_design.rds")

if (!file.exists("GSE158422_design.rds")) {
 download.file("https://ziemann-lab.net/public/gmea_prototype/GSE158422_mx.rds","GSE158422_mx.rds")
}
mval <- readRDS("GSE158422_mx.rds")

boxplot(list("normal"=matrix(colMeans(mval),ncol=2)[,2],"tumor"=matrix(colMeans(mval),ncol=2)[,1]),
  main="mean probe methylation mval")

```

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,gsdl}

gs_entrez <- gmt_import("c2.cp.reactome.v2023.1.Hs.entrez.gmt")

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

# GMEA FUNCTIONS

## LA

This process runs limma first and then aggregates the results before doing an enrichment test.

```{r,lafunc}

# limma
runlimma <- function(mval,design,myann) {
  fit.reduced <- lmFit(mval,design)
  fit.reduced <- eBayes(fit.reduced)
  summary(decideTests(fit.reduced))
  dm <- topTable(fit.reduced,coef=4, number = Inf)
  dm <- merge(myann,dm,by=0)
  dm <- dm[order(dm$P.Value),]
  rownames(dm) <- dm$Row.names
  dm$Row.names=NULL
  return(dm)
}

# aggregate
agg <- function(dm,cores=1) {
  gn <- unique(unlist(strsplit( dm$UCSC_RefGene_Name ,";")))
  gnl <- strsplit( dm$UCSC_RefGene_Name ,";")
  gnl <- mclapply(gnl,unique,mc.cores=cores)
  dm$UCSC_RefGene_Name <- gnl
  l <- mclapply(1:nrow(dm), function(i) {
    a <- dm[i,]
    len <- length(a[[1]][[1]])
    tvals <- as.numeric(rep(a["t"],len))
    genes <- a[[1]][[1]]
    data.frame(genes,tvals)
  },mc.cores=cores)
  df <- do.call(rbind,l)
  keep <- names(which(table(df$genes)>1))
  df <- df[df$genes %in% keep,]
  gn <- unique(df$genes)
  gme_res <- lapply( 1:length(gn), function(i) {
    g <- gn[i]
    tstats <- df[which(df$genes==g),"tvals"]
    myn <- length(tstats)
    mymean <- mean(tstats)
    mymedian <- median(tstats)
    if ( length(tstats) > 2 ) {
      ttest <- t.test(tstats)
      pval <- ttest$p.value
    } else {
      pval = 1
    }
    res <- c("gene"=g,"nprobes"=myn,"mean"=mymean,
      "median"=mymedian, pval=pval)
  } )
  gme_res_df <- do.call(rbind, gme_res)
  rownames(gme_res_df) <- gme_res_df[,1]
  gme_res_df <- gme_res_df[,-1]
  tmp <- apply(gme_res_df,2,as.numeric)
  rownames(tmp) <- rownames(gme_res_df)
  gme_res_df <- as.data.frame(tmp)
  gme_res_df$sig <- -log10(gme_res_df[,4])
  gme_res_df <- gme_res_df[order(-gme_res_df$sig),]
  gme_res_df$fdr <- p.adjust(gme_res_df$pval)
  out <- list("df"=df,"gme_res_df"=gme_res_df)
  return(out)
}

# enrich
ttenrich <- function(m,genesets,cores=1) {
  res <- mclapply( 1:length(genesets), function(i) {
    gs <- genesets[i]
    name <- names(gs)
    n_members <- length(which(rownames(m) %in% gs[[1]]))
    if ( n_members > 4 ) {
      tstats <- m[which(rownames(m) %in% gs[[1]]),]
      myn <- length(tstats)
      mymean <- mean(tstats)
      mymedian <- median(tstats)
      wt <- t.test(tstats)
      res <- c(name,myn,mymean,mymedian,wt$p.value)
    }
  } , mc.cores = cores)
  res_df <- do.call(rbind, res)
  rownames(res_df) <- res_df[,1]
  res_df <- res_df[,-1]
  colnames(res_df) <- c("n_genes","t_mean","t_median","pval")
  tmp <- apply(res_df,2,as.numeric)
  rownames(tmp) <- rownames(res_df)
  res_df <- tmp
  res_df <- as.data.frame(res_df)
  res_df <- res_df[order(res_df$pval),]
  res_df$logp <- -log10(res_df$pval )
  res_df$fdr <- p.adjust(res_df$pval,method="fdr")
  res_df[order(abs(res_df$pval)),]
  return(res_df)
}

```

## AL: Aggregate limma

Functions for aggregate-limma-enrich approach.

```{r,alfunc}

# chromosome by chromosome will be much faster
magg <- function(mval,myann,cores=1){
  gn <- unique(unlist(strsplit( myann$UCSC_RefGene_Name ,";")))
  gnl <- strsplit( myann$UCSC_RefGene_Name ,";")
  gnl <- mclapply(gnl,unique,mc.cores=cores)
  myann$gnl <- gnl
  keep <- rownames(subset(myann,UCSC_RefGene_Name!=""))
  mx <- mval[rownames(mval) %in% keep,]
  mymed <- function(g) {
    probes <- rownames(myann[grep(g,myann$gnl),])
    rows <- which(rownames(mx) %in% probes)
    if ( length(rows) > 1 ) {
      b <- mx[rows,]
      med <- apply(b,2,median)
      med <- matrix(med,nrow=1)
      colnames(med) <- colnames(b)
      rownames(med) <- g
      return(med)
    }
  }
  med <- mclapply(gn,mymed,mc.cores=cores)
  med <- med[lapply(med,length)>0]
  medf <- do.call(rbind,med)
  return(medf)
}

chragg <- function(mval,myann,cores=1){
  annodf <- as.data.frame(anno)
  keep <- rownames(subset(myann,UCSC_RefGene_Name!=""))
  mx <- mval[rownames(mval) %in% keep,]
  chrs <- unique(anno$chr)
  myorder <- unlist(lapply(chrs,function(mychr) { nrow( annodf[annodf$chr==mychr,] ) } ))
  chrs <- chrs[order(-myorder)]
  leadercores <- floor(sqrt(cores))
  workercores <- ceiling(sqrt(cores))
  chrmedf <- mclapply(chrs,function(chr) {
    chrfrag <- annodf[annodf$chr==chr,]
    chrprobes <-rownames(chrfrag)
    chrmx <- mx[rownames(mx) %in% chrprobes,]
    chranno <- myann[rownames(myann) %in% chrprobes,]
    chrmedf <- magg(mval=chrmx,myann=chranno,cores=workercores)
    return(chrmedf)
  },mc.cores=leadercores)
  medf <- do.call(rbind, chrmedf)
  return(medf)
}

agglimma <- function(medf,design) {
  fit.reduced <- lmFit(medf,design)
  fit.reduced <- eBayes(fit.reduced)
  dmagg <- topTable(fit.reduced,coef=ncol(design), number = Inf)
  nondup <- !duplicated(dmagg$ID)
  dmagg <- dmagg[nondup,]
  rownames(dmagg) <- dmagg$ID
  dmagg$ID = NULL
  return(dmagg)
}

#ttenrich is same

```

## Prep

```{r,prep}

# prep
sex <- as.data.frame(design)$sex
tumor <- as.data.frame(design)$tumor
patient <- as.character(unlist(lapply(1:ncol(mval),function(i) {c(i,i)})))
patient <- head(patient,ncol(mval))
design <- model.matrix(~ patient + tumor )
rownames(design) <- colnames(mval)

```

## LA workflow split and analyse

split 74 samples 37 patients split into 18 + 18
Conduct LA workflow analysis.
Then make overlap.

```{r,laflow}

splitanalyze_la <- function(seed){
  set.seed(seed)
  a <- sample(1:37,18)
  a <- a[order(a)]
  nona <- setdiff(1:37,a)
  set.seed(seed)
  b <- sample(nona,18)
  b <- b[order(b)]
  a <- c( (a*2)-1 ,a*2 )
  a <- a[order(a)]
  b <- c( (b*2)-1 ,b*2 )
  b <- b[order(b)]

  design_a <- design[a,]
  design_a <- design_a[,colSums(design_a)>0]
  mval_a <- mval[,a]
  design_b <- design[b,]
  design_b <- design_b[,colSums(design_b)>0]
  mval_b <- mval[,b]

  # la1
  dm1 <- runlimma(mval_a,design_a,myann)
  dmagg1 <- agg(dm1,cores=CORES)
  m1 <- dmagg1$gme_res_df[,"mean",drop=FALSE]
  lares1 <- ttenrich(m=m1,genesets=gs_symbols,cores=CORES)

  # la2
  dm2 <- runlimma(mval_b,design_b,myann)
  dmagg2 <- agg(dm2,cores=CORES)
  m2 <- dmagg2$gme_res_df[,"mean",drop=FALSE]
  lares2 <- ttenrich(m=m2,genesets=gs_symbols,cores=CORES)
}

lares <- lapply( seq(100,1000,100) ,splitanalyze_la)

lares

```

## AL workflow split and analyse

```{r,alflow}

splitanalyze_al <- function(seed){
  set.seed(seed)
  a <- sample(1:37,18)
  a <- a[order(a)]
  nona <- setdiff(1:37,a)
  set.seed(seed)
  b <- sample(nona,18)
  b <- b[order(b)]
  a <- c( (a*2)-1 ,a*2 )
  a <- a[order(a)]
  b <- c( (b*2)-1 ,b*2 )
  b <- b[order(b)]

  design_a <- design[a,]
  design_a <- design_a[,colSums(design_a)>0]
  mval_a <- mval[,a]
  design_b <- design[b,]
  design_b <- design_b[,colSums(design_b)>0]
  mval_b <- mval[,b]

  #al
  medf1 <- chragg(mval_a,myann,cores=CORES)
  magg1 <- agglimma(medf1,design_a)
  m1 <- as.data.frame(magg1$t)
  rownames(m1) <- rownames(magg1)
  colnames(m1) <- "t"
  alres1 <- ttenrich(m=m1,genesets=gs_symbols,cores=CORES)

  #al
  medf2 <- chragg(mval_b,myann,cores=CORES)
  magg2 <- agglimma(medf2,design_b)
  m2 <- as.data.frame(magg2$t)
  rownames(m2) <- rownames(magg2)
  colnames(m2) <- "t"
  alres2 <- ttenrich(m=m2,genesets=gs_symbols,cores=CORES)

  list("alres1"=alres1,"alres2"=alres2)
}

alres <- lapply( seq(100,1000,100) ,splitanalyze_al)

alres

```

## Compare

```{r,compare}

lacompare <- function(lares){
  lares1 <- lares[[1]]
  lares2 <- lares[[2]]
  lasig1 <- subset(lares1,fdr<0.05)
  lasig2 <- subset(lares2,fdr<0.05)
  lasig1_up <- rownames(subset(lasig1,t_median>0))
  lasig1_dn <- rownames(subset(lasig1,t_median<0))
  lasig2_up <- rownames(subset(lasig2,t_median>0))
  lasig2_dn <- rownames(subset(lasig2,t_median<0))

  total <- length(unique(c(lasig1_up,lasig2_up,lasig1_dn,lasig2_dn)))
  common <- length(intersect(lasig1_up, lasig2_up)) + length(intersect(lasig1_dn, lasig2_dn))
  discordant <- length(intersect(lasig1_up, lasig2_dn)) + length(intersect(lasig1_dn, lasig2_up))
  uncommon = total - common - discordant

  lares <- as.data.frame(do.call(rbind,lares))
  lares$p_comm <-lares$common / lares$total
  lares$p_disc <-lares$discordant / lares$total
  lares$p_uncom <-lares$uncommon / lares$total
  lares
}

alcompare <-function(alres){
  alres1 <- alres[[1]]
  alres2 <- alres[[2]]
  alsig1 <- subset(alres1,fdr<0.05)
  alsig2 <- subset(alres2,fdr<0.05)

  alsig1_up <- rownames(subset(alsig1,t_median>0))
  alsig1_dn <- rownames(subset(alsig1,t_median<0))
  alsig2_up <- rownames(subset(alsig2,t_median>0))
  alsig2_dn <- rownames(subset(alsig2,t_median<0))

  total <- length(unique(c(alsig1_up,alsig2_up,alsig1_dn,alsig2_dn)))
  common <- length(intersect(alsig1_up, alsig2_up)) + length(intersect(alsig1_dn, alsig2_dn))
  discordant <- length(intersect(alsig1_up, alsig2_dn)) + length(intersect(alsig1_dn, alsig2_up))
  uncommon = total - common - discordant
  result <- c("total"=total,"common"=common,"discordant"=discordant,"uncommon"=uncommon)

  alres <- as.data.frame(do.call(rbind,alres))
  alres$p_comm <-alres$common / alres$total
  alres$p_disc <-alres$discordant / alres$total
  alres$p_uncom <-alres$uncommon / alres$total
  alres
}

```

## Session information

```{r,save}

save.image("GSE158422_gmea.Rdata")

sessionInfo()

```