### 2023-10-16 figs/GSE158422_simulate.Rmd
bugfix

---
title: "Testing gene methylation enrichment analysis approaches using simulated data"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

In this report we establish a new method for generating simulated data with known ground truth.
This will be used to test different gene methylation enrichment approaches systematically.

The general steps are:

1. Import GSE158422 data corresponding to control (non-tumour tissue).

2. From the 37 samples, create two groups of 18 samples.
One of these will be considered "control" and the other "case".

3. Create random gene sets that have similar sized to Reactome pathways.

3. Some gene sets will be selected to be differentially methylated.
Half of these will be hypermethylated and the others will be hypomethylated.

4. The changes will be incorporated into the "case" samples.

5. The enrichment analysis will be conducted.

6. The accuracy will be calculated.

```{r,packages}

suppressPackageStartupMessages({
  library("stringi")
  library("limma")
  library("missMethyl")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library('org.Hs.eg.db')
  library("tictoc")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("eulerr")
})

# optimised for 128 GB sever with 32 threads
CORES=6

```

## Load data

* annotations

* probe sets

* gene sets

* design matrix

* mval matrix

```{r,annotation}

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","Regulatory_Feature_Group","Islands_Name","Relation_to_Island")])

gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
sets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))

summary(unlist(lapply(sets,length)))

#genesets <- gmt_import("https://ziemann-lab.net/public/gmea_prototype/ReactomePathways.gmt")

if (!file.exists("GSE158422_design.rds")) {
  download.file("https://ziemann-lab.net/public/gmea_prototype/GSE158422_design.rds", "GSE158422_design.rds")
}
design <- readRDS("GSE158422_design.rds")

if (!file.exists("GSE158422_design.rds")) {
 download.file("https://ziemann-lab.net/public/gmea_prototype/GSE158422_mx.rds","GSE158422_mx.rds")
}
mval <- readRDS("GSE158422_mx.rds")

boxplot(list("normal"=matrix(colMeans(mval),ncol=2)[,2],"tumor"=matrix(colMeans(mval),ncol=2)[,1]),
  main="mean probe methylation mval")

```

# Gene set database

We could use Reactome pathways, however these have a lot of overlapping sets, which could cause inflated
false positives.
A better solution could be to select random gene sets with sizes similar to Reactomes.
Reactomes were downloaded on the 14th Sept 2023 from MsigDB.

```{r,gsdl}

gs_entrez <- gmt_import("c2.cp.reactome.v2023.1.Hs.entrez.gmt")

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

Get a vector of genes, and the length of gene sets from Reactome.

Then generate random gene sets.

```{r,gset_prep}

gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))

lengths <- unname(unlist(lapply(gs_symbols,length)))

randomGeneSets <- function(gene_catalog, lengths, seed){
  num_gsets <- length(lengths)
  set.seed(seed) ; seeds <- sample(1:1e6, num_gsets)
  gsets <- lapply(1:num_gsets,function(i) {
    set.seed(seeds[i]) ; gs <- sample(gene_catalog,lengths[i])
    return(gs)
  } )
  names(gsets)<-stri_rand_strings(length(gsets), 15, pattern = "[A-Za-z]")
  return(gsets)
}

gsets <- randomGeneSets(gene_catalog,lengths,seed=100)

```

Select nonoverlapping gene sets.

Not in use now. Might delete later.

```{r,gset_selection}

gset_selection <- function(genesetdatabase,seed,gene_catalog) {
  set.seed(seed)
  gsets <- genesetdatabase[sample(1:length(genesetdatabase))]
  genelist=NULL
  genesets=NULL
  for ( i in 1:length(gsets) ) {
    gs <- gsets[i]
    inx <- length(intersect(unlist(gs),genelist))
    num <- length(intersect(unlist(gs),gene_catalog))
    if ( inx == 0 & num >= 10 ) {
      genesets <- c(genesets,gs)
      genelist <- c(unname(unlist(gs)),genelist)
    }
  }
  return(genesets)
}

gset_mod <- gset_selection(genesetdatabase=gsets,seed=100,gene_catalog=gene_catalog)

```

# Incorporate changes

TODO: to incorporate changes to case samples.

Need to figure out what magnitude to change.
Will refer to the cancer/normal comparison.

Select genes and probes to alter.

```{r,incorp}

seed=100
frac_genes=0.5
frac_probes=0.5
delta=1
nsamples=10
normal_mval <- mval[,(1:(ncol(mval)/2)*2)]

incorp_dm <- function(genesets,myann,mval,seed,
  frac_genes,frac_probes,groupsize,delta=1) {

  # divide gene sets between hyper and hypomethylated
  nset <- floor(length(genesets)/2)
  set.seed(seed) ; gtup <-sample(genesets,nset)
  set.seed(seed) ; gtdn <- sample(setdiff(genesets,gtup),nset)
  gup <- unname(unlist(gtup))
  gdn <- unname(unlist(gtdn))
  # make probe-gene vector
  probe2gene <- strsplit(myann$UCSC_RefGene_Name,";")
  names(probe2gene) <- rownames(myann)
  probe2gene <- unlist(probe2gene)
  # select probes hypermethylated
  set.seed(seed) ; gup2 <- sample(gup,floor(length(gup)*frac_genes))
  pup <- names(probe2gene[which(probe2gene %in% gup2)])
  set.seed(seed) ; pup2 <- sample(pup,floor(length(pup)*frac_probes))
  # select probes hypomethylated
  set.seed(seed) ; gdn2 <- sample(gdn,floor(length(gdn)*frac_genes))
  pdn <- names(probe2gene[which(probe2gene %in% gdn2)])
  set.seed(seed) ; pdn2 <- sample(pdn,floor(length(pdn)*frac_probes))
  # divide samples between ctrl and case
  ncols <- ncol(mval)
  maxgroupsize=floor(ncols/2)
  if ( groupsize > maxgroupsize ) { stop("groupsize cannot be larger than half the ncols of mval") }
  set.seed(seed) ; ctrl <- sample(1:ncols,groupsize)
  set.seed(seed) ; case <- sample(setdiff(1:ncols,ctrl),groupsize)
  mval_ctrl <- mval[,ctrl]
  mval_case <- mval[,case]
  # incorporate altered signals - change by +1 or -1
  mval_case[rownames(mval_case) %in% pup2,] <-  mval_case[rownames(mval_case) %in% pup2,] + delta
  mval_case[rownames(mval_case) %in% pdn2,] <-  mval_case[rownames(mval_case) %in% pdn2,] - delta
  mval2 <- cbind(mval_ctrl,mval_case)
  result <- list("mval"=mval2,"probes up"=pup2,"probes down"=pdn2,
    "genes up"=gup2,"genes down"=gdn2,
    "genesets up"=gtup,"genesets down"=gtdn)
  return(result)
}

```

## GSAMETH function

```{r,gsameth_func}

# limma
runlimma <- function(mval,design,myann) {
  fit.reduced <- lmFit(mval,design)
  fit.reduced <- eBayes(fit.reduced)
  dm <- topTable(fit.reduced,coef=ncol(design), number = Inf)
  dm <- merge(myann,dm,by=0)
  dm <- dm[order(dm$P.Value),]
  rownames(dm) <- dm$Row.names
  dm$Row.names=NULL
  return(dm)
}

```

This is how to use the function

This could be complicated as it requires translation of symbols to entrez IDs, but
could be simplified if the entrez translation is done at the gsameth step.

```{r,rungsameth}

simgsa <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  dm3 <- runlimma(mval=mval2,design=d,myann=myann)
  pup3 <- rownames(subset(dm3,adj.P.Val<0.05 & logFC>0))
  pdn3 <- rownames(subset(dm3,adj.P.Val<0.05 & logFC<0))
  if ( length(pup3) < 250 ) { pup3 <- head(rownames(subset(dm3, logFC > 0)), 250) }
  if ( length(pdn3) < 250 ) { pdn3 <- head(rownames(subset(dm3, logFC < 0)), 250) }
  # convert gene sets to entrez
  suppressWarnings(suppressMessages({ gene2entrez <- mapIds(org.Hs.eg.db, gene_catalog, 'ENTREZID', 'SYMBOL') }))
  gsets_entrez <- lapply(gsets,function(gs) {
    gs2 <- unique(gene2entrez[names(gene2entrez) %in% gs])
    gs2 <- gs2[!is.na(gs2)]
    return(gs2)
  })
  suppressWarnings(suppressMessages({
    gsaup3 <- gsameth(sig.cpg=pup3, all.cpg=rownames(dm3), collection=gsets_entrez)
    gsadn3 <- gsameth(sig.cpg=pdn3, all.cpg=rownames(dm3), collection=gsets_entrez)
  }))
  gsig_up3 <- rownames(subset(gsaup3,FDR<0.05))
  gsig_dn3 <- rownames(subset(gsadn3,FDR<0.05))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(gsadn3)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

```

## LA function

This process runs limma first and then aggregates the results before doing an enrichment test.

```{r,lafunc}

# aggregate
agg <- function(dm,cores=1) {
  gn <- unique(unlist(strsplit( dm$UCSC_RefGene_Name ,";")))
  gnl <- strsplit( dm$UCSC_RefGene_Name ,";")
  gnl <- mclapply(gnl,unique,mc.cores=cores)
  dm$UCSC_RefGene_Name <- gnl
  l <- mclapply(1:nrow(dm), function(i) {
    a <- dm[i,]
    len <- length(a[[1]][[1]])
    tvals <- as.numeric(rep(a["t"],len))
    genes <- a[[1]][[1]]
    data.frame(genes,tvals)
  },mc.cores=cores)
  df <- do.call(rbind,l)
  keep <- names(which(table(df$genes)>1))
  df <- df[df$genes %in% keep,]
  gn <- unique(df$genes)
  gme_res <- lapply( 1:length(gn), function(i) {
    g <- gn[i]
    tstats <- df[which(df$genes==g),"tvals"]
    myn <- length(tstats)
    mymean <- mean(tstats)
    mymedian <- median(tstats)
    if ( length(tstats) > 2 ) {
      ttest <- t.test(tstats)
      pval <- ttest$p.value
    } else {
      pval = 1
    }
    res <- c("gene"=g,"nprobes"=myn,"mean"=mymean,
      "median"=mymedian, pval=pval)
  } )
  gme_res_df <- do.call(rbind, gme_res)
  rownames(gme_res_df) <- gme_res_df[,1]
  gme_res_df <- gme_res_df[,-1]
  tmp <- apply(gme_res_df,2,as.numeric)
  rownames(tmp) <- rownames(gme_res_df)
  gme_res_df <- as.data.frame(tmp)
  gme_res_df$sig <- -log10(gme_res_df[,4])
  gme_res_df <- gme_res_df[order(-gme_res_df$sig),]
  gme_res_df$fdr <- p.adjust(gme_res_df$pval)
  out <- list("df"=df,"gme_res_df"=gme_res_df)
  return(out)
}

# enrich
ttenrich <- function(m,genesets,cores=1) {
  res <- mclapply( 1:length(genesets), function(i) {
    gs <- genesets[i]
    name <- names(gs)
    n_members <- length(which(rownames(m) %in% gs[[1]]))
    if ( n_members > 4 ) {
      tstats <- m[which(rownames(m) %in% gs[[1]]),]
      myn <- length(tstats)
      mymean <- mean(tstats)
      mymedian <- median(tstats)
      wt <- t.test(tstats)
      res <- c(name,myn,mymean,mymedian,wt$p.value)
    }
  } , mc.cores = cores)
  res_df <- do.call(rbind, res)
  rownames(res_df) <- res_df[,1]
  res_df <- res_df[,-1]
  colnames(res_df) <- c("n_genes","t_mean","t_median","pval")
  tmp <- apply(res_df,2,as.numeric)
  rownames(tmp) <- rownames(res_df)
  res_df <- tmp
  res_df <- as.data.frame(res_df)
  res_df <- res_df[order(res_df$pval),]
  res_df$logp <- -log10(res_df$pval )
  res_df$fdr <- p.adjust(res_df$pval,method="fdr")
  res_df[order(abs(res_df$pval)),]
  return(res_df)
}

simla <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  dm3 <- runlimma(mval=mval2,design=d,myann=myann)
  dmagg1 <- agg(dm3,cores=4)
  m1 <- dmagg1$gme_res_df[,"mean",drop=FALSE]
  lares1 <- ttenrich(m=m1,genesets=gsets,cores=10)
  gsig_up3 <- rownames(subset(lares1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(lares1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(lares1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

```

## AL: Aggregate limma functions

Functions for aggregate-limma-enrich approach.

```{r,alfunc}

# chromosome by chromosome will be much faster
magg <- function(mval,myann,cores=1){
  gn <- unique(unlist(strsplit( myann$UCSC_RefGene_Name ,";")))
  gnl <- strsplit( myann$UCSC_RefGene_Name ,";")
  gnl <- mclapply(gnl,unique,mc.cores=cores)
  myann$gnl <- gnl
  keep <- rownames(subset(myann,UCSC_RefGene_Name!=""))
  mx <- mval[rownames(mval) %in% keep,]
  mymed <- function(g) {
    probes <- rownames(myann[grep(g,myann$gnl),])
    rows <- which(rownames(mx) %in% probes)
    if ( length(rows) > 1 ) {
      b <- mx[rows,]
      med <- apply(b,2,mean)
      med <- matrix(med,nrow=1)
      colnames(med) <- colnames(b)
      rownames(med) <- g
      return(med)
    }
  }
  med <- mclapply(gn,mymed,mc.cores=cores)
  med <- med[lapply(med,length)>0]
  medf <- do.call(rbind,med)
  return(medf)
}

chragg <- function(mval,myann,cores=1){
  annodf <- as.data.frame(anno)
  keep <- rownames(subset(myann,UCSC_RefGene_Name!=""))
  mx <- mval[rownames(mval) %in% keep,]
  chrs <- unique(anno$chr)
  myorder <- unlist(lapply(chrs,function(mychr) { nrow( annodf[annodf$chr==mychr,] ) } ))
  chrs <- chrs[order(-myorder)]
  leadercores <- floor(sqrt(cores))
  workercores <- ceiling(sqrt(cores))
  chrmedf <- mclapply(chrs,function(chr) {
    chrfrag <- annodf[annodf$chr==chr,]
    chrprobes <-rownames(chrfrag)
    chrmx <- mx[rownames(mx) %in% chrprobes,]
    chranno <- myann[rownames(myann) %in% chrprobes,]
    chrmedf <- magg(mval=chrmx,myann=chranno,cores=workercores)
    return(chrmedf)
  },mc.cores=leadercores)
  medf <- do.call(rbind, chrmedf)
  return(medf)
}

agglimma <- function(medf,design) {
  fit.reduced <- lmFit(medf,design)
  fit.reduced <- eBayes(fit.reduced)
  dmagg <- topTable(fit.reduced,coef=ncol(design), number = Inf)
  nondup <- !duplicated(dmagg$ID)
  dmagg <- dmagg[nondup,]
  rownames(dmagg) <- dmagg$ID
  dmagg$ID = NULL
  return(dmagg)
}

simal <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  # al pipeline
  medf1 <- chragg(mval=mval2,myann=myann,cores=4)
  magg1 <- agglimma(medf1,d)
  m1 <- as.data.frame(magg1$t)
  rownames(m1) <- rownames(magg1)
  colnames(m1) <- "t"
  alres1 <- ttenrich(m=m1,genesets=gsets,cores=10)
  # summarise results
  gsig_up3 <- rownames(subset(alres1, fdr < 0.05 & t_mean > 0))
  gsig_dn3 <- rownames(subset(alres1, fdr < 0.05 & t_mean < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(alres1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

```

## AA Aggregate-aggregate-limma functions

Use mean value works well here.

```{r,aafunc}

gsagg <- function(x,genesets,cores=1) {
  meds <- mclapply(1:length(genesets), function(i) {
    gs = genesets[[i]]
    xx <- x[rownames(x) %in% gs,]
    med <- apply(xx,2,mean)
  },mc.cores=cores)
  mymed <- do.call(rbind,meds)
  rownames(mymed) <- names(genesets)
  as.data.frame(mymed)
}

aalimma <- function(agag,design) {
  fit.reduced <- lmFit(agag,design)
  fit.reduced <- eBayes(fit.reduced)
  dmagg <- topTable(fit.reduced,coef=ncol(design), number = Inf)
  return(dmagg)
}

aal <- function(mval,myann,genesets,design,cores=1) {
  medf <- chragg(mval,myann,cores=cores)
  agag <- gsagg(x=medf,genesets=genesets,cores=cores)
  aalres <- aalimma(agag=agag,design=design)
  return(aalres)
}

simaa <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  # aa pipeline
  aares1 <- aal(mval=mval2, myann=myann, genesets=gsets, design=d, cores=4)
  # summarise results
  gsig_up3 <- rownames(subset(aares1, adj.P.Val < 0.05 & logFC > 0))
  gsig_dn3 <- rownames(subset(aares1, adj.P.Val < 0.05 & logFC < 0))
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(aares1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

```

## Agg-limma-mitch function

This approach uses the aggregated mvals, limma and instead of a 1-sample t-test it uses mitch
which is a competitive test and could give more interpretable results.

```{r,alm}

runmitch <- function(m,genesets,cores=1) {
  suppressMessages({ mres <- mitch_calc(m,genesets,minsetsize=5,cores=cores) })
  mres <- mres$enrichment_result
  rownames(mres) <- mres$set
  mres$set=NULL
  return(mres)
}

simalm <- function(genesetdatabase, myann, mval, seed, frac_genes, frac_probes, groupsize, delta=1, num_dm_sets=50) {
  # generate gene sets
  gene_catalog <- unique(unlist(strsplit(myann$UCSC_RefGene_Name,";")))
  lengths <- unname(unlist(lapply(genesetdatabase,length)))
  gsets <- randomGeneSets(gene_catalog,lengths,seed=seed)
  # select gene sets to alter
  set.seed(seed) ; gset_mod <- sample(gsets,num_dm_sets)
  # incorporate select changes
  sim <- incorp_dm(genesets=gset_mod, myann=myann, mval=mval, seed=seed,
    frac_genes=0.5,frac_probes=0.5,groupsize=groupsize,delta=delta)
  # set up limma
  mval2 <- sim$mval
  ncols <- ncol(mval2)
  groupsize <- ncols/2
  ss <- data.frame(colnames(mval2))
  colnames(ss) <- "sample"
  ss$case <- c(rep(0,groupsize),rep(1,groupsize))
  d <- model.matrix(~ ss$case )
  # alm pipeline
  medf1 <- chragg(mval2,myann,cores=4)
  magg1 <- agglimma(medf1,d)
  m1 <- as.data.frame(magg1$t)
  rownames(m1) <- rownames(magg1)
  colnames(m1) <- "t"
  almres1 <- runmitch(m=m1,genesets=gsets,cores=4)
  # summarise results
  gsig_up3 <- rownames( subset( almres1, p.adjustANOVA < 0.05 & s.dist > 0 ) )
  gsig_dn3 <- rownames( subset( almres1, p.adjustANOVA < 0.05 & s.dist < 0 ) )
  gtup <- names(sim[[6]])
  gtdn <- names(sim[[7]])
  UPTP=length(intersect(gsig_up3 ,gtup))
  UPFP=length(setdiff(gsig_up3 ,gtup))
  UPFN=length(setdiff(gtup,gsig_up3))
  DNTP=length(intersect(gsig_dn3 ,gtdn))
  DNFP=length(setdiff(gsig_dn3 ,gtdn))
  DNFN=length(setdiff(gtdn,gsig_dn3))
  TP=UPTP+DNTP
  FP=UPFP+DNFP
  FN=UPFN+DNFN
  TN=nrow(almres1)-DNTP-DNFP-DNFN-UPTP-UPFP-UPFN
  PREC=TP/(TP+FP)
  REC=TP/(TP+FN)
  F1=TP/(TP+(0.5*(FP+FN)))
  result <- c("TP"=TP,"FP"=FP,"FN"=FN,"TN"=TN,"PREC"=PREC,"REC"=REC)
  return(result)
}

```

# Run analyses

Set assumptions.

```{r,set}

groupsize=3
delta=0.6
num_dm_sets=50

```

## GSA meth sim

Cannot be run in multicore due to fragility of AnnotationDbi SQLite objects.

```{r,rungsa}

gres <- lapply(1:100,function(i) {
  simgsa(genesetdatabase=gs_symbols, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
    groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
})

gres <- do.call(rbind,gres)
gres

```

## LA sim

```{r,runla}

lares <- mclapply(1:100,function(i) {
  simla(genesetdatabase=gs_symbols, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
    groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
},mc.cores=6)

lares <- do.call(rbind,lares)
lares

```

## AL sim

```{r,runal}

alres <- mclapply(1:100,function(i) {
  simal(genesetdatabase=gs_symbols, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
    groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
},mc.cores=6)

alres <- do.call(rbind,alres)
alres

```

## AA sim

```{r,runaa}

aares <- mclapply(1:100,function(i) {
  simaa(genesetdatabase=gs_symbols, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
    groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
},mc.cores=6)

aares <- do.call(rbind,aares)
aares

```

## ALM competitive test

```{r,runalm}

almres <- mclapply(1:100,function(i) {
  simalm(genesetdatabase=gs_symbols, myann=myann, mval=normal_mval, seed=i*100, frac_genes=0.5, frac_probes=0.5,
    groupsize=groupsize, delta=delta, num_dm_sets=num_dm_sets)
},mc.cores=6)

almres <- do.call(rbind,almres)
almres

```

## Session information

```{r,save}

save.image("GSE158422_simulate.Rdata")

sessionInfo()

```