### 2023-11-02 figs/GSE158422_simulate_summary.Rmd
summary of simulation results good

---
title: "Testing gene methylation enrichment analysis approaches using simulated data - summarise data"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

In this report we establish a new method for generating simulated data with known ground truth.
This will be used to test different gene methylation enrichment approaches systematically.

The general steps are:

1. Import GSE158422 data corresponding to control (non-tumour tissue).

2. From the 37 samples, create two groups of 18 samples.
One of these will be considered "control" and the other "case".

3. Create random gene sets that have similar sized to Reactome pathways.

3. Some gene sets will be selected to be differentially methylated.
Half of these will be hypermethylated and the others will be hypomethylated.

4. The changes will be incorporated into the "case" samples.

5. The enrichment analysis will be conducted.

6. The accuracy will be calculated.

```{r,packages}

suppressPackageStartupMessages({
  library("stringi")
  library("limma")
  library("missMethyl")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library('org.Hs.eg.db')
  library("psych")
  library("mitch")
  library("kableExtra")
})

```

## Functions

```{r,functions}

F1 <- function(x,y) {
  ( 2 * x * y ) / ( x + y )
}

```

# Geneset size = 100

```{r,gsa}

load("GSE158422_simulate100.Rdata")

par(mfrow=c(2,3))

gres2[,"PREC"] <- gres2[,"TP"] / ( gres2[,"TP"] + gres2[,"FP"] )
gres3p <- do.call(rbind,lapply(groupsizes, function (g) { gres2[params$groupsizes==g,"PREC"] }))
colnames(gres3p) <- deltas
rownames(gres3p) <- groupsizes
gres3p %>% kbl(caption="GSAmeth precision") %>% kable_paper("hover", full_width = F)

gres3r <- do.call(rbind,lapply(groupsizes, function (g) { gres2[params$groupsizes==g,"REC"] }))
colnames(gres3r) <- deltas
rownames(gres3r) <- groupsizes
gres3r %>% kbl(caption="GSAmeth recall") %>% kable_paper("hover", full_width = F)

gres3f <- F1(gres3p,gres3r)
gres3f %>% kbl(caption="GSAmeth F1") %>% kable_paper("hover", full_width = F)

plot(rownames(gres3p),gres3p[,"0.5"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(rownames(gres3p),gres3p[,"0.4"],type="b",col="red")
lines(rownames(gres3p),gres3p[,"0.3"],type="b",col="blue")
lines(rownames(gres3p),gres3p[,"0.2"],type="b",col="darkgreen")
grid()

legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen"), cex=1)

plot(rownames(gres3r),gres3r[,"0.5"],type="b",las=1,ylab="recall",xlab="group size",
  main="GSA",ylim=c(0,1))
lines(rownames(gres3r),gres3r[,"0.4"],type="b",col="red")
lines(rownames(gres3r),gres3r[,"0.3"],type="b",col="blue")
lines(rownames(gres3r),gres3r[,"0.2"],type="b",col="darkgreen")
grid()

plot(rownames(gres3r),gres3f[,"0.5"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(rownames(gres3r),gres3f[,"0.4"],type="b",col="red")
lines(rownames(gres3r),gres3f[,"0.3"],type="b",col="blue")
lines(rownames(gres3r),gres3f[,"0.2"],type="b",col="darkgreen")
grid()

lacmres2[,"PREC"] <- lacmres2[,"TP"] / ( lacmres2[,"TP"] + lacmres2[,"FP"] )
lacmres3p <- do.call(rbind,lapply(groupsizes, function (g) { lacmres2[params$groupsizes==g,"PREC"] }))
colnames(lacmres3p) <- deltas
rownames(lacmres3p) <- groupsizes
lacmres3p %>% kbl(caption="LAM precision") %>% kable_paper("hover", full_width = F)

lacmres3r <- do.call(rbind,lapply(groupsizes, function (g) { lacmres2[params$groupsizes==g,"REC"] }))
colnames(lacmres3r) <- deltas
rownames(lacmres3r) <- groupsizes
lacmres3r %>% kbl(caption="LAM recall") %>% kable_paper("hover", full_width = F)

lacmres3f <- F1(lacmres3p,lacmres3r)
lacmres3f %>% kbl(caption="LAM F1") %>% kable_paper("hover", full_width = F)

plot(rownames(lacmres3p),lacmres3p[,"0.5"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(rownames(lacmres3p),lacmres3p[,"0.4"],type="b",col="red")
lines(rownames(lacmres3p),lacmres3p[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3p),lacmres3p[,"0.2"],type="b",col="darkgreen")
grid()

legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen"), cex=1)

plot(rownames(lacmres3r),lacmres3r[,"0.5"],type="b",las=1,ylab="recall",xlab="group size",
  main="LAM",ylim=c(0,1))
lines(rownames(lacmres3r),lacmres3r[,"0.4"],type="b",col="red")
lines(rownames(lacmres3r),lacmres3r[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3r),lacmres3r[,"0.2"],type="b",,col="darkgreen")
grid()

plot(rownames(lacmres3r),lacmres3f[,"0.5"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(rownames(lacmres3r),lacmres3f[,"0.4"],type="b",col="red")
lines(rownames(lacmres3r),lacmres3f[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3r),lacmres3f[,"0.2"],type="b",col="darkgreen")
grid()

dev.off()

```

## Session information

```{r,save}

sessionInfo()

```

## Notes

LA

* simla: parametric self-contained

* simlac: parametric competitive

* simnla: nonparametric self-contained

* simnlac: nonparametric competitive

AL

* simal: parametric self-contained

* simalc: parametric competitive

* simnal: nonparametric self-contained

* simnalc: nonparametric competitive