### 2023-11-04 figs/GSE158422_simulate_summary.Rmd
fig1

---
title: "Testing gene methylation enrichment analysis approaches using simulated data - summarise data"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Here is an explanation of the different objects:

0. GSA: gres

1. LAT: lacres

2. LTT: lactopres

3. LAW: nlacres

4. LAM: lacmres

5. ALT: alcres

6. ALW: nalcres

7. ALM: almres

```{r,packages}

suppressPackageStartupMessages({
  library("stringi")
  library("limma")
  library("missMethyl")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library('org.Hs.eg.db')
  library("psych")
  library("mitch")
  library("kableExtra")
})

```

## Functions

```{r,functions}

F1 <- function(x,y) {
  ( 2 * x * y ) / ( x + y )
}

summarise <- function(x) {
  PREC <- sum(x[,"TP"]) / ( sum(x[,"TP"]) + sum(x[,"FP"]) )
  REC <- sum(x[,"TP"]) / ( sum(x[,"TP"]) + sum(x[,"FN"]) )
  F <- F1(PREC,REC)
  return(c("PRECISION"=PREC,"RECALL"=REC,"F1"=F))
}

```

## Overall summary

Make a set of three barplots which summarise method performance over the range of group size and
delta values, for gene set sizes of 20, 50 and 200.

```{r,summary,fig.width=5,fig.height=4}

mar = c(5.1, 4.1, 5.1, 2.1)

load("GSE158422_simulate020.Rdata")

l1 <- list("GSA"=gres2,"LAT"=lacres2,"LTT"=lactopres2,"LAW"=nlacres2,
  "LAM"=lacmres2,"ALT"=alcres2,"ALW"=nalcres2,"ALM"=almres2)

x1 <- do.call(rbind,lapply(l1,summarise))

x1 %>% kbl(caption="sim 20 res") %>% kable_paper("hover", full_width = F)

cols <- c("white","gray","black")

barplot(t(x1),beside=TRUE,ylim=c(0,1),ylab="index", col=cols, cex.names=0.7, cex.axis=0.7, main="20 genes per set")
grid()

legend("topleft", inset=.1, c("Precision","Recall","F1"), fill=cols, cex=0.7)

load("GSE158422_simulate050.Rdata")

l2 <- list("GSA"=gres2,"LAT"=lacres2,"LTT"=lactopres2,"LAW"=nlacres2,
  "LAM"=lacmres2,"ALT"=alcres2,"ALW"=nalcres2,"ALM"=almres2)

x2 <- do.call(rbind,lapply(l2,summarise))

x2 %>% kbl(caption="sim 50 res") %>% kable_paper("hover", full_width = F)

barplot(t(x2),beside=TRUE,ylim=c(0,1),ylab="index", col=cols, cex.names=0.7, cex.axis=0.7, main="50 genes per set")
grid()

load("GSE158422_simulate100.Rdata")

l3 <- list("GSA"=gres2,"LAT"=lacres2,"LTT"=lactopres2,"LAW"=nlacres2,
  "LAM"=lacmres2,"ALT"=alcres2,"ALW"=nalcres2,"ALM"=almres2)

x3 <- do.call(rbind,lapply(l3,summarise))

x %>% kbl(caption="sim 100 res") %>% kable_paper("hover", full_width = F)

barplot(t(x),beside=TRUE,ylim=c(0,1),ylab="index", col=cols, cex.names=0.7, cex.axis=0.7, main="100 genes per set")
grid()

```

Make a PNG copy of the data.

```{r,summary_png,fig.width=5,fig.height=4}

png(filename="fig1.png",height=70,width=210,units="mm",res=200)

par(mar = c(2.5, 4.1, 2.1, 0.5))

par(mfrow=c(1,3))

cols <- c("white","gray","black")

barplot(t(x1),beside=TRUE,ylim=c(0,1),ylab="index", col=cols, cex.names=0.75, cex.axis=0.75, main="20 genes per set")
grid()

legend("topleft", inset=.1, c("Precision","Recall","F1"), fill=cols, cex=0.75)

barplot(t(x2),beside=TRUE,ylim=c(0,1),ylab="index", col=cols, cex.names=0.75, cex.axis=0.75, main="50 genes per set")
grid()

barplot(t(x3),beside=TRUE,ylim=c(0,1),ylab="index", col=cols, cex.names=0.75, cex.axis=0.75, main="100 genes per set")
grid()

dev.off()

par(mfrow=c(1,1))

```

## In depth analysis of GSA vs LAM

As GSA appears to be the best-performing existing method for enrichment analysis, it is a good idea
to compare it to LAM which is the best one out of the new methods evaluated here.
The following charts show the performance of GSA and LAM including precision and recall over the
range of group sizes and delta values when the sets contain 20, 50 and 100 genes each.

```{r,020_gsa_vs_lam}

# 20
load("GSE158422_simulate020.Rdata")

par(mfrow=c(2,3))

gres2[,"PREC"] <- gres2[,"TP"] / ( gres2[,"TP"] + gres2[,"FP"] )
gres3p <- do.call(rbind,lapply(groupsizes, function (g) { gres2[params$groupsizes==g,"PREC"] }))
colnames(gres3p) <- deltas
rownames(gres3p) <- groupsizes
gres3p %>% kbl(caption="GSAmeth precision") %>% kable_paper("hover", full_width = F)

gres3r <- do.call(rbind,lapply(groupsizes, function (g) { gres2[params$groupsizes==g,"REC"] }))
colnames(gres3r) <- deltas
rownames(gres3r) <- groupsizes
gres3r %>% kbl(caption="GSAmeth recall") %>% kable_paper("hover", full_width = F)

gres3f <- F1(gres3p,gres3r)
gres3f %>% kbl(caption="GSAmeth F1") %>% kable_paper("hover", full_width = F)

plot(rownames(gres3p),gres3p[,"0.5"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(rownames(gres3p),gres3p[,"0.4"],type="b",col="red")
lines(rownames(gres3p),gres3p[,"0.3"],type="b",col="blue")
lines(rownames(gres3p),gres3p[,"0.2"],type="b",col="darkgreen")
grid()

legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen"), cex=1)

plot(rownames(gres3r),gres3r[,"0.5"],type="b",las=1,ylab="recall",xlab="group size",
  main="GSA",ylim=c(0,1))
lines(rownames(gres3r),gres3r[,"0.4"],type="b",col="red")
lines(rownames(gres3r),gres3r[,"0.3"],type="b",col="blue")
lines(rownames(gres3r),gres3r[,"0.2"],type="b",col="darkgreen")
grid()

plot(rownames(gres3r),gres3f[,"0.5"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(rownames(gres3r),gres3f[,"0.4"],type="b",col="red")
lines(rownames(gres3r),gres3f[,"0.3"],type="b",col="blue")
lines(rownames(gres3r),gres3f[,"0.2"],type="b",col="darkgreen")
grid()

lacmres2[,"PREC"] <- lacmres2[,"TP"] / ( lacmres2[,"TP"] + lacmres2[,"FP"] )
lacmres3p <- do.call(rbind,lapply(groupsizes, function (g) { lacmres2[params$groupsizes==g,"PREC"] }))
colnames(lacmres3p) <- deltas
rownames(lacmres3p) <- groupsizes
lacmres3p %>% kbl(caption="LAM precision") %>% kable_paper("hover", full_width = F)

lacmres3r <- do.call(rbind,lapply(groupsizes, function (g) { lacmres2[params$groupsizes==g,"REC"] }))
colnames(lacmres3r) <- deltas
rownames(lacmres3r) <- groupsizes
lacmres3r %>% kbl(caption="LAM recall") %>% kable_paper("hover", full_width = F)

lacmres3f <- F1(lacmres3p,lacmres3r)
lacmres3f %>% kbl(caption="LAM F1") %>% kable_paper("hover", full_width = F)

plot(rownames(lacmres3p),lacmres3p[,"0.5"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(rownames(lacmres3p),lacmres3p[,"0.4"],type="b",col="red")
lines(rownames(lacmres3p),lacmres3p[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3p),lacmres3p[,"0.2"],type="b",col="darkgreen")
grid()

legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen"), cex=1)

plot(rownames(lacmres3r),lacmres3r[,"0.5"],type="b",las=1,ylab="recall",xlab="group size",
  main="LAM",ylim=c(0,1))
lines(rownames(lacmres3r),lacmres3r[,"0.4"],type="b",col="red")
lines(rownames(lacmres3r),lacmres3r[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3r),lacmres3r[,"0.2"],type="b",,col="darkgreen")
grid()

plot(rownames(lacmres3r),lacmres3f[,"0.5"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(rownames(lacmres3r),lacmres3f[,"0.4"],type="b",col="red")
lines(rownames(lacmres3r),lacmres3f[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3r),lacmres3f[,"0.2"],type="b",col="darkgreen")
grid()

# 50
load("GSE158422_simulate050.Rdata")

gres2[,"PREC"] <- gres2[,"TP"] / ( gres2[,"TP"] + gres2[,"FP"] )
gres3p <- do.call(rbind,lapply(groupsizes, function (g) { gres2[params$groupsizes==g,"PREC"] }))
colnames(gres3p) <- deltas
rownames(gres3p) <- groupsizes
gres3p %>% kbl(caption="GSAmeth precision") %>% kable_paper("hover", full_width = F)

gres3r <- do.call(rbind,lapply(groupsizes, function (g) { gres2[params$groupsizes==g,"REC"] }))
colnames(gres3r) <- deltas
rownames(gres3r) <- groupsizes
gres3r %>% kbl(caption="GSAmeth recall") %>% kable_paper("hover", full_width = F)

gres3f <- F1(gres3p,gres3r)
gres3f %>% kbl(caption="GSAmeth F1") %>% kable_paper("hover", full_width = F)

plot(rownames(gres3p),gres3p[,"0.5"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(rownames(gres3p),gres3p[,"0.4"],type="b",col="red")
lines(rownames(gres3p),gres3p[,"0.3"],type="b",col="blue")
lines(rownames(gres3p),gres3p[,"0.2"],type="b",col="darkgreen")
grid()

legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen"), cex=1)

plot(rownames(gres3r),gres3r[,"0.5"],type="b",las=1,ylab="recall",xlab="group size",
  main="GSA",ylim=c(0,1))
lines(rownames(gres3r),gres3r[,"0.4"],type="b",col="red")
lines(rownames(gres3r),gres3r[,"0.3"],type="b",col="blue")
lines(rownames(gres3r),gres3r[,"0.2"],type="b",col="darkgreen")
grid()

plot(rownames(gres3r),gres3f[,"0.5"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(rownames(gres3r),gres3f[,"0.4"],type="b",col="red")
lines(rownames(gres3r),gres3f[,"0.3"],type="b",col="blue")
lines(rownames(gres3r),gres3f[,"0.2"],type="b",col="darkgreen")
grid()

lacmres2[,"PREC"] <- lacmres2[,"TP"] / ( lacmres2[,"TP"] + lacmres2[,"FP"] )
lacmres3p <- do.call(rbind,lapply(groupsizes, function (g) { lacmres2[params$groupsizes==g,"PREC"] }))
colnames(lacmres3p) <- deltas
rownames(lacmres3p) <- groupsizes
lacmres3p %>% kbl(caption="LAM precision") %>% kable_paper("hover", full_width = F)

lacmres3r <- do.call(rbind,lapply(groupsizes, function (g) { lacmres2[params$groupsizes==g,"REC"] }))
colnames(lacmres3r) <- deltas
rownames(lacmres3r) <- groupsizes
lacmres3r %>% kbl(caption="LAM recall") %>% kable_paper("hover", full_width = F)

lacmres3f <- F1(lacmres3p,lacmres3r)
lacmres3f %>% kbl(caption="LAM F1") %>% kable_paper("hover", full_width = F)

plot(rownames(lacmres3p),lacmres3p[,"0.5"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(rownames(lacmres3p),lacmres3p[,"0.4"],type="b",col="red")
lines(rownames(lacmres3p),lacmres3p[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3p),lacmres3p[,"0.2"],type="b",col="darkgreen")
grid()

legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen"), cex=1)

plot(rownames(lacmres3r),lacmres3r[,"0.5"],type="b",las=1,ylab="recall",xlab="group size",
  main="LAM",ylim=c(0,1))
lines(rownames(lacmres3r),lacmres3r[,"0.4"],type="b",col="red")
lines(rownames(lacmres3r),lacmres3r[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3r),lacmres3r[,"0.2"],type="b",,col="darkgreen")
grid()

plot(rownames(lacmres3r),lacmres3f[,"0.5"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(rownames(lacmres3r),lacmres3f[,"0.4"],type="b",col="red")
lines(rownames(lacmres3r),lacmres3f[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3r),lacmres3f[,"0.2"],type="b",col="darkgreen")
grid()

# 100
load("GSE158422_simulate100.Rdata")

gres2[,"PREC"] <- gres2[,"TP"] / ( gres2[,"TP"] + gres2[,"FP"] )
gres3p <- do.call(rbind,lapply(groupsizes, function (g) { gres2[params$groupsizes==g,"PREC"] }))
colnames(gres3p) <- deltas
rownames(gres3p) <- groupsizes
gres3p %>% kbl(caption="GSAmeth precision") %>% kable_paper("hover", full_width = F)

gres3r <- do.call(rbind,lapply(groupsizes, function (g) { gres2[params$groupsizes==g,"REC"] }))
colnames(gres3r) <- deltas
rownames(gres3r) <- groupsizes
gres3r %>% kbl(caption="GSAmeth recall") %>% kable_paper("hover", full_width = F)

gres3f <- F1(gres3p,gres3r)
gres3f %>% kbl(caption="GSAmeth F1") %>% kable_paper("hover", full_width = F)

plot(rownames(gres3p),gres3p[,"0.5"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(rownames(gres3p),gres3p[,"0.4"],type="b",col="red")
lines(rownames(gres3p),gres3p[,"0.3"],type="b",col="blue")
lines(rownames(gres3p),gres3p[,"0.2"],type="b",col="darkgreen")
grid()

legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen"), cex=1)

plot(rownames(gres3r),gres3r[,"0.5"],type="b",las=1,ylab="recall",xlab="group size",
  main="GSA",ylim=c(0,1))
lines(rownames(gres3r),gres3r[,"0.4"],type="b",col="red")
lines(rownames(gres3r),gres3r[,"0.3"],type="b",col="blue")
lines(rownames(gres3r),gres3r[,"0.2"],type="b",col="darkgreen")
grid()

plot(rownames(gres3r),gres3f[,"0.5"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(rownames(gres3r),gres3f[,"0.4"],type="b",col="red")
lines(rownames(gres3r),gres3f[,"0.3"],type="b",col="blue")
lines(rownames(gres3r),gres3f[,"0.2"],type="b",col="darkgreen")
grid()

lacmres2[,"PREC"] <- lacmres2[,"TP"] / ( lacmres2[,"TP"] + lacmres2[,"FP"] )
lacmres3p <- do.call(rbind,lapply(groupsizes, function (g) { lacmres2[params$groupsizes==g,"PREC"] }))
colnames(lacmres3p) <- deltas
rownames(lacmres3p) <- groupsizes
lacmres3p %>% kbl(caption="LAM precision") %>% kable_paper("hover", full_width = F)

lacmres3r <- do.call(rbind,lapply(groupsizes, function (g) { lacmres2[params$groupsizes==g,"REC"] }))
colnames(lacmres3r) <- deltas
rownames(lacmres3r) <- groupsizes
lacmres3r %>% kbl(caption="LAM recall") %>% kable_paper("hover", full_width = F)

lacmres3f <- F1(lacmres3p,lacmres3r)
lacmres3f %>% kbl(caption="LAM F1") %>% kable_paper("hover", full_width = F)

plot(rownames(lacmres3p),lacmres3p[,"0.5"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(rownames(lacmres3p),lacmres3p[,"0.4"],type="b",col="red")
lines(rownames(lacmres3p),lacmres3p[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3p),lacmres3p[,"0.2"],type="b",col="darkgreen")
grid()

legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen"), cex=1)

plot(rownames(lacmres3r),lacmres3r[,"0.5"],type="b",las=1,ylab="recall",xlab="group size",
  main="LAM",ylim=c(0,1))
lines(rownames(lacmres3r),lacmres3r[,"0.4"],type="b",col="red")
lines(rownames(lacmres3r),lacmres3r[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3r),lacmres3r[,"0.2"],type="b",,col="darkgreen")
grid()

plot(rownames(lacmres3r),lacmres3f[,"0.5"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(rownames(lacmres3r),lacmres3f[,"0.4"],type="b",col="red")
lines(rownames(lacmres3r),lacmres3f[,"0.3"],type="b",col="blue")
lines(rownames(lacmres3r),lacmres3f[,"0.2"],type="b",col="darkgreen")
grid()

par(mfrow=c(1,1))

```

## Session information

```{r,save}

sessionInfo()

```

## Notes

LA

* simla: parametric self-contained

* simlac: parametric competitive

* simnla: nonparametric self-contained

* simnlac: nonparametric competitive

AL

* simal: parametric self-contained

* simalc: parametric competitive

* simnal: nonparametric self-contained

* simnalc: nonparametric competitive