### 2023-11-17 figs/GSE158422_rna_int.Rmd
force use of EPIC annotation for GSAmeth

---
title: "GSE158422: DNA methylation in Lung squamous cell carcinoma - RNA expression and integration"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Let's take a look at how pathway methylation associates with gene expression.

Partition methylation analysis between promoter and body.

Compare to GSAmeth.

```{r,packages}

suppressPackageStartupMessages({
  library("DESeq2")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("HGNChelper")
  library("gplots")
})

CORES=12

```

## Load annotations and methylation data

```{r,annotation}

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","Regulatory_Feature_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
sets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))
summary(unlist(lapply(sets,length)))

design <- readRDS("GSE158422_design.rds")
mval <- readRDS("GSE158422_mx.rds")
dm <- readRDS("GSE158422_dm.rds")

```

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,gsdl}

gs_entrez <- gmt_import("c2.cp.reactome.v2023.1.Hs.entrez.gmt")

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

## Load RNA expression

From GEO https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE158420
GSE158420_counts.txt.gz

## DESEQ2

```{r,deseq1}

x <- read.table("GSE158420_counts.txt.gz",row.names=NULL)
dim(x)
x <- aggregate(. ~ row.names,x,sum)
dim(x)

#RUFKM
head(x)

```

There are some gene name errors, so I will use HGNChelper to repair them.

```{r,gene_name_fix}

human=x$row.names
fix <- checkGeneSymbols(human)
x$gene <- fix$Suggested.Symbol
x$row.names=NULL
x <- aggregate(. ~ gene,x,sum)
dim(x)
row.names(x)=x$gene
x$gene=NULL
head(x)

```

Now do DESeq2

```{r,deseq2}

patient <- head(as.character(unlist(lapply(1:ncol(x),function(i) {c(i,i)}))),ncol(x))
tumor <- rep(c(0,1),ncol(x)/2)
ss <- data.frame(patient,tumor)

dds <- DESeqDataSetFromMatrix(countData = x , colData=ss, design = ~ patient + tumor)
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <-cbind(as.data.frame(z),assay(vsd))
dge <- as.data.frame(zz[order(zz$pvalue),])

sig <- subset(dge,padj<0.05)
SIG = nrow(sig)
DN = nrow(subset(sig,log2FoldChange<0))
UP = nrow(subset(sig,log2FoldChange>0))
HEADER = paste("mRNA", SIG , "DEGs,", UP ,"upregulated,", DN, "downregulated")
# smear
plot(log2(dge$baseMean),dge$log2FoldChange,cex=0.6,cex.axis=1.2,cex.lab=1.3,
  xlab="log2 base mean",
  ylab="log2 fold change" ,pch=19,col="#838383")
points(log2(sig$baseMean),sig$log2FoldChange,cex=0.6,pch=19,col="red")
mtext((HEADER),cex=1.2)
top <- head(sig,20)
# volcano
plot(dge$log2FoldChange, -log2(dge$pvalue) ,cex=0.6, cex.lab=1.3,cex.axis=1.2,
 ,xlab="log2 fold change", ylab="-log2 p-value" ,pch=19,col="#838383")
points(sig$log2FoldChange, -log2(sig$pvalue),cex=0.6,pch=19,col="red")
mtext((HEADER),cex=1.2)
# top N gene heatmap
colCols <- rep(c(0,1),ncol(x)/2)
colCols <- gsub("0","gray",colCols)
colCols <- gsub("1","black",colCols)

colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2(  as.matrix(dge[1:50,c(7:ncol(dge))]), col=colfunc(25),scale="row",
  ColSideColors=colCols ,
  trace="none",margins = c(6,10), cexRow=.6, cexCol=.5, main="Top 50 genes")

```

RNA expression pathway enrichment with mitch.

```{r,mitch}

rna <- mitch_import(dge, DEtype="deseq2")

res <- mitch_calc(rna, genesets=gs_symbols, priority="effect",cores=32)

mres <- res$enrichment_result
msig <- subset(mres,p.adjustANOVA<0.05)

head(subset(mres,p.adjustANOVA<0.05 & s.dist>0),20) %>%
  kbl(caption="mRNA upregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

head(subset(mres,p.adjustANOVA<0.05 & s.dist<0),20) %>%
  kbl(caption="mRNA downregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

```

## Load methylation enrichment data and calculate overlap

Use Euler diagrams to find the similarity.
As medians are normally distributed about the zero, one would expect equal numbers of up and down-
methylated gene sets.

```{r,lamfunc}

# LAM aggregate
agg <- function(dm,cores=1) {
  dm <- dm[which(dm$UCSC_RefGene_Name!=""),]
  gn <- unique(unlist(strsplit( dm$UCSC_RefGene_Name ,";")))
  gnl <- strsplit( dm$UCSC_RefGene_Name ,";")
  gnl <- mclapply(gnl,unique,mc.cores=cores)
  dm$UCSC_RefGene_Name <- gnl
  l <- mclapply(1:nrow(dm), function(i) {
    a <- dm[i,]
    len <- length(a[[1]][[1]])
    tvals <- as.numeric(rep(a["t"],len))
    genes <- a[[1]][[1]]
    data.frame(genes,tvals)
  },mc.cores=cores)
  df <- do.call(rbind,l)
  keep <- names(which(table(df$genes)>1))
  df <- df[df$genes %in% keep,]
  gn <- unique(df$genes)
  gme_res <- lapply( 1:length(gn), function(i) {
    g <- gn[i]
    tstats <- df[which(df$genes==g),"tvals"]
    myn <- length(tstats)
    mymean <- mean(tstats)
    mymedian <- median(tstats)
    top <- tstats[order(-abs(tstats))][1]
    if ( length(tstats) > 2 ) {
      ttest <- t.test(tstats)
      pval <- ttest$p.value
    } else {
      pval = 1
    }
    res <- c("gene"=g,"nprobes"=myn,"mean"=mymean,
      "median"=mymedian, "top"=top, "pval"=pval)
  } )
  gme_res_df <- do.call(rbind, gme_res)
  rownames(gme_res_df) <- gme_res_df[,1]
  gme_res_df <- gme_res_df[,-1]
  tmp <- apply(gme_res_df,2,as.numeric)
  rownames(tmp) <- rownames(gme_res_df)
  gme_res_df <- as.data.frame(tmp)
  gme_res_df$sig <- -log10(gme_res_df[,"pval"])
  gme_res_df <- gme_res_df[order(-gme_res_df$sig),]
  gme_res_df$fdr <- p.adjust(gme_res_df$pval)
  out <- list("df"=df,"gme_res_df"=gme_res_df)
  return(out)
}

```

Here we will look at parallel analysis of gene body and promoter.

```{r,promoterbody}

table(dm$Regulatory_Feature_Group)

# all gene
dmagg <- agg(dm=dm,cores=16)
str(dmagg)
head(dmagg$gme_res_df)
dim(dm)
dim(dmagg$gme_res_df)
mm <- dmagg$gme_res_df[,"mean",drop=FALSE]
mmres <- mitch_calc(mm, genesets=gs_symbols, priority="effect",cores=32)
mmres_up <- subset(mmres$enrichment_result,s.dist>0 & p.adjustANOVA < 0.05)
mmres_dn <- subset(mmres$enrichment_result,s.dist<0 & p.adjustANOVA < 0.05)
nrow(mmres_up)
nrow(mmres_dn)
head(mmres_up,20)
head(mmres_dn,20)

# promoter
dmpr <- dm[grep("Promoter",dm$Regulatory_Feature_Group),]
dmpragg <- agg(dm=dmpr,cores=16)
str(dmpragg)
head(dmpragg$gme_res_df)
dim(dmpr)
dim(dmpragg$gme_res_df)
mmpr <- dmpragg$gme_res_df[,"mean",drop=FALSE]
mmprres <- mitch_calc(mmpr, genesets=gs_symbols, priority="effect",cores=32)
mmprres_up <- subset(mmprres$enrichment_result,s.dist>0 & p.adjustANOVA < 0.05)
mmprres_dn <- subset(mmprres$enrichment_result,s.dist<0 & p.adjustANOVA < 0.05)
nrow(mmprres_up)
nrow(mmprres_dn)
head(mmprres_up,20)
head(mmprres_dn,20)

# ex-promoter
dmx <- dm[grep("Promoter",dm$Regulatory_Feature_Group,invert=TRUE),]
dmxagg <- agg(dm=dmx,cores=16)
str(dmxagg)
head(dmxagg$gme_res_df)
dim(dmx)
dim(dmxagg$gme_res_df)
mmx <- dmxagg$gme_res_df[,"mean",drop=FALSE]
mmxres <- mitch_calc(mmx, genesets=gs_symbols, priority="effect",cores=32)
mmxres_up <- subset(mmxres$enrichment_result,s.dist>0 & p.adjustANOVA < 0.05)
mmxres_dn <- subset(mmxres$enrichment_result,s.dist<0 & p.adjustANOVA < 0.05)
nrow(mmxres_up)
nrow(mmxres_dn)
head(mmxres_up,20)
head(mmxres_dn,20)

# combined methylation
mmm <- merge(mmpr,mmx,by=0)
rownames(mmm) <- mmm$Row.names
mmm$Row.names = NULL
colnames(mmm) <- c("pr","x")
mmmres <- mitch_calc(mmm, genesets=gs_symbols, priority="effect",cores=32)
mmmres_sig <- subset(mmmres$enrichment_result, p.adjustMANOVA < 0.05)
head(mmmres_sig,20)
table(sign(mmmres_sig$s.pr) == sign(mmmres_sig$s.x))

```

## RNA expression and methylation enrichment

```{r,rnaoverlap2}

# three way
mmrna <- merge(mmm,rna,by=0)
rownames(mmrna) <- mmrna$Row.names
mmrna$Row.names = NULL
colnames(mmrna) <- c("pr","x","rna")
mmrnares <- mitch_calc(mmrna, genesets=gs_symbols, priority="effect",cores=32)
mmrnares_sig <- subset(mmrnares$enrichment_result, p.adjustMANOVA < 0.05)
head(mmrnares_sig,20)
table(sign(mmrnares_sig$s.pr) == sign(mmrnares_sig$s.x))
table(sign(mmrnares_sig$s.pr) == sign(mmrnares_sig$s.rna))
table(sign(mmrnares_sig$s.x) == sign(mmrnares_sig$s.rna))
mtop <- head(mmrnares_sig,30)
mtop <- mtop[,c("set","s.pr","s.x","s.rna")]
rownames(mtop) <- gsub("REACTOME_","",mtop$set)
rownames(mtop) <- gsub("_"," ",rownames(mtop))
mtop$set=NULL
colnames(mtop) <- c("promoter","non-promoter","RNA")
colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2( as.matrix(mtop) , trace="none",margins = c(6,25), col=colfunc(25), cexRow=.6, cexCol=1)

# rna vs promoter
mmrna <- merge(mmpr,rna,by=0)
rownames(mmrna) <- mmrna$Row.names
mmrna$Row.names = NULL
colnames(mmrna) <- c("pr","rna")
mmrnares <- mitch_calc(mmrna, genesets=gs_symbols, priority="effect",cores=32)
mmrnares_sig <- subset(mmrnares$enrichment_result, p.adjustMANOVA < 0.05)
head(mmrnares_sig,20)
table(sign(mmrnares_sig$s.pr) == sign(mmrnares_sig$s.rna))
mmrnares_sig <- subset(mmrnares$enrichment_result, p.adjustMANOVA < 0.05)
mylm <- lm(mmrnares$enrichment_result$s.rna ~ mmrnares$enrichment_result$s.pr)
summary(mylm)
mylm2 <- lm(mmrnares_sig$s.rna ~ mmrnares_sig$s.pr)
summary(mylm2)
plot(mmrnares$enrichment_result$s.pr,mmrnares$enrichment_result$s.rna,xlab="promoter",ylab="RNA",main="S dist",col="gray",pch=19)
abline(v=0,h=0,lty=2,lwd=2)
points(mmrnares_sig$s.pr,mmrnares_sig$s.rna,col="black",pch=19,cex=1.01)
abline(mylm,col="blue",lty=3,lwd=3)
abline(mylm2,col="red",lty=3,lwd=3)

# rna vs expromoter
mmrna <- merge(mmx,rna,by=0)
rownames(mmrna) <- mmrna$Row.names
mmrna$Row.names = NULL
colnames(mmrna) <- c("x","rna")
mmrnares <- mitch_calc(mmrna, genesets=gs_symbols, priority="effect",cores=32)
mmrnares_sig <- subset(mmrnares$enrichment_result, p.adjustMANOVA < 0.05)
head(mmrnares_sig,20)
table(sign(mmrnares_sig$s.x) == sign(mmrnares_sig$s.rna))
mmrnares_sig <- subset(mmrnares$enrichment_result, p.adjustMANOVA < 0.05)
mylm <- lm(mmrnares$enrichment_result$s.rna ~ mmrnares$enrichment_result$s.x)
summary(mylm)
mylm2 <- lm(mmrnares_sig$s.rna ~ mmrnares_sig$s.x)
summary(mylm2)
plot(mmrnares$enrichment_result$s.x,mmrnares$enrichment_result$s.rna,xlab="ex-promoter",ylab="RNA",main="S dist",col="gray",pch=19)
abline(v=0,h=0,lty=2,lwd=2)
points(mmrnares_sig$s.x,mmrnares_sig$s.rna,col="black",pch=19,cex=1.01)
abline(mylm,col="blue",lty=3,lwd=3)
abline(mylm2,col="red",lty=3,lwd=3)

```

Now look at GSAmeth.

```{r,rnagsa}

# all meth
sigup <- rownames(subset(dm,logFC > 0 & adj.P.Val < 0.05))
sigdn <- rownames(subset(dm,logFC < 0 & adj.P.Val < 0.05))
gsaup <- gsameth(sig.cpg=sigup, all.cpg=rownames(dm), collection=gs_entrez, array.type="EPIC")
rownames(gsaup) <- paste(rownames(gsaup),"UP")
gsadn <- gsameth(sig.cpg=sigdn, all.cpg=rownames(dm), collection=gs_entrez, array.type="EPIC")
rownames(gsadn) <- paste(rownames(gsadn),"DN")
gsig_up <- subset(gsaup,FDR<0.05)
gsig_dn <- subset(gsadn,FDR<0.05)
gsig <- rbind(gsig_up,gsig_dn)
nrow(gsig)

# all meth
sigup <- rownames(subset(dm,logFC > 0 & adj.P.Val < 0.05))
sigdn <- rownames(subset(dm,logFC < 0 & adj.P.Val < 0.05))
gsaup <- gsameth(sig.cpg=sigup, all.cpg=rownames(dm), collection=gs_entrez, array.type="EPIC")
rownames(gsaup) <- paste(rownames(gsaup),"UP")
gsadn <- gsameth(sig.cpg=sigdn, all.cpg=rownames(dm), collection=gs_entrez, array.type="EPIC")
rownames(gsadn) <- paste(rownames(gsadn),"DN")
gsig_up <- subset(gsaup,FDR<0.05)
gsig_dn <- subset(gsadn,FDR<0.05)
gsig <- rbind(gsig_up,gsig_dn) 
nrow(gsig)

```

## Conclusion

Although there were some differentially methylated pathways analysed with different methods,
they were not concordant with differential expression.

## Session information

```{r,save}

sessionInfo()

```