### 2023-11-21 figs/GSE158422_GSAfix.Rmd
Prefer UCSC annotations

---
title: "Can GSAmeth be fixed?"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

In the previous analysis, it was shown that sensitivity of GSAmeth was poor compared to LAM with real data.
This was surprising as GSAmeth performed well with simulated data.

Possible explanations to this:

* GSAmeth filters out genes that do not have annotations.
Ideally this filtering should be deactivated.

* GSAmeth only performs FDR correction on gene sets where a minimum number of genes in the foreground.
Ideally, FDR correction should be applied to genesets with a minimum number of genes in the background.

Let's see if we can fix it.

## Requirements

```{r,packages}

suppressPackageStartupMessages({
  library("limma")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("gridExtra")
  library("png")
  source("gsameth_fix.R")
})

CORES=24

```

## Load data

* annotations

* probe sets

* gene sets

* design matrix

* mval matrix

```{r,annotation}

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])

gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
sets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))

summary(unlist(lapply(sets,length)))

genesets <- gmt_import("https://ziemann-lab.net/public/gmea_prototype/ReactomePathways.gmt")

if (!file.exists("GSE158422_design.rds")) {
  download.file("https://ziemann-lab.net/public/gmea_prototype/GSE158422_design.rds", "GSE158422_design.rds")
}
design <- readRDS("GSE158422_design.rds")

if (!file.exists("GSE158422_design.rds")) {
 download.file("https://ziemann-lab.net/public/gmea_prototype/GSE158422_mx.rds","GSE158422_mx.rds")
}
mval <- readRDS("GSE158422_mx.rds")

mval_means <- list("tumor"=matrix(colMeans(mval),ncol=2)[,1],"normal"=matrix(colMeans(mval),ncol=2)[,2])
boxplot(mval_means, main="mean probe methylation mval",cex=0,col="white")
beeswarm(mval_means,add=TRUE,pch=1,cex=1.5)
lapply(mval_means,mean)

```

We will assume mean probe methylation is a proxy measure for global methylation.
The boxplot appears to show higher values in tumour samples, based on median and quartiles.
However the mean values are actually lower in tumour samples compared to normal.

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,gsdl}

gs_entrez <- gmt_import("c2.cp.reactome.v2023.1.Hs.entrez.gmt")

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

# AL GMEA: all probes

Start with curating the sample info and design matrix.

```{r,samplesheet}

sex <- as.data.frame(design)$sex
tumor <- as.data.frame(design)$tumor
patient <- as.character(unlist(lapply(1:ncol(mval),function(i) {c(i,i)})))
patient <- head(patient,ncol(mval))
design <- model.matrix(~ patient + tumor )

```

## Functions

limma-average-mitch (LAM).

```{r,limmafunc}

# limma
runlimma <- function(mval,design,myann) {
  fit.reduced <- lmFit(mval,design)
  fit.reduced <- eBayes(fit.reduced)
  dm <- topTable(fit.reduced,coef=ncol(design), number = Inf)
  dm <- merge(myann,dm,by=0)
  dm <- dm[order(dm$P.Value),]
  rownames(dm) <- dm$Row.names
  dm$Row.names=NULL
  return(dm)
}

```

The following function is key for the aggregation of probe t-statistic values by gene using the arithmetic mean.
In this step, ~800k probe values are summarised to ~28k genes.

```{r,lamfunction}

# LAM aggregate
agg <- function(dm,cores=1) {
  dm <- dm[which(dm$UCSC_RefGene_Name!=""),]
  gn <- unique(unlist(strsplit( dm$UCSC_RefGene_Name ,";")))
  gnl <- strsplit( dm$UCSC_RefGene_Name ,";")
  gnl <- mclapply(gnl,unique,mc.cores=cores)
  dm$UCSC_RefGene_Name <- gnl
  l <- mclapply(1:nrow(dm), function(i) {
    a <- dm[i,]
    len <- length(a[[1]][[1]])
    tvals <- as.numeric(rep(a["t"],len))
    genes <- a[[1]][[1]]
    data.frame(genes,tvals)
  },mc.cores=cores)
  df <- do.call(rbind,l)
  keep <- names(which(table(df$genes)>1))
  df <- df[df$genes %in% keep,]
  gn <- unique(df$genes)
  gme_res <- lapply( 1:length(gn), function(i) {
    g <- gn[i]
    tstats <- df[which(df$genes==g),"tvals"]
    myn <- length(tstats)
    mymean <- mean(tstats)
    mymedian <- median(tstats)
    top <- tstats[order(-abs(tstats))][1]
    if ( length(tstats) > 2 ) {
      ttest <- t.test(tstats)
      pval <- ttest$p.value
    } else {
      pval = 1
    }
    res <- c("gene"=g,"nprobes"=myn,"mean"=mymean,
      "median"=mymedian, "top"=top, "pval"=pval)
  } )
  gme_res_df <- do.call(rbind, gme_res)
  rownames(gme_res_df) <- gme_res_df[,1]
  gme_res_df <- gme_res_df[,-1]
  tmp <- apply(gme_res_df,2,as.numeric)
  rownames(tmp) <- rownames(gme_res_df)
  gme_res_df <- as.data.frame(tmp)
  gme_res_df$sig <- -log10(gme_res_df[,"pval"])
  gme_res_df <- gme_res_df[order(-gme_res_df$sig),]
  gme_res_df$fdr <- p.adjust(gme_res_df$pval)
  out <- list("df"=df,"gme_res_df"=gme_res_df)
  return(out)
}

```

Function to run mitch analysis.

```{r,mitchfunc}

runmitch <- function(m,genesets,cores=1) {
  suppressMessages({ mres <- mitch_calc(m,genesets,minsetsize=5,cores=cores, priority="effect") })
  mres <- mres$enrichment_result
  rownames(mres) <- mres$set
  mres$set=NULL
  return(mres)
}

```
## Run limma

Run limma and look at the absolute effect sizes.

```{r,limma1}

dim(mval)
dim(design)

tic()
dm <- runlimma(mval=mval,design=design,myann=myann)
toc() ## mean(23.7 26.8 25.4)=23.7
top <- head(dm,20)
top <- top[order(top$logFC),]

dim(dm)
nrow(subset(dm,adj.P.Val<0.05 & logFC>0))
nrow(subset(dm,adj.P.Val<0.05 & logFC<0))

length(unique(unlist(strsplit(subset(dm,adj.P.Val<0.05 & logFC>0)$UCSC_RefGene_Name,";"))))
length(unique(unlist(strsplit(subset(dm,adj.P.Val<0.05 & logFC<0)$UCSC_RefGene_Name,";"))))

length(unique(unlist(strsplit(dm$UCSC_RefGene_Name,";"))))

```

# GSAMETH

Run GSA meth on the full dataset.

```{r,rungsameth1}

dm <- runlimma(mval,design,myann)

hist(dm$t,breaks=100,xlim=c(-10,10)) ; grid()

head(dm, 10) %>% kbl(caption="Top 10 probes by p-value") %>% kable_paper("hover", full_width = F)

# the number of DM probes impact the results drastically
sizes=c(1000,5000,10000,20000,50000,100000,200000,300000,350000,400000,450000,500000)
nsigs <- lapply(sizes, function(z) {
  dmsig <- head(dm,z)
  dmsigup <- rownames(subset(dmsig,logFC>0))
  dmsigdn <- rownames(subset(dmsig,logFC<0))
  gsaup <- gsameth(sig.cpg=dmsigup, all.cpg=rownames(dm), collection=gs_entrez)
  rownames(gsaup) <- paste(rownames(gsaup),"UP")
  gsadn <- gsameth(sig.cpg=dmsigdn, all.cpg=rownames(dm), collection=gs_entrez)
  rownames(gsadn) <- paste(rownames(gsadn),"DN")
  nsig <- rownames(subset(rbind(gsaup,gsadn),FDR<0.05))
  return(nsig)
})

names(nsigs) <- sizes
nsigs

par( mar = c(5.1, 6.1, 4.1, 2.1) )
barplot(unlist(lapply(nsigs,length)),horiz=TRUE,las=1,xlab="no. significant sets") ; grid()
mtext("no. probes selected", 2, 4)

# the fold change threshold also impacts the results
FCLIM <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
ngsig <- lapply(FCLIM,function(F) {
  sigup <- rownames(subset(dm,logFC > F & adj.P.Val < 0.05))
  sigdn <- rownames(subset(dm,logFC < -F & adj.P.Val < 0.05))
  gsaup <- gsameth(sig.cpg=sigup, all.cpg=rownames(dm), collection=gs_entrez)
  rownames(gsaup) <- paste(rownames(gsaup),"UP")
  gsadn <- gsameth(sig.cpg=sigdn, all.cpg=rownames(dm), collection=gs_entrez)
  rownames(gsadn) <- paste(rownames(gsadn),"DN")
  ngsig <- rownames(subset(rbind(gsaup,gsadn),FDR<0.05))
  return(ngsig)
})

names(ngsig) <- FCLIM
ngsig

par( mar = c(5.1, 6.1, 4.1, 2.1) )
barplot(unlist(lapply(ngsig,length)),horiz=TRUE,las=1,xlab="no. significant sets",xlim=c(0,60)) ; grid()
mtext("fold change threshold", 2, 4)

sigup <- rownames(subset(dm,logFC > 0 & adj.P.Val < 0.05))
sigdn <- rownames(subset(dm,logFC < 0 & adj.P.Val < 0.05))

tic()
gsaup <- gsameth(sig.cpg=sigup, all.cpg=rownames(dm), collection=gs_entrez)
rownames(gsaup) <- paste(rownames(gsaup),"UP")
gsadn <- gsameth(sig.cpg=sigdn, all.cpg=rownames(dm), collection=gs_entrez)
rownames(gsadn) <- paste(rownames(gsadn),"DN")
toc() # 12.9, 12.6, 11.5

gsaup <- gsaup[order(gsaup$P.DE),]
gsadn <- gsadn[order(gsadn$P.DE),]

gsig_up <- subset(gsaup,FDR<0.05)
head(gsig_up,20) %>% kbl(caption="Hypermethylated pathways") %>% kable_paper("hover", full_width = F)

gsig_dn <- subset(gsadn,FDR<0.05)
head(gsig_dn,20) %>% kbl(caption="Hypomethylated pathways") %>% kable_paper("hover", full_width = F)

gsig <- rbind(gsig_up,gsig_dn)
gsigsets <- rownames(gsig)

```

Unload missmethyl install new one with fix.

```{r,gsafgix}

detach("package:missMethyl", unload=TRUE)
install.packages("../pkgs/missMethyl", repos = NULL, type = "source")
library("missMethyl")

gsafixup <- gsameth(sig.cpg=sigup, all.cpg=rownames(dm), collection=gs_entrez)
rownames(gsafixup) <- paste(rownames(gsafixup),"UP")
gsafixdn <- gsameth(sig.cpg=sigdn, all.cpg=rownames(dm), collection=gs_entrez)
rownames(gsafixdn) <- paste(rownames(gsafixdn),"DN")

gsafixup <- gsafixup[order(gsafixup$P.DE),]
gsafixdn <- gsafixdn[order(gsafixdn$P.DE),]

gsigfix_up <- subset(gsafixup,FDR<0.05)
head(gsigfix_up,20) %>% kbl(caption="Hypermethylated pathways (fix)") %>% kable_paper("hover", full_width = F)

gsigfix_dn <- subset(gsafixdn,FDR<0.05)
head(gsigfix_dn,20) %>% kbl(caption="Hypomethylated pathways (fix)") %>% kable_paper("hover", full_width = F)

gsigfix <- rbind(gsigfix_up,gsigfix_dn)

gsigfixsets <- rownames(gsigfix)

# the number of DM probes impact the results drastically
sizes=c(1000,5000,10000,20000,50000,100000,200000,300000,350000,400000,450000,500000)
nsigs <- lapply(sizes, function(z) {
  dmsig <- head(dm,z)
  dmsigup <- rownames(subset(dmsig,logFC>0))
  dmsigdn <- rownames(subset(dmsig,logFC<0))
  gsaup <- gsameth(sig.cpg=dmsigup, all.cpg=rownames(dm), collection=gs_entrez)
  rownames(gsaup) <- paste(rownames(gsaup),"UP")
  gsadn <- gsameth(sig.cpg=dmsigdn, all.cpg=rownames(dm), collection=gs_entrez)
  rownames(gsadn) <- paste(rownames(gsadn),"DN")
  nsig <- rownames(subset(rbind(gsaup,gsadn),FDR<0.05))
  return(nsig)
})

names(nsigs) <- sizes
nsigs

par( mar = c(5.1, 6.1, 4.1, 2.1) )
barplot(unlist(lapply(nsigs,length)),horiz=TRUE,las=1,xlab="no. significant sets") ; grid()
mtext("no. probes selected", 2, 4)

# the fold change threshold also impacts the results
FCLIM <- c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1)
ngsig <- lapply(FCLIM,function(F) {
  sigup <- rownames(subset(dm,logFC > F & adj.P.Val < 0.05))
  sigdn <- rownames(subset(dm,logFC < -F & adj.P.Val < 0.05))
  gsaup <- gsameth(sig.cpg=sigup, all.cpg=rownames(dm), collection=gs_entrez)
  rownames(gsaup) <- paste(rownames(gsaup),"UP")
  gsadn <- gsameth(sig.cpg=sigdn, all.cpg=rownames(dm), collection=gs_entrez)
  rownames(gsadn) <- paste(rownames(gsadn),"DN")
  ngsig <- rownames(subset(rbind(gsaup,gsadn),FDR<0.05))
  return(ngsig)
})

names(ngsig) <- FCLIM
ngsig

par( mar = c(5.1, 6.1, 4.1, 2.1) )
barplot(unlist(lapply(ngsig,length)),horiz=TRUE,las=1,xlab="no. significant sets",xlim=c(0,60)) ; grid()
mtext("fold change threshold", 2, 4)

```

## Session information

```{r,save}

save.image("GSE158422_sensitivity.Rdata")

sessionInfo()

```