### 2023-11-21 figs/GSE158422_pkgdev.Rmd
Prefer UCSC annotations

---
title: "GSE158422: Package Development"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Here we are looking at putting together the easiest possible workflow for the package.

## Requirements

```{r,packages}

suppressPackageStartupMessages({
  library("limma")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("gridExtra")
  library("png")
})

CORES=24

```

## Load data

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,load1}

if (!file.exists("GSE158422_design.rds")) {
  download.file("https://ziemann-lab.net/public/gmea_prototype/GSE158422_design.rds", "GSE158422_design.rds")
}
design <- readRDS("GSE158422_design.rds")

if (!file.exists("GSE158422_design.rds")) {
 download.file("https://ziemann-lab.net/public/gmea_prototype/GSE158422_mx.rds","GSE158422_mx.rds")
}
mval <- readRDS("GSE158422_mx.rds")

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

## Curate the annotation

Use all probes on the chip.

```{r,anno1}

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","GencodeCompV12_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #9.0s

```

# AL GMEA: all probes

Start with curating the sample info and design matrix.

```{r,samplesheet}

sex <- as.data.frame(design)$sex
tumor <- as.data.frame(design)$tumor
patient <- as.character(unlist(lapply(1:ncol(mval),function(i) {c(i,i)})))
patient <- head(patient,ncol(mval))
design <- model.matrix(~ patient + tumor )

```

## Prep limma

```{r,limma1}

tic()
fit.reduced <- lmFit(mval,design)
fit.reduced <- eBayes(fit.reduced)
dm <- topTable(fit.reduced,coef=ncol(design), number = Inf)
head(dm)
toc() #25.53, 25.50, 25.51

```

## Merge gene name and aggregate

This only works properly when mitch's mapGeneIds function uses `mean` to aggregate (line 69 of mitch.R).

```{r,mg1}

# mitch aggregate function needs to be changes to `mean`
tic()
m <- mitch_import(x=dm,DEtype="limma",geneTable=gt)
toc() #9.58s, 9.71, 9.69
head(m)

```

This is the alternative aggregation approach which is faster but more complicated.

```{r,mg2,eval=FALSE}

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","GencodeCompV12_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
dim(gt)
str(gt)
toc() #2.44, 2.41, 2.41s

```

##  Run mitch analysis.

```{r,mitch1}

tic()
mres <- mitch_calc(x=m,genesets=gs_symbols,minsetsize=5,cores=16, priority="effect")
toc() #10.92, 10.733, 10.91

mtable <- mres$enrichment_result
up <- subset(mtable,s.dist>0 & p.adjustANOVA<0.05)
dn <- subset(mtable,s.dist<0 & p.adjustANOVA<0.05)
nrow(up)
nrow(dn)
head(up,10)
head(dn,10)

top <- rbind(up[1:10,],dn[1:10,])
top <- top[order(top$s.dist),]

barnames <- gsub("_"," ",gsub("REACTOME_","",top$set))
cols <- as.character(((sign(top$s.dist)+1)/2)+1)
cols <- gsub("1","blue",gsub("2","red",cols))

par(mar = c(5.1, 29.1, 4.1, 2.1))
barplot(abs(top$s.dist),horiz=TRUE,las=1,names.arg=barnames, cex.names=0.8,
  cex.axis=0.8,col=cols, xlab="Enrichment score",main="Reactomes")
mtext("whole gene")
grid()

par( mar = c(5.1, 4.1, 4.1, 2.1) )

```

Make a html report and some charts.

```{r,report1}

mitch_report(res=mres,outfile="GSE158422_pkgdev_mitchreport.html",overwrite=TRUE)
mitch_plots(res=mres,outfile="GSE158422_pkgdev_mitchcharts.pdf")

```

## Promoter only analysis

Now do the same but filter for promoters.
Using the UCSC_RefGene_Name field of the annotation and using TSS probes only.

```{r,promo}

head(myann)
table(anno$UCSC_RefGene_Group)
table(unlist(strsplit(myann$GencodeCompV12_Group,";")))

# promoter prep
tic()
keep=c("TSS200","TSS1500")
annos <- strsplit(myann$GencodeCompV12_Group,";")
genes <- strsplit(myann$UCSC_RefGene_Name,";")
keeps <- lapply(1:nrow(myann), function(i) {
  unique(genes[[i]][annos[[i]] %in% keep])
})
names(keeps) <- rownames(myann)
#gt <- as.data.frame((unlist(gp2)))
gt2 <- stack(gp2)
colnames(gt2) <- c("gene","probe")
toc() #20.37, 19.56, 21.23

tic()
m2 <- mitch_import(x=dm,DEtype="limma",geneTable=gt2)
toc() #9.68, 9.78, 9.62
head(m)
dim(m)

tic()
mres <- mitch_calc(x=m2,genesets=gs_symbols,minsetsize=5,cores=16, priority="effect")
toc() #11.12, 10.99, 11.01

mtable <- mres$enrichment_result
up <- subset(mtable,s.dist>0 & p.adjustANOVA<0.05)
dn <- subset(mtable,s.dist<0 & p.adjustANOVA<0.05)
nrow(up)
nrow(dn)
head(up,10)
head(dn,10)

top <- rbind(up[1:10,],dn[1:10,])
top<- top[order(top$s.dist),]

barnames <- gsub("_"," ",gsub("REACTOME_","",top$set))
cols <- as.character(((sign(top$s.dist)+1)/2)+1)
cols <- gsub("1","blue",gsub("2","red",cols))

par(mar = c(5.1, 29.1, 4.1, 2.1))
barplot(abs(top$s.dist),horiz=TRUE,las=1,names.arg=barnames, cex.names=0.8,
  cex.axis=0.8,col=cols, xlab="Enrichment score",main="Reactomes")
mtext("promoter only")
grid()

par( mar = c(5.1, 4.1, 4.1, 2.1) )

```

## Session information

```{r,save}

save.image("GSE158422_pkgdev.Rdata")

sessionInfo()

```