### 2023-11-28 figs/GSE158422_rna_int.Rmd
progress on promoter versus body analysis

---
title: "GSE158422: DNA methylation in Lung squamous cell carcinoma - RNA expression and integration"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Let's take a look at how pathway methylation associates with gene expression.

Partition methylation analysis between promoter and body.

Will need to do this at the annotation level.

Compare to GSAmeth.

```{r,packages}

suppressPackageStartupMessages({
  library("DESeq2")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("limma")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("HGNChelper")
  library("gplots")
})

CORES=12

```

## Load annotations and methylation data

```{r,annotation}

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
sets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))
summary(unlist(lapply(sets,length)))

design <- readRDS("GSE158422_design.rds")
mval <- readRDS("GSE158422_mx.rds")

runlimma <- function(mval,design,myann) {
  fit.reduced <- lmFit(mval,design)
  fit.reduced <- eBayes(fit.reduced)
  dm <- topTable(fit.reduced,coef=ncol(design), number = Inf)
  dm <- merge(myann,dm,by=0)
  dm <- dm[order(dm$P.Value),]
  rownames(dm) <- dm$Row.names
  dm$Row.names=NULL
  return(dm)
}

dm <- runlimma(mval=mval,design=design,myann=myann)

```

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,gsdl}

gs_entrez <- gmt_import("c2.cp.reactome.v2023.1.Hs.entrez.gmt")

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

## Load RNA expression

From GEO https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE158420
GSE158420_counts.txt.gz

## DESEQ2

```{r,deseq1}

x <- read.table("GSE158420_counts.txt.gz",row.names=NULL)
dim(x)
x <- aggregate(. ~ row.names,x,sum)
dim(x)

#RUFKM
head(x)

```

There are some gene name errors, so I will use HGNChelper to repair them.

```{r,gene_name_fix}

human=x$row.names
fix <- checkGeneSymbols(human)
x$gene <- fix$Suggested.Symbol
x$row.names=NULL
x <- aggregate(. ~ gene,x,sum)
dim(x)
row.names(x)=x$gene
x$gene=NULL
head(x)

```

Now do DESeq2

```{r,deseq2}

patient <- head(as.character(unlist(lapply(1:ncol(x),function(i) {c(i,i)}))),ncol(x))
tumor <- rep(c(0,1),ncol(x)/2)
ss <- data.frame(patient,tumor)

dds <- DESeqDataSetFromMatrix(countData = x , colData=ss, design = ~ patient + tumor)
res <- DESeq(dds)
z <- results(res)
vsd <- vst(dds, blind=FALSE)
zz <-cbind(as.data.frame(z),assay(vsd))
dge <- as.data.frame(zz[order(zz$pvalue),])

sig <- subset(dge,padj<0.05)
SIG = nrow(sig)
DN = nrow(subset(sig,log2FoldChange<0))
UP = nrow(subset(sig,log2FoldChange>0))
HEADER = paste("mRNA", SIG , "DEGs,", UP ,"upregulated,", DN, "downregulated")
# smear
plot(log2(dge$baseMean),dge$log2FoldChange,cex=0.6,cex.axis=1.2,cex.lab=1.3,
  xlab="log2 base mean",
  ylab="log2 fold change" ,pch=19,col="#838383")
points(log2(sig$baseMean),sig$log2FoldChange,cex=0.6,pch=19,col="red")
mtext((HEADER),cex=1.2)
top <- head(sig,20)
# volcano
plot(dge$log2FoldChange, -log2(dge$pvalue) ,cex=0.6, cex.lab=1.3,cex.axis=1.2,
 ,xlab="log2 fold change", ylab="-log2 p-value" ,pch=19,col="#838383")
points(sig$log2FoldChange, -log2(sig$pvalue),cex=0.6,pch=19,col="red")
mtext((HEADER),cex=1.2)
# top N gene heatmap
colCols <- rep(c(0,1),ncol(x)/2)
colCols <- gsub("0","gray",colCols)
colCols <- gsub("1","black",colCols)

colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2(  as.matrix(dge[1:50,c(7:ncol(dge))]), col=colfunc(25),scale="row",
  ColSideColors=colCols ,
  trace="none",margins = c(6,10), cexRow=.6, cexCol=.5, main="Top 50 genes")

```

RNA expression pathway enrichment with mitch.

```{r,mitch}

rna <- mitch_import(dge, DEtype="deseq2")

res <- mitch_calc(rna, genesets=gs_symbols, priority="effect",cores=32)

mres <- res$enrichment_result
msig <- subset(mres,p.adjustANOVA<0.05)

head(subset(mres,p.adjustANOVA<0.05 & s.dist>0),20) %>%
  kbl(caption="mRNA upregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

head(subset(mres,p.adjustANOVA<0.05 & s.dist<0),20) %>%
  kbl(caption="mRNA downregulated pathways") %>%
  kable_styling("hover",full_width=FALSE)

```

## Load methylation enrichment data and calculate overlap

Use Euler diagrams to find the similarity.
As medians are normally distributed about the zero, one would expect equal numbers of up and down-
methylated gene sets.

## Curate the annotation

Use all probes on the chip.

```{r,anno1}

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
gt$probe <- as.character(gt$probe)
dim(gt)
str(gt)

```

Here we will look at parallel analysis of gene body and promoter.

The approach here is to generate two mapping tables of probes to genes based on the annotation which is split
into gene body and promoter (TSS).

gene body=gb

gene promoter (TSS) = gp

```{r,promoterbody}

# partition body and promoter data(probes)
dmg <- strsplit(dm$UCSC_RefGene_Name,";")

body <- lapply(1:nrow(dm),function(i) {
  a <- dm[i,"UCSC_RefGene_Group"]
  a <- unlist(strsplit(a,";"))
  dmg[[i]][a=="Body"]
})
names(body) <- rownames(dm)
body <- lapply(body,unique)
gb <- stack(body)
colnames(gb) <- c("gene","probe")
gb$probe <- as.character(gb$probe)
dim(gb)
str(gb)

tss <- lapply(1:nrow(dm),function(i) {
  a <- dm[i,"UCSC_RefGene_Group"]
  a <- unlist(strsplit(a,";"))
  dmg[[i]][grep("TSS",a)]
})
names(tss) <- rownames(dm)
tss <- lapply(tss,unique)
gp <- stack(tss)
colnames(gp) <- c("gene","probe")
gp$probe <- as.character(gp$probe)
dim(gp)
str(gp)

# HERENOW

# all gene
dmagg <- agg(dm=dm,cores=16)
str(dmagg)
head(dmagg$gme_res_df)
dim(dm)
dim(dmagg$gme_res_df)
mm <- dmagg$gme_res_df[,"mean",drop=FALSE]
mmres <- mitch_calc(mm, genesets=gs_symbols, priority="effect",cores=32)
mmres_up <- subset(mmres$enrichment_result,s.dist>0 & p.adjustANOVA < 0.05)
mmres_dn <- subset(mmres$enrichment_result,s.dist<0 & p.adjustANOVA < 0.05)
nrow(mmres_up)
nrow(mmres_dn)
head(mmres_up,20)
head(mmres_dn,20)

# promoter
dmpr <- dm[grep("Promoter",dm$UCSC_RefGene_Group),]
dmpragg <- agg(dm=dmpr,cores=16)
str(dmpragg)
head(dmpragg$gme_res_df)
dim(dmpr)
dim(dmpragg$gme_res_df)
mmpr <- dmpragg$gme_res_df[,"mean",drop=FALSE]
mmprres <- mitch_calc(mmpr, genesets=gs_symbols, priority="effect",cores=32)
mmprres_up <- subset(mmprres$enrichment_result,s.dist>0 & p.adjustANOVA < 0.05)
mmprres_dn <- subset(mmprres$enrichment_result,s.dist<0 & p.adjustANOVA < 0.05)
nrow(mmprres_up)
nrow(mmprres_dn)
head(mmprres_up,20)
head(mmprres_dn,20)

# ex-promoter
dmx <- dm[grep("Promoter",dm$UCSC_RefGene_Group,invert=TRUE),]
dmxagg <- agg(dm=dmx,cores=16)
str(dmxagg)
head(dmxagg$gme_res_df)
dim(dmx)
dim(dmxagg$gme_res_df)
mmx <- dmxagg$gme_res_df[,"mean",drop=FALSE]
mmxres <- mitch_calc(mmx, genesets=gs_symbols, priority="effect",cores=32)
mmxres_up <- subset(mmxres$enrichment_result,s.dist>0 & p.adjustANOVA < 0.05)
mmxres_dn <- subset(mmxres$enrichment_result,s.dist<0 & p.adjustANOVA < 0.05)
nrow(mmxres_up)
nrow(mmxres_dn)
head(mmxres_up,20)
head(mmxres_dn,20)

# combined methylation
mmm <- merge(mmpr,mmx,by=0)
rownames(mmm) <- mmm$Row.names
mmm$Row.names = NULL
colnames(mmm) <- c("pr","x")
mmmres <- mitch_calc(mmm, genesets=gs_symbols, priority="effect",cores=32)
mmmres_sig <- subset(mmmres$enrichment_result, p.adjustMANOVA < 0.05)
head(mmmres_sig,20)
table(sign(mmmres_sig$s.pr) == sign(mmmres_sig$s.x))

```

## RNA expression and methylation enrichment

```{r,rnaoverlap2}

# three way
mmrna <- merge(mmm,rna,by=0)
rownames(mmrna) <- mmrna$Row.names
mmrna$Row.names = NULL
colnames(mmrna) <- c("pr","x","rna")
mmrnares <- mitch_calc(mmrna, genesets=gs_symbols, priority="effect",cores=32)
mmrnares_sig <- subset(mmrnares$enrichment_result, p.adjustMANOVA < 0.05)
head(mmrnares_sig,20)
table(sign(mmrnares_sig$s.pr) == sign(mmrnares_sig$s.x))
table(sign(mmrnares_sig$s.pr) == sign(mmrnares_sig$s.rna))
table(sign(mmrnares_sig$s.x) == sign(mmrnares_sig$s.rna))
mtop <- head(mmrnares_sig,30)
mtop <- mtop[,c("set","s.pr","s.x","s.rna")]
rownames(mtop) <- gsub("REACTOME_","",mtop$set)
rownames(mtop) <- gsub("_"," ",rownames(mtop))
mtop$set=NULL
colnames(mtop) <- c("promoter","non-promoter","RNA")
colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2( as.matrix(mtop) , trace="none",margins = c(6,25), col=colfunc(25), cexRow=.6, cexCol=1)

# rna vs promoter
mmrna <- merge(mmpr,rna,by=0)
rownames(mmrna) <- mmrna$Row.names
mmrna$Row.names = NULL
colnames(mmrna) <- c("pr","rna")
mmrnares <- mitch_calc(mmrna, genesets=gs_symbols, priority="effect",cores=32)
mmrnares_sig <- subset(mmrnares$enrichment_result, p.adjustMANOVA < 0.05)
head(mmrnares_sig,20)
table(sign(mmrnares_sig$s.pr) == sign(mmrnares_sig$s.rna))
mmrnares_sig <- subset(mmrnares$enrichment_result, p.adjustMANOVA < 0.05)
mylm <- lm(mmrnares$enrichment_result$s.rna ~ mmrnares$enrichment_result$s.pr)
summary(mylm)
mylm2 <- lm(mmrnares_sig$s.rna ~ mmrnares_sig$s.pr)
summary(mylm2)
plot(mmrnares$enrichment_result$s.pr,mmrnares$enrichment_result$s.rna,xlab="promoter",ylab="RNA",main="S dist",col="gray",pch=19)
abline(v=0,h=0,lty=2,lwd=2)
points(mmrnares_sig$s.pr,mmrnares_sig$s.rna,col="black",pch=19,cex=1.01)
abline(mylm,col="blue",lty=3,lwd=3)
abline(mylm2,col="red",lty=3,lwd=3)

# rna vs expromoter
mmrna <- merge(mmx,rna,by=0)
rownames(mmrna) <- mmrna$Row.names
mmrna$Row.names = NULL
colnames(mmrna) <- c("x","rna")
mmrnares <- mitch_calc(mmrna, genesets=gs_symbols, priority="effect",cores=32)
mmrnares_sig <- subset(mmrnares$enrichment_result, p.adjustMANOVA < 0.05)
head(mmrnares_sig,20)
table(sign(mmrnares_sig$s.x) == sign(mmrnares_sig$s.rna))
mmrnares_sig <- subset(mmrnares$enrichment_result, p.adjustMANOVA < 0.05)
mylm <- lm(mmrnares$enrichment_result$s.rna ~ mmrnares$enrichment_result$s.x)
summary(mylm)
mylm2 <- lm(mmrnares_sig$s.rna ~ mmrnares_sig$s.x)
summary(mylm2)
plot(mmrnares$enrichment_result$s.x,mmrnares$enrichment_result$s.rna,xlab="ex-promoter",ylab="RNA",main="S dist",col="gray",pch=19)
abline(v=0,h=0,lty=2,lwd=2)
points(mmrnares_sig$s.x,mmrnares_sig$s.rna,col="black",pch=19,cex=1.01)
abline(mylm,col="blue",lty=3,lwd=3)
abline(mylm2,col="red",lty=3,lwd=3)

```

Now look at GSAmeth.

```{r,rnagsa}

# all meth
sigup <- rownames(subset(dm,logFC > 0 & adj.P.Val < 0.05))
sigdn <- rownames(subset(dm,logFC < 0 & adj.P.Val < 0.05))
gsaup <- gsameth(sig.cpg=sigup, all.cpg=rownames(dm), collection=gs_entrez, array.type="EPIC")
rownames(gsaup) <- paste(rownames(gsaup),"UP")
gsadn <- gsameth(sig.cpg=sigdn, all.cpg=rownames(dm), collection=gs_entrez, array.type="EPIC")
rownames(gsadn) <- paste(rownames(gsadn),"DN")
gsig_up <- subset(gsaup,FDR<0.05)
gsig_dn <- subset(gsadn,FDR<0.05)
gsig <- rbind(gsig_up,gsig_dn)
nrow(gsig)

# all meth
sigup <- rownames(subset(dm,logFC > 0 & adj.P.Val < 0.05))
sigdn <- rownames(subset(dm,logFC < 0 & adj.P.Val < 0.05))
gsaup <- gsameth(sig.cpg=sigup, all.cpg=rownames(dm), collection=gs_entrez, array.type="EPIC")
rownames(gsaup) <- paste(rownames(gsaup),"UP")
gsadn <- gsameth(sig.cpg=sigdn, all.cpg=rownames(dm), collection=gs_entrez, array.type="EPIC")
rownames(gsadn) <- paste(rownames(gsadn),"DN")
gsig_up <- subset(gsaup,FDR<0.05)
gsig_dn <- subset(gsadn,FDR<0.05)
gsig <- rbind(gsig_up,gsig_dn) 
nrow(gsig)

```

## Conclusion

Although there were some differentially methylated pathways analysed with different methods,
they were not concordant with differential expression.

## Session information

```{r,save}

sessionInfo()

```