### 2023-12-18 figs/aging_example.Rmd
adding aging example

---
title: "Aging and DNA meth (https://doi.org/10.1186/s13073-019-0693-z)"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Here we are looking at putting together the easiest possible workflow for the package.

## Requirements

```{r,packages}

suppressPackageStartupMessages({
  library("limma")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("mitch")
  library("gplots")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("gridExtra")
  library("png")
})

CORES=16

```

## Load data

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

We'll also score the probes by the signed -log pvalue

```{r,load1}

if(!file.exists("aging_TableS1.csv")){
  options(timeout=1000000)
  download.file("https://datashare.ed.ac.uk/bitstream/handle/10283/3507/Supplementary_Table1_Limma_Age_EWAS_w1w3_Autosomes.csv?sequence=10&isAllowed=y",
    destfile="aging_TableS1.csv")
}

x <- read.csv("aging_TableS1.csv",header=TRUE)
head(x)
y1 <- x[,"Discovery.t",drop=FALSE]
rownames(y1) <- x$ProbeID
y2 <- x[,"Replication.t",drop=FALSE]
rownames(y2) <- x$ProbeID
head(y1)
head(y2)
l <- list("Discovery"=y1,"Replication"=y2)
str(l)

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

## Curate the annotation

Use all probes on the chip.

```{r,anno1}

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
gt$probe <- as.character(gt$probe)
dim(gt)
str(gt)
toc() #9.0s

```

## Merge gene name and aggregate

This only works properly when mitch's mapGeneIds function uses `mean` to aggregate (line 69 of mitch.R).

```{r,mg1}

# mitch aggregate function needs to be changes to `mean`
tic()
m <- mitch_import(x=l,DEtype="prescored",geneTable=gt)
toc() #9.0s
head(m)

```

```{r,mitch1}

tic()
mres <- mitch_calc(x=m,genesets=gs_symbols,minsetsize=5,cores=16, priority="effect")
toc() #11.45s

mtable <- mres$enrichment_result
head(mtable,20)

```

Rank-rank

```{r,rankrank}

mrank <- mres$ranked_profile

palette <- colorRampPalette(c("white", "yellow", "orange", "red", "darkred", "black"))
k <- MASS::kde2d(mrank[,1], mrank[,2])
X_AXIS = "Discovery"
Y_AXIS = "Replication"

filled.contour(k, color.palette = palette, plot.title = {
        abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")
        title(main = "rank-rank plot", xlab = X_AXIS,
        ylab = Y_AXIS)
    })

pdf("fig6a.pdf")
palette <- colorRampPalette(c("white", "yellow", "orange", "red", "darkred", "black"))
k <- MASS::kde2d(mrank[,1], mrank[,2])
X_AXIS = "Discovery"
Y_AXIS = "Replication"

filled.contour(k, color.palette = palette, plot.title = {
        abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")
        title(main = "rank-rank plot", xlab = X_AXIS,
        ylab = Y_AXIS)
    })
dev.off()

```

PW Scatter

```{r,pwscatter}

head(mtable)

sig <- subset(mtable,p.adjustMANOVA<0.05)
plot(mtable$s.Discovery,mtable$s.Replication,pch=19,col="gray",
  xlab="Discovery",ylab="Replication",main="pathway enrichment score")
points(sig$s.Discovery,sig$s.Replication,pch=19,col="red")
abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")

pdf("fig6b.pdf")
plot(mtable$s.Discovery,mtable$s.Replication,pch=19,col="gray",
  xlab="Discovery",ylab="Replication",main="pathway enrichment score")
points(sig$s.Discovery,sig$s.Replication,pch=19,col="red")
abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")
dev.off()

```

Heatmap

```{r,mheat}

mt <- head(mtable,30)
rownames(mt) <- mt$set
mt <- mt[,c("s.Discovery","s.Replication")]
rownames(mt) <- gsub("REACTOME_","",rownames(mt))
rownames(mt) <- gsub("_"," ",rownames(mt))
colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2(as.matrix(mt),scale="none",trace="none",margins=c(6,25),col=colfunc(25),cexRow=0.6,cexCol=1)

pdf("fig6c.pdf")
heatmap.2(as.matrix(mt),scale="none",trace="none",margins=c(6,25),col=colfunc(25),cexRow=0.6,cexCol=1)
dev.off()

```

Example 2d

```{r,example}

det <- mres$detailed_sets$REACTOME_CREATION_OF_C4_AND_C2_ACTIVATORS

k <- MASS::kde2d(det[,1], det[,2])
X_AXIS = "Discovery"
Y_AXIS = "Replication"

XMAX=max(mrank[,1])
XMIN=min(mrank[,1])
YMAX=max(mrank[,2])
YMIN=min(mrank[,2])

filled.contour(k, color.palette = palette, xlim=c(XMIN,XMAX), ylim=c(YMIN,YMAX),
      plot.title = {
        abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")
        title(main = "Creation of C4 and C2 Activators", xlab = X_AXIS,
        ylab = Y_AXIS)
    })

pdf("fig6d.pdf")
filled.contour(k, color.palette = palette, xlim=c(XMIN,XMAX), ylim=c(YMIN,YMAX),
      plot.title = {
        abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")
        title(main = "Creation of C4 and C2 Activators", xlab = X_AXIS,
        ylab = Y_AXIS)
    })
dev.off()

```

Make a html report and some charts.

```{r,report1}

mitch_report(res=mres,outfile="aging_mitchreport.html",overwrite=TRUE)
mitch_plots(res=mres,outfile="aging_mitchcharts.pdf")

```

## Session information

```{r,save}

save.image("aging_example.Rdata")

sessionInfo()

```