### 2023-12-18 figs/ART_estill2016.Rmd
added estill 450k dataset

---
title: "Estill 2016 reanalysis"
author: "The methylation in ART meta-analysis group"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

In this report, I will take you through a re-analysis methylation data first described by 
[Estill et al (2016)](https://www.ncbi.nlm.nih.gov/pubmed/27288894).

In their study, they analysed the DNA methylation patterns of 137 neonatal blood spots
conceived naturally (NAT), through through insemination (IUI), or through ICSI using fresh or
cryopreserved (frozen) embryo transfer (FH and FZ respectively).

The platform used in the study is the Illumina Infinium HumanMethylation450k BeadChip assay.
The authors used a pipeline based on ChAMP 
([Tian et al, 2017](https://www.ncbi.nlm.nih.gov/pubmed/28961746)), together with Adjacent
Site Clustering ([Sofer et al, 2013](https://www.ncbi.nlm.nih.gov/pubmed/23990415)).

The methylation data have been deposited to NCBI GEO repository accession number 
[GSE79257](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE79257).

The main conclusions from the original study were:

* The methylation profiles of assisted reproductive technology and IUI newborns were dramatically
different from those of naturally (in vivo) conceived newborns.

* Profiles of ICSI-frozen (FET) and IUI infants were strikingly similar, suggesting that
cryopreservation may temper some of the epigenetic aberrations induced by IVF or ICSI.

*  The DNA methylation changes associated with IVF/ICSI culture conditions and/or parental
infertility were detected at metastable epialleles, suggesting a lasting impact on a child's
epigenome.

* Both infertility and ICSI alter DNA methylation at specific genomic loci, an effect that 
is mitigated to some extent by FET (freezing).

The aim of this work is to;

1. develop the analytical pipelines required for efficient re-analysis of 450K array data,

2. to confirm that we are able to obtain differential methylation results that are similar
to those obtained in the original study, and

3. to critically evaluate the conclusions made in the original study.

In this report I will be using the missMethyl vignette as a guide to analyse this dataset 
([Phipson et al, 2015](https://www.ncbi.nlm.nih.gov/pubmed/26424855)).

Previously I curated a list of good and bad probes using the script "filter_probes.Rmd".

## Loading packages and functions

These packackes will help us to perform vital steps such as normalisation, filtering, 
differential analysis, etc, and provide information about the array probe annotaions.

These functions provide shortcuts to help with charts and other analysis. They will
eventually be shoved into another Rscript or package but can stay here for now.

```{r,packages}

source("meth_functions.R")

# Annotation
anno <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
anno <- data.frame(anno[,c("UCSC_RefGene_Name","Regulatory_Feature_Group",
  "Islands_Name","Relation_to_Island")])
promoters <- grep("Prom",myann$Regulatory_Feature_Group)

```

## Probe sets

For each gene, extract out the probes.

```{r,probesets}

promoters <- grep("Prom",anno$Regulatory_Feature_Group)
promoters <- anno[promoters,]
gp <- anno[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
sets <- split(rep(names(gp2), lengths(gp2)), unlist(gp2))
summary(unlist(lapply(sets,length)))

```

## Data import

TODO- add explanation

```{r,data}

dir.create("GSE79257")

ARRAY_SAMPLESHEET="GSE79257/GSE79257_Illumina_samplesheet.csv"
# only download it if it is not present on the system
if ( !file.exists(ARRAY_SAMPLESHEET ) ) {
    DLFILE=paste(ARRAY_SAMPLESHEET,".gz",sep="")
    download.file("https://ftp.ncbi.nlm.nih.gov/geo/series/GSE79nnn/GSE79257/suppl/GSE79257_Illumina_samplesheet.csv.gz",
        destfile = DLFILE)
    gunzip(DLFILE, overwrite = TRUE )
}

ARRAY_DATA="GSE79257/GSE79257_RAW.tar"
# only download it if it is not present on the system
if ( !dir.exists("GSE79257/IDAT") ) {
  dir.create("GSE79257/IDAT")
  if ( !file.exists(ARRAY_DATA)  ) {
    system('wget -O idats.tar.gz "https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE79257&format=file"')
    file.rename("idats.tar.gz",ARRAY_DATA)
    untar(exdir = "GSE79257/IDAT", tarfile = ARRAY_DATA)
  }
}

baseDir <- "GSE79257"
targets <- read.metharray.sheet(baseDir)
head(targets)
rgSet <- read.metharray.exp(targets = targets)

```

## Normalisation

```{r,norm,fig.cap="Figure 1. Normalisation of bead-array data with SWAN."}

mSet <- preprocessRaw(rgSet)
mSetSw <- SWAN(mSet,verbose=FALSE)
par(mfrow=c(1,2), cex=0.8)
densityByProbeType(mSet[,1], main = "Raw")
densityByProbeType(mSetSw[,1], main = "SWAN")

```

## Filter probes

Here we are running parallel analyses, both including and excluding sex chromosomes.

```{r,filterprobes}

# include sex chromosomes
detP <- detectionP(rgSet)
keep <- rowSums(detP < 0.01) == ncol(rgSet)
mSetSw <- mSetSw[keep,]

# exclude SNP probes
mSetSw <- mapToGenome(mSetSw)
mSetSw_nosnp <- dropLociWithSnps(mSetSw)
dim(mSetSw)
dim(mSetSw_nosnp)
mSetSw <- mSetSw_nosnp

# exclude sex chromosomes
keep <- !(featureNames(mSetSw) %in% ann450k$Name[ann450k$chr %in% c("chrX","chrY")])
mSetFlt <- mSetSw[keep,]
head(mSetFlt)
dim(mSetFlt)

```

## Extracting Beta and M-values

```{r,beta_m_vals}

# include sex chromosomes
meth <- getMeth(mSetSw)
unmeth <- getUnmeth(mSetSw)
Mval <- log2((meth + 100)/(unmeth + 100))
beta <- getBeta(mSetSw)

# exclude sex chromosomes
meth <- getMeth(mSetFlt)
unmeth <- getUnmeth(mSetFlt)
Mval_flt <- log2((meth + 100)/(unmeth + 100))
beta_flt <- getBeta(mSetFlt)

```

## MDS analysis

[Multidimensional scaling(https://en.wikipedia.org/wiki/Multidimensional_scaling)  plot is a method used
to identify the major sources of variation in a dataset. In the MDS plots below, I will be plotting the 
first two dimensions (principal components [PCs]), with each sample label coloured either by ART 
classification, sex, ART and sex, and then array chip and then sample plate.

We wil begin with MDS analysis including the sex chromosomes and then exclude them.

First, let's quantify the contribution of the major principal components. with a scree plot, we can see 
whether most of the variation is captured in the first two PCs or whether it is spread over more PCs.
As we can see in Figure 2, the main source of variation is what is shown in PC1, and a much lesser
extent on the other dimensions. Interestingly, excluding sex chromosomes does not seem to change the 
relative contributions of PCs very much.

```{r,scree,fig.cap="Figure 2. Screeplot shows contribution of top principal components to the overall variation in the experiment."}

par(mfrow=c(2,1))
myscree(Mval,main="incl sex chr")
myscree(Mval_flt,main="excl sex chr")

```

Here is the MDS plot by ART classification (Figure 3). You can see that there are four clusters 
of samples when MDS is projected this way. Natural birth infants (NAT) are mostly seen in the 
two clusters at the bottom of the chart and the IVF frozen, IVF fresh and in intrauterine 
insemination (IUI) appear mostly as two clusters at the top of the chart. IUI seem to have a
few samples that appear as intermediate between the upper and lower clusters.

When excluding sex chromosomes, the samples tend to form two clusters with the NAT samples on
the left of the chart and others on the right of the chart. IUI and FH have some samples which
have intermediate methylation, which points to a slight effect of freezing.

```{r,mds1,fig.width = 8 ,fig.height = 8,fig.cap="Figure 3. MDS plot coloured by ART classification."}

targets$sex <- factor(sapply(strsplit(targets$Sample_Group,"_"),"[[",2))
targets$art <- factor(sapply(strsplit(targets$Sample_Group,"_"),"[[",1))
sample_groups <- factor(targets$art)
colour_palette=brewer.pal(n = length(levels(sample_groups)), name = "Paired")
colours <- colour_palette[as.integer(factor(targets$art))]
plot(1,axes = FALSE,xlab="",ylab="",main="MDS by ART type")
legend("center",legend=levels(sample_groups),pch=16,cex=1.2,col=colour_palette)
mydist <- plotMDS(Mval, labels=targets$Sample_Name,col=colours,main="sex chromosomes included")
mydist_flt <- plotMDS(Mval_flt, labels=targets$Sample_Name,col=colours,main="sex chromosomes excluded")

```

Next, we created an MDS plot by sex (Figure 4). The female samples appear in the two cluster on the 
right hand side of the chart and the male sample on the left. This could be simply due to the
chromosomal makeup of the male and female samples or could also be due to autosomes. To check this,
the MDS plot needs to be repeated with autosomes excluded.

When excluding the sex chromosomes, there is no clear clustering of the samples by sex. It appears 
that ART classification is the dominant source of variation.

```{r,mds2,fig.width = 8 ,fig.height = 8, fig.cap="Figure 4. MDS plot coloured by sex."}

sample_groups <- factor(targets$sex)
colour_palette=brewer.pal(n = length(levels(sample_groups)), name = "Paired")
colours <- colour_palette[as.integer(factor(targets$sex))]
plot(1,axes = FALSE,xlab="",ylab="",main="MDS by sex")
legend("center",legend=levels(sample_groups),pch=16,cex=1.2,col=colour_palette)
plotMDS(mydist, labels=targets$Sample_Name,col=colours,main="sex chromosomes included")
plotMDS(mydist_flt, labels=targets$Sample_Name,col=colours,main="sex chromosomes excluded")

```

For completeness, we show the MDS by ART and sex (Figure 5). It confirms the trends seen above where the 
samples are split on the x axis by sex and on the y axis by ART classification.

When sex chromosomes are removed, it is clear that ART classification is the dominant source of variance.

```{r,mds3,fig.width = 8 ,fig.height = 8, fig.cap="Figure 5. MDS plot coloured by ART classification and sex."}

sample_groups <- factor(targets$Sample_Group)
colour_palette=brewer.pal(n = length(levels(sample_groups)), name = "Paired")
colours <- colour_palette[as.integer(factor(targets$Sample_Group))]
plot(1,axes = FALSE,xlab="",ylab="",main="MDS by ART and sex")
legend("center",legend=levels(sample_groups),pch=16,cex=1.2,col=colour_palette)
plotMDS(mydist, labels=targets$Sample_Name,col=colours,main="sex chromosomes included")
plotMDS(mydist_flt, labels=targets$Sample_Name,col=colours,main="sex chromosomes excluded")

```

To acertain whether technical factors like batch effects account for variance on the two main PCs, we
create an MDS by array chip (Figure 6). There appears to be no relationship between array chip number
and the top two PCs in the two MDS plots shown below.

```{r,mds4,fig.width = 8 ,fig.height = 8, fig.cap="Figure 6. MDS plot coloured by array chip."}

sample_groups <- factor(targets$Array)
colour_palette=brewer.pal(n = length(levels(sample_groups)), name = "Paired")
colours <- colour_palette[as.integer(factor(targets$Array))]
plot(1,axes = FALSE,xlab="",ylab="",main="MDS by array chip")
legend("center",legend=levels(sample_groups),pch=16,cex=1.2,col=colour_palette)
plotMDS(mydist, labels=targets$Sample_Name,col=colours,main="sex chromosomes included")
plotMDS(mydist_flt, labels=targets$Sample_Name,col=colours,main="sex chromosomes excluded")

```

This was also performed for sample plate (Figure 7). It appears that the laboratory did not randomise
samples over plates 1 and 2. For example the samples on plate 2 appear to be mostly NAT. This does
not appear to be a major concern because NAT samples from plate 2 seem to cluster with NAT samples on
plate 1 in both MDS plots.

```{r,mds5,fig.width = 8 ,fig.height = 8, fig.cap="Figure 7. MDS plot coloured by sample plate."}

sample_groups <- factor(targets$Sample_Plate)
colour_palette=brewer.pal(n = length(levels(sample_groups)), name = "Paired")
colours <- colour_palette[as.integer(factor(targets$Sample_Plate))]
plot(1,axes = FALSE,xlab="",ylab="",main="MDS by sample plate")
legend("center",legend=levels(sample_groups),pch=16,cex=1.2,col=colour_palette)
plotMDS(mydist, labels=targets$Sample_Name,col=colours,main="sex chromosomes included")
plotMDS(mydist_flt, labels=targets$Sample_Name,col=colours,main="sex chromosomes excluded")

```

## Differential analysis

There are several differential contrasts that would be of interest to us in this study:

* __NAT (43) vs FH (38)__: to see what probes are affected by IVF where the embryos are not frozen.

* __NAT (43) vs FZ (38)__: to see what probes are affected by IVF where the embryos are frozen.

* __NAT (43) vs IUI (18)__: It will be interesting to see if the probes affected by IVF are also sensitive to
IUI, a process that involves in utero injection of sperm.

* __FH (38) vs FZ (38)__: to see whether freezing results in differential methylation.

* __NAT (43) vs FH+FZ (76)__: if freezing results in only a small effect, then these groups can be 
lumped together and provide more power to determine subtle effects of IVF.

* __IUI (18) vs FH+FZ (76)__: It will be interesting to see whether the process of extracting eggs, culturing
and fertilising in vitro has an effect that is separate to IUI.

The differential analysis is centred around limma to identify differentially methylated probes. 
TopConfects was also run to obtain the probes with the largest confident effect (topconfect).
There are five outputs below:

1. Volcano plot (limma result).

2. Beeswarm plot (top probes by limma p-value).

3. Heatmap (top probes by limma p-value).

4. Beeswarm plot (top probes by topconfect ranking).

5. Heatmap (top probes by topconfect ranking). (TODO)

### NAT vs FH

In this first analysis I will look at the effect of natural conception versus IVF (fresh).
I have conducted the differential analysis in parallel, both includng and excluding sex
chromosomes. Even when excluding autosomes, there are ~2000 probes on the autosomes that 
show differential methylation between sexes. Based on this data and MDS analysis, we will
now exclude sex chromosomes from downstream analysis.

In the comparison of NAT and FH samples there were ~150k probes with differential methylation
(FDR<0.05), with ~120k loss of methylation and ~30k with increased methylation.

```{r,nat_vs_fh_incl, fig.width = 8 ,fig.height = 8 }

# include sex chromosomes
samplesheet <- subset(targets,art=="NAT" | art=="FH")
samplesheet$Basename <- sapply(strsplit(samplesheet$Basename, "/"), "[[", 3)
groups <- factor(samplesheet$art,levels=c("NAT","FH"))
sex <- factor(samplesheet$sex,levels=c("Male","Female"))

top_nat_vs_fh_inc <- dm_analysis(samplesheet=samplesheet,
    sex=sex,groups=groups,mx=Mval,name="top_nat_vs_fh_inc",
    myann=myann ,beta= beta)

head(top_nat_vs_fh_inc$dma)
head(top_nat_vs_fh_inc$dmr)
top_nat_vs_fh_inc$comp
top_nat_vs_fh_inc$cgi
names(top_nat_vs_fh_inc)
head(top_nat_vs_fh_inc$confects$table$name)

```

```{r,nat_vs_fh_excl, fig.width = 8 ,fig.height = 8}

# exclude sex chromosomes
samplesheet <- subset(targets,art=="NAT" | art=="FH")
samplesheet$Basename <- sapply(strsplit(samplesheet$Basename, "/"), "[[", 3)
groups <- factor(samplesheet$art,levels=c("NAT","FH"))
sex <- factor(samplesheet$sex,levels=c("Male","Female"))

top_nat_vs_fh <- dm_analysis(samplesheet=samplesheet,
    sex=sex,groups=groups,mx=Mval_flt,name="top_nat_vs_fh",
    myann=myann, beta=beta_flt)

head(top_nat_vs_fh$dma)
 head(top_nat_vs_fh$dmr)
top_nat_vs_fh$comp
top_nat_vs_fh$cgi
write.table(top_nat_vs_fh$dma,file="estill_nat_vs_fh.tsv",sep="\t",quote=FALSE)

```

### NAT vs FZ

In the comparison of NAT and FZ samples, there were ~130k probes with lower methylation
and ~37k with higher methylation.

```{r,nat_vs_fz, fig.width = 8 ,fig.height = 8}

samplesheet <- subset(targets,art=="NAT" | art=="FZ")
samplesheet$Basename <- sapply(strsplit(samplesheet$Basename, "/"), "[[", 3)
groups <- factor(samplesheet$art,levels=c("NAT","FZ"))
sex <- factor(samplesheet$sex,levels=c("Male","Female"))

top_nat_vs_fz <- dm_analysis(samplesheet=samplesheet,
    sex=sex,groups=groups,mx=Mval_flt,name="top_nat_vs_fz",
    myann=myann, beta=beta_flt)

head(top_nat_vs_fz$dma)
head(top_nat_vs_fz$dmr)
top_nat_vs_fz$comp
write.table(top_nat_vs_fz$dma,file="estill_nat_vs_fz.tsv",sep="\t",quote=FALSE)

```

### NAT vs IUI

In the comparison of NAT and IUI samples there were ~114k probes with reduced methylation
and ~36k with higher methylation.

```{r,nat_vs_iui, fig.width = 8 ,fig.height = 8}

samplesheet <- subset(targets,art=="NAT" | art=="IUI")
groups <- factor(samplesheet$art,levels=c("NAT","IUI"))
samplesheet$Basename <- sapply(strsplit(samplesheet$Basename, "/"), "[[", 3)
sex <- factor(samplesheet$sex,levels=c("Male","Female"))
table(groups)
top_nat_vs_iui <- dm_analysis(samplesheet=samplesheet,
    sex=sex,groups=groups,mx=Mval_flt,name="top_nat_vs_iui",
    myann=myann, beta=beta_flt)

head(top_nat_vs_iui$dma)
head(top_nat_vs_iui$dmr)
top_nat_vs_iui$comp
write.table(top_nat_vs_iui$dma,file="estill_nat_vs_iui.tsv",sep="\t",quote=FALSE)

```

### FH vs FZ

In the comparison of FH and FZ, only ~13k probes were differentially methylated with roughly
half lower and higher. This suggests that freezing has only a small effect on methylation.

```{r,fh_vs_fz, fig.width = 8 ,fig.height = 8}

samplesheet <- subset(targets,art=="FH" | art=="FZ")
groups <- factor(samplesheet$art,levels=c("FH","FZ"))
samplesheet$Basename <- sapply(strsplit(samplesheet$Basename, "/"), "[[", 3)
sex <- factor(samplesheet$sex,levels=c("Male","Female"))
table(groups)
top_fh_vs_fz <- dm_analysis(samplesheet=samplesheet,
    sex=sex,groups=groups,mx=Mval_flt,name="top_fh_vs_fz",
    myann=myann, beta=beta_flt)

head(top_fh_vs_fz$dma)
head(top_fh_vs_fz$dmr)
top_fh_vs_fz$comp
write.table(top_fh_vs_fz$dma,file="estill_fh_vs_fz.tsv",sep="\t",quote=FALSE)

```

### NAT vs FH/FZ

Because the effect of embryo freezing was found to be very small, we compared the NAT samples
to all IVF samples (union of FH and FZ). There were ~200k probes with altered methylation
~157k with lower methylation and ~50k higher. This result points to a major effect of IVF on
infant genome methylation.

```{r,nat_vs_fhfz, fig.width = 8 ,fig.height = 8}

samplesheet <- subset(targets,art=="NAT" | art=="FH" | art=="FZ")
samplesheet$Basename <- sapply(strsplit(samplesheet$Basename, "/"), "[[", 3)
samplesheet$art <- gsub("FH","FX",samplesheet$art)
samplesheet$art <- gsub("FZ","FX",samplesheet$art)
groups <- factor(samplesheet$art,levels=c("NAT","FX"))
sex <- factor(samplesheet$sex,levels=c("Male","Female"))
table(groups)
top_nat_vs_fx <- dm_analysis(samplesheet=samplesheet,
    sex=sex,groups=groups,mx=Mval_flt,name="top_nat_vs_fx",
    myann=myann, beta=beta_flt)

head(top_nat_vs_fx$dma)
head(top_nat_vs_fx$dmr)
top_nat_vs_fx$comp
write.table(top_nat_vs_fx$dma,file="estill_nat_vs_fx.tsv",sep="\t",quote=FALSE)

```

### IUI vs FH/FZ

We know that IVF conceived infants have a major shift in genome methylation, but we don't know
exactly whether it is caused by the IVF process itself. To eliminate whether it might be potentially
caused by subfertility we looked at whether the differences between IUI and FH+FZ were large.

We observe only ~1500 differentiallymethylated probes which is drastically smaller than the NAT vs
FH/FZ comparison which suggests that the process of IVF accounts for only a small fraction (1%) of the 
variance in methylation. The difference in methylomes could be due to other factors such as subfertility
or even the effect of hormone therapy.

```{r,iui_vs_fhfz, fig.width = 8 ,fig.height = 8}

samplesheet <- subset(targets,art=="IUI" | art=="FH" | art=="FZ")
samplesheet$Basename <- sapply(strsplit(samplesheet$Basename, "/"), "[[", 3)
samplesheet$art <- gsub("FH","FX",samplesheet$art)
samplesheet$art <- gsub("FZ","FX",samplesheet$art)
groups <- factor(samplesheet$art,levels=c("IUI","FX"))
sex <- factor(samplesheet$sex,levels=c("Male","Female"))
table(groups)
top_iui_vs_fx <- dm_analysis(samplesheet=samplesheet,
    sex=sex,groups=groups,mx=Mval_flt,name="top_iui_vs_fx",
    myann=myann, beta=beta_flt)

head(top_iui_vs_fx$dma)
head(top_iui_vs_fx$dmr)
top_iui_vs_fx$comp
write.table(top_iui_vs_fx$dma,file="estill_iui_vs_fx.tsv",sep="\t",quote=FALSE)

```

## Venn diagrams of the differential methylated probes for each contrast

First, we look at the similarity of DMPs altered by fresh and frozen procedures. The overlap is very large.
This indicates a high degree of similarity between the profiles.

```{r,venn1,fig.cap="Comparison of probes altered by fresh and frozen IVF procedures", fig.width = 8 ,fig.height = 8}

v1 <- list("FH up" = top_nat_vs_fh$dm_up , 
           "FZ up" = top_nat_vs_fz$dm_up ,
           "FH dn" = top_nat_vs_fh$dm_dn ,
           "FZ dn" = top_nat_vs_fz$dm_dn )
plot(euler(v1, shape = "ellipse"), quantities = TRUE)

```

Now to take a look at whether IUI is similar to IVF fresh and frozen. The overlap is just a large as 
FH vs FZ.

```{r,venn2,fig.cap="Comparison of probes altered by IUI, fresh and frozen IVF procedures", fig.width = 8 ,fig.height = 8}

v2 <- list("FH up" = top_nat_vs_fh$dm_up , 
           "FH dn" = top_nat_vs_fh$dm_dn ,
           "IUI up" = top_nat_vs_iui$dm_up ,
           "IUI dn" = top_nat_vs_iui$dm_dn )
plot(euler(v2, shape = "ellipse"), quantities = TRUE)

v3 <- list("FZ up" = top_nat_vs_fz$dm_up , 
           "FZ dn" = top_nat_vs_fz$dm_dn ,
           "IUI up" = top_nat_vs_iui$dm_up ,
           "IUI dn" = top_nat_vs_iui$dm_dn )
plot(euler(v3, shape = "ellipse"), quantities = FALSE)

```

## Spearman correlations of each contrast

While the above Venn diagrams are suggestive of similarity between contrasts, the best 
way to assess this is with correlation analysis. I have chosen Spearman as based on
the directional p-value ranking metric.

```{r,spearman, fig.width = 8 ,fig.height = 8}

mycontrasts <- list("top_nat_vs_fh"=top_nat_vs_fh, "top_nat_vs_fz"=top_nat_vs_fz, "top_nat_vs_fx"=top_nat_vs_fx ,
  "top_nat_vs_iui"=top_nat_vs_iui, "top_fh_vs_fz"=top_fh_vs_fz, "top_iui_vs_fx"=top_iui_vs_fx)

myrnks <- lapply(X = mycontrasts, FUN = myranks)
df <- join_all(myrnks,by="rn")
rownames(df) <- df$rn
df$rn=NULL
colnames(df) <- names(mycontrasts)
head(df)

mycors <- cor(df,method = "spearman")

my_palette <- colorRampPalette(c("darkred","red", "orange", "yellow","white"))(n = 25)
heatmap.2(mycors,scale="none",margin=c(10, 10),cexRow=0.8,trace="none",cexCol=0.8,
    col=my_palette,main="Spearman correlations")

mycors

```

## References

Estill MS, Bolnick JM, Waterland RA, Bolnick AD, Diamond MP, Krawetz SA.
Assisted reproductive technology alters deoxyribonucleic acid methylation
profiles in bloodspots of newborn infants. Fertil Steril. 2016 Sep
1;106(3):629-639.e10. doi: 10.1016/j.fertnstert.2016.05.006. Epub 2016 Jun 8.
PubMed PMID: 27288894.

Phipson B, Maksimovic J, Oshlack A. missMethyl: an R package for analyzing
data from Illumina's HumanMethylation450 platform. Bioinformatics. 2016 Jan
15;32(2):286-8. doi: 10.1093/bioinformatics/btv560. Epub 2015 Sep 30. PubMed
PMID: 26424855.

Sofer T, Schifano ED, Hoppin JA, Hou L, Baccarelli AA. A-clustering: a novel
method for the detection of co-regulated methylation regions, and regions
associated with exposure. Bioinformatics. 2013 Nov 15;29(22):2884-91. doi:
10.1093/bioinformatics/btt498. Epub 2013 Aug 29. PubMed PMID: 23990415; PubMed
Central PMCID: PMC3810849.

Tian Y, Morris TJ, Webster AP, Yang Z, Beck S, Feber A, Teschendorff AE.
ChAMP: updated methylation analysis pipeline for Illumina BeadChips.
Bioinformatics. 2017 Dec 15;33(24):3982-3984. doi: 10.1093/bioinformatics/btx513.
PubMed PMID: 28961746; PubMed Central PMCID: PMC5860089.

## Session info

```{r,sessioninfo}

sessionInfo()

```
END OF REPORT