### 2023-12-19 figs/art_example.Rmd
integrative example

---
title: "DNA meth and IVF"
author: "The GMEA team and ART epigenetics team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Here we are looking at putting together the easiest possible workflow for the package.

We will analyze methylation data for IVF conceived infants compared to naturally conceived.

Data from:

1. Estill MS, Bolnick JM, Waterland RA, Bolnick AD, Diamond MP, Krawetz SA. Assisted reproductive technology alters deoxyribonucleic acid methylation profiles in bloodspots of newborn infants. Fertil Steril. 2016 Sep 1;106(3):629-639.e10. doi: 10.1016/j.fertnstert.2016.05.006. Epub 2016 Jun 8. PMID: 27288894.

2. Novakovic B, Lewis S, Halliday J, Kennedy J, Burgner DP, Czajko A, Kim B, Sexton-Oates A, Juonala M, Hammarberg K, Amor DJ, Doyle LW, Ranganathan S, Welsh L, Cheung M, McBain J, McLachlan R, Saffery R. Assisted reproductive technologies are associated with limited epigenetic variation at birth that largely resolves by adulthood. Nat Commun. 2019 Sep 2;10(1):3922. doi: 10.1038/s41467-019-11929-9. PMID: 31477727; PMCID: PMC6718382.

## Requirements

```{r,packages}

suppressPackageStartupMessages({
  library("limma")
  library("eulerr")
  library("IlluminaHumanMethylation450kmanifest")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("tictoc")
  library("mitch")
  library("gplots")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("gridExtra")
  library("png")
})

CORES=16

```

## Load data

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

We'll also score the probes by the signed -log pvalue

```{r,load1}

# estill 2016 450k
art1 <- readRDS("estill_nat_vs_fh.rds")
art1 <- art1[[1]]
rownames(art1) <- art1$Row.names
str(art1)

# novakovic 2019 EPIC
art2 <- readRDS("novakovic_nat_vs_fh.rds")
art2 <-art2[[1]]
rownames(art2) <- art2$Row.names
str(art2)

gs_symbols <- gmt_import("c2.cp.reactome.v2023.1.Hs.symbols.gmt")

```

## Curate the annotation

Use all probes on the chip. Start with 450k chip.

```{r,anno1}

tic()
anno <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt1 <- stack(gp2)
colnames(gt1) <- c("gene","probe")
gt1$probe <- as.character(gt1$probe)
dim(gt1)
str(gt1)
length(unique(gt1$gene))
toc() #6.3s

```

Now EPIC chip.

```{r,anno2}

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt2 <- stack(gp2)
colnames(gt2) <- c("gene","probe")
gt2$probe <- as.character(gt2$probe)
dim(gt2)
str(gt2)
length(unique(gt2$gene))
toc() #10.6s

```

## Import for mitch

```{r,import1}

m1 <- mitch_import(x=art1,DEtype="limma",geneTable=gt1)
dim(m1)
head(m1)

m2 <- mitch_import(x=art2,DEtype="limma",geneTable=gt2)
dim(m2)
head(m2)

m <- merge(m1,m2,by=0)
rownames(m) <- m[,1]
m[,1]=NULL
colnames(m) <- c("Estill","Novakovic")
dim(m)
head(m)

```

## Run mitch

```{r,mitch1}

tic()
mres <- mitch_calc(x=m,genesets=gs_symbols,minsetsize=5,cores=16, priority="effect")
toc() #11.45s

mtable <- mres$enrichment_result
sig <- subset(mtable,p.adjustMANOVA<0.05)
head(sig,20)

```

Rank-rank

```{r,rankrank}

mrank <- mres$ranked_profile

palette <- colorRampPalette(c("white", "yellow", "orange", "red", "darkred", "black"))
k <- MASS::kde2d(mrank[,1], mrank[,2])
X_AXIS = "Estill 450k"
Y_AXIS = "Novakovic EPIC"

filled.contour(k, color.palette = palette, plot.title = {
        abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")
        title(main = "rank-rank plot", xlab = X_AXIS,
        ylab = Y_AXIS)
    })

pdf("fig7a.pdf")
palette <- colorRampPalette(c("white", "yellow", "orange", "red", "darkred", "black"))
k <- MASS::kde2d(mrank[,1], mrank[,2])
X_AXIS = "Estill 450k"
Y_AXIS = "Novakovic EPIC"

filled.contour(k, color.palette = palette, plot.title = {
        abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")
        title(main = "rank-rank plot", xlab = X_AXIS,
        ylab = Y_AXIS)
    })
dev.off()

```

PW Scatter

```{r,pwscatter}

head(mtable)

sig <- subset(mtable,p.adjustMANOVA<0.05)
plot(mtable$s.Discovery,mtable$s.Replication,pch=19,col="gray",
  xlab="Estill 450k",ylab="Novakovic EPIC",main="pathway enrichment score")
points(sig$s.Discovery,sig$s.Replication,pch=19,col="red")
abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")

pdf("fig7b.pdf")
plot(mtable$s.Discovery,mtable$s.Replication,pch=19,col="gray",
  xlab="Estill 450k",ylab="Novakovic EPIC",main="pathway enrichment score")
points(sig$s.Discovery,sig$s.Replication,pch=19,col="red")
abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")
dev.off()

```

Heatmap

```{r,mheat}

mt <- head(mtable,30)
rownames(mt) <- mt$set
mt <- mt[,c("Estill 450k","Novakovic EPIC")]
rownames(mt) <- gsub("REACTOME_","",rownames(mt))
rownames(mt) <- gsub("_"," ",rownames(mt))
colfunc <- colorRampPalette(c("blue", "white", "red"))
heatmap.2(as.matrix(mt),scale="none",trace="none",margins=c(6,25),col=colfunc(25),cexRow=0.6,cexCol=1)

pdf("fig7c.pdf")
heatmap.2(as.matrix(mt),scale="none",trace="none",margins=c(6,25),col=colfunc(25),cexRow=0.6,cexCol=1)
dev.off()

```

Example 2d

```{r,example}

det <- mres$detailed_sets$REACTOME_SENSORY_PERCEPTION

k <- MASS::kde2d(det[,1], det[,2])
X_AXIS = "Discovery"
Y_AXIS = "Replication"

XMAX=max(mrank[,1])
XMIN=min(mrank[,1])
YMAX=max(mrank[,2])
YMIN=min(mrank[,2])

filled.contour(k, color.palette = palette, xlim=c(XMIN,XMAX), ylim=c(YMIN,YMAX),
      plot.title = {
        abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")
        title(main = "Creation of C4 and C2 Activators", xlab = X_AXIS,
        ylab = Y_AXIS)
    })

pdf("fig7d.pdf")
filled.contour(k, color.palette = palette, xlim=c(XMIN,XMAX), ylim=c(YMIN,YMAX),
      plot.title = {
        abline(v = 0, h = 0, lty = 2, lwd = 2, col = "blue")
        title(main = "Creation of C4 and C2 Activators", xlab = X_AXIS,
        ylab = Y_AXIS)
    })
dev.off()

```

Make a html report and some charts.

```{r,report1}

mitch_report(res=mres,outfile="aging_mitchreport.html",overwrite=TRUE)
mitch_plots(res=mres,outfile="aging_mitchcharts.pdf")

```

## Session information

```{r,save}

save.image("aging_example.Rdata")

sessionInfo()

```