### 2023-12-20 figs/GSE158422.Rmd
added main script

---
title: "GSE158422: DNA methylation in Lung squamous cell carcinoma"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
  toc: true
theme: cosmo
---

## Introduction

This work is to conduct the methylation data processing and differential analysis of GSE158422, which is matched
lung squamous cell carcinoma (LUSC) and control tissue.

The data is going to be saved as an Rdata file and then used for enrichment analysis in a subsequent Rmd script.

```{r,packages}

suppressPackageStartupMessages({
  library("plyr")
  library("R.utils")
  library("missMethyl")
  library("limma")
  library("DMRcate")
  library("DMRcatedata")
  library("topconfects")
  library("minfi")
  library("IlluminaHumanMethylation450kmanifest")
  library("RColorBrewer")
  library("IlluminaHumanMethylation450kanno.ilmn12.hg19")
  library("GEOquery")
  library("eulerr")
  library("plyr")
  library("gplots")
  library("reshape2")
  library("forestplot")
  library("beeswarm")
  library("RCircos")
  library("qqman")
  library("ENmix")
})

source("../meth_functions.R")

```

## Obtaining array annotations

```{r,annotation}

anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","Regulatory_Feature_Group","Islands_Name","Relation_to_Island")])
promoters <- grep("Prom",myann$Regulatory_Feature_Group)

```

## Obtain array data

GSE158422 is the accession number for the array data.

```{r,arraydata}

ARRAY_DATA="GSE158422_RAW.tar"
if(!file.exists(ARRAY_DATA)){
  download.file("https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE158422&format=file",
    destfile=ARRAY_DATA)
  untar(tarfile = ARRAY_DATA)
}

SERIES_MATRIX="GSE158422_series_matrix.txt.gz"
if(!file.exists(SERIES_MATRIX)){
download.file("https://ftp.ncbi.nlm.nih.gov/geo/series/GSE158nnn/GSE158422/matrix/GSE158422_series_matrix.txt.gz",
  destfile=SERIES_MATRIX)
}
gse <- getGEO(filename=SERIES_MATRIX)
baseDir <- "IDAT"
sample_metadata <- pData(phenoData(gse))
sample_metadata$patient <- factor(sapply(strsplit(sample_metadata$title,"_"),"[[",1))

targets <- sample_metadata
mybase <- unique(gsub("_Red.idat.gz" ,"",  gsub("_Grn.idat.gz", "" ,list.files(".",pattern = "GSM",recursive = TRUE))))
targets$Basename <- mybase
head(targets)
rgSet <- read.metharray.exp(targets = targets)
rgSet

targets$Basename <- gsub("IDAT/","",mybase)
head(targets)

```

## Filtering and normalisation

```{r,norm,fig.cap="Figure 1. Normalisation of bead-array data with SWAN."}

mSet <- preprocessRaw(rgSet)
mSetSw <- SWAN(mSet,verbose=FALSE)
par(mfrow=c(1,2), cex=0.8)
densityByProbeType(mSet[,1], main = "Raw")
densityByProbeType(mSetSw[,1], main = "SWAN")

```

```{r,filterprobes}

# include sex chromosomes
detP <- detectionP(rgSet)
keep <- rowSums(detP < 0.01) == ncol(rgSet)
mSetSw <- mSetSw[keep,]

# exclude SNP probes
mSetSw <- mapToGenome(mSetSw)
mSetSw_nosnp <- dropLociWithSnps(mSetSw)

# exclude sex chromosomes
xyprobes <- anno$Name[anno$chr %in% c("chrX","chrY")]
mSetFlt <- mSetSw[which(!rownames(mSetSw) %in% xyprobes),]

```

```{r,beta_m_vals}

# include sex chromosomes
meth <- getMeth(mSetSw)
unmeth <- getUnmeth(mSetSw)
Mval <- log2((meth + 100)/(unmeth + 100))
beta <- getBeta(mSetSw)
dim(Mval)

# exclude sex chromosomes
meth <- getMeth(mSetFlt)
unmeth <- getUnmeth(mSetFlt)
Mval_flt <- log2((meth + 100)/(unmeth + 100))
beta_flt <- getBeta(mSetFlt)
dim(Mval_flt)

```

# Infer sex

```{r,infersex}

xprobes <- anno$Name[anno$chr %in% "chrX"]
yprobes <- anno$Name[anno$chr %in% "chrY"]

xdat <- Mval[which( rownames(Mval) %in% xprobes ),]
ydat <- Mval[which( rownames(Mval) %in% yprobes ),]
gsm <- sapply( strsplit(colnames(Mval),"_" ) , "[[" , 1)
plot(colMeans(xdat) , colMeans(ydat), cex=1)
text(colMeans(xdat) , colMeans(ydat),labels=gsm )
grid()
abline(v=-0.1,col="red",lty=2)

which( colMeans(xdat)  > -0.1 )

sex_female <- as.numeric(colMeans(xdat)  > -0.1)
names(sex_female) <- colnames(xdat)

```

# MDS analysis

```{r,scree,fig.cap="Figure 2. Screeplot shows contribution of top principal components to the overall variation in the experiment."}

# scree plot shows the amount of variation in a dataset that is accounted
# for by the first N principal components
myscree <- function(mx,n=10,main="") {
  pc <- princomp(mx)$sdev
  pcp <- pc/sum(pc)*100
  pcp <- pcp[1:10]
  barplot(pcp,cex.names = 1,las=2,ylim=c(0,60),
      ylab="percent (%) variance explained", main=main)
  text((0.5:length(pcp)*1.2),pcp,label=signif(pcp,3),pos=3,cex=0.8)
}

par(mfrow=c(2,1))
myscree(Mval,main="incl sex chr")
myscree(Mval_flt,main="excl sex chr")
par(mfrow=c(1,1))

```

MDS plot 1

```{r,mds1,fig.width = 8 ,fig.height = 8,fig.cap="Figure 3. MDS plot coloured by ART classification."}

colnames(targets) <- gsub(" ","_",colnames(targets))
colnames(targets) <- gsub(":","_",colnames(targets))
targets$sex<-factor(sex_female)
targets$case<-factor(targets$`characteristics_ch1`)
sample_groups <- factor(targets$case)
colour_palette=brewer.pal(n = length(levels(sample_groups)), name = "Paired")
colours <- colour_palette[as.integer(factor(targets$case))]
plot(1,axes = FALSE,xlab="",ylab="",main="MDS by case/ctrl",cex=0)
legend("center",legend=levels(sample_groups),pch=16,cex=1.2,col=colour_palette)
mydist <- plotMDS(Mval, labels=gsm,col=colours,main="sex chromosomes included")
mydist_flt <- plotMDS(Mval_flt, labels=gsm,col=colours,main="sex chromosomes excluded")

```

## Differential analysis

Paired normal versus cancer tissue.

```{r,dm}

samplesheet <- targets

patient <- factor(sapply(strsplit(samplesheet$title,"_"),"[[",1))

tumor <- factor(samplesheet$`tumor:ch1`,levels=c("normal","tumor"))

sex <- sex_female

dim(Mval)
dim(beta)

design <- model.matrix(~ sex + patient + tumor )
mx <- Mval_flt
name="control vs tumor"

mxs <- mx[,which( colnames(mx) %in% samplesheet$Basename )]
fit.reduced <- lmFit(mxs,design)
fit.reduced <- eBayes(fit.reduced)
summary(decideTests(fit.reduced))
dm <- topTable(fit.reduced,coef=ncol(design), number = Inf)
dma <- merge(myann,dm,by=0)
dma <- dma[order(dma$P.Value),]
dm_up <- rownames(subset(dm,adj.P.Val<0.05 & logFC>0))
dm_dn <- rownames(subset(dm,adj.P.Val<0.05 & logFC<0))
sig <- min(length(dm_up),length(dm_dn))

confects <- limma_confects(fit.reduced, coef=3, fdr=0.05)

if (sig > 10) {
  dmr <- run_dmrcate(mx=mxs,design=design)
  head(dmr)
  comp <- compartment_enrichment(dma)
  comp
  cgi <- cgi_enrichment(dma)
  cgi
} else {
  dmr <- NULL
  comp <- NULL
  cgi <- NULL
}
#make_dm_plots(dm = dm ,name=name , mx=beta,mxs=mxs, groups = groups,
  confects=confects,dmr = dmr, comp=comp, cgi=cgi)

GSE158422 <- list("dma"=dma, "dm_up"=dm_up, "dm_dn"=dm_dn, "confects"=confects,
  "dmr"= dmr, "comp"=comp, "cgi"=cgi, "fit"=fit.reduced)

head(GSE158422$dma,10)
head(GSE158422$dmr,10)

saveRDS(GSE158422,file="GSE158422.rds")
saveRDS(design,file="GSE158422_design.rds")
saveRDS(mx,file="GSE158422_mx.rds")

write.table(GSE158422$dma,file="GSE158422_limma.tsv",sep="\t",quote=FALSE)

```

## save data object

```{r,save}

save.image("GSE158422.Rdata")

```