### 2023-12-20 figs/main.Rmd

---
title: "Regenerate figures for the GMEA paper"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
  toc: true
theme: cosmo
---

## Introduction

Here we evaluate the performance of different FCS methods for Infinium methylation analysis.

# Workflow

## Simulation: Figure 1 and 2

```{r,fig12}

GSE158422.Rmd
GSE158422_simulate020.Rmd
GSE158422_simulate050.Rmd
GSE158422_simulate100.Rmd
GSE158422_simulate_summary.Rmd

```

## Sensitivity: Figure 3 and 4

```{r,fig34}

GSE158422_sensitivity.Rmd

```

## Integration with RNA: Figure 5

```{r,fig5}

GSE158422_rna_int.Rmd

```

## Aging study

```{r,fig6}

aging_example.Rmd

```

## IVF example

```{r,fig7}

ART_estill2016.Rmd
ART_novakovic.Rmd
art_example.Rmd

```

# Session information

```{r,session}

sessionInfo()

```