### 2024-01-12 figs/GSE158422_simulate_rslurm_summary.Rmd
added HPC results

---
title: "Testing gene methylation enrichment analysis approaches using simulated data - summarise data"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Here is an explanation of the different objects:

0. GSA: gres

1. LAT: lacres

2. LTT: lactopres

3. LAW: nlacres

4. LAM: lacmres

5. LRM: lrmres

6. ALT: alcres

7. ALW: nalcres

8. ALM: almres

```{r,packages}

suppressPackageStartupMessages({
  library("kableExtra")
})

```

## Functions

```{r,functions}

F1 <- function(x,y) {
  ( 2 * x * y ) / ( x + y )
}

PRECREC <- function(x){
  y <- aggregate(.~groupsize+delta,x,mean)
  y$PREC <- y$TP / ( y$TP + y$FP )
  y$REC <- y$TP / ( y$TP + y$FN )
  y$F1 <- F1(y$PREC,y$REC)
  y
}
# usage: lapply(x20,PRECREC)

smrz <- function(x) {
  P <- mean(x$PREC)
  R <- mean(x$REC)
  F <- F1(P,R)
  c(P,R,F)
}
# usage: smrz(x20a)

```

## Load data

```{r,load}

load("GSE158422_simulate020_rslurm.Rdata")

x20 <- list("GSA"=gres,"LAT"=latres,"LTT"=lttres,"LAW"=lawres,
  "LAM"=lamres,"LRM"=lrmres,"ALT"=altres,"ALW"=alwres,"ALM"=almres)

load("GSE158422_simulate050_rslurm.Rdata")

x50 <- list("GSA"=gres,"LAT"=latres,"LTT"=lttres,"LAW"=lawres,
  "LAM"=lamres,"LRM"=lrmres,"ALT"=altres,"ALW"=alwres,"ALM"=almres)

load("GSE158422_simulate100_rslurm.Rdata")

x100 <- list("GSA"=gres,"LAT"=latres,"LTT"=lttres,"LAW"=lawres,
  "LAM"=lamres,"LRM"=lrmres,"ALT"=altres,"ALW"=alwres,"ALM"=almres)

rm(myann)
rm(gp2)
rm(normal_mval)
rm(mval,anno)

```

## Overall summary

Merge the data at gene set sizes of 20, 50 and 100 into a single result.

```{r,merge,fig.width=5,fig.height=4}

all <- lapply(1:length(x20),function(i) {
  mg <- rbind(x20[[i]],x50[[i]],x100[[i]])
  m <- apply(mg,2,mean)
  P <- m["TP"] / ( m["TP"] + m["FP"])
  R <- m["TP"] / ( m["TP"] + m["FN"] )
  F <- F1(P,R)
  res <- c(P,R,F)
  names(res) <- c("P","R","F")
  return(res)
} )
names(all) <- names(x20)
all <- do.call(rbind,all)
colnames(all) <- c("PREC","REC","F1")

all %>% kbl(caption="Overall simulation results") %>% kable_paper("hover", full_width = F)

par(mar = c(5.1, 4.1, 5.1, 2.1))
cols <- c("white","gray","black")

barplot(t(all),beside=TRUE,ylim=c(0,1),ylab="index",
  col=cols, cex.names=0.7, cex.axis=0.7, main="Overall simulation results")
grid()
legend("right", inset=.01, c("Precision","Recall","F1"), fill=cols, cex=0.7)

```

## Focus on differences caused byu gene set size

Make a set of three barplots which summarise method performance over the range of group size and
delta values, for gene set sizes of 20, 50 and 200.

```{r,x20summary,fig.width=5,fig.height=4}

groupsizes=unique(x20[[1]]$groupsize)
deltas=unique(x20[[1]]$delta)

x20a <- lapply(x20,PRECREC)
x20aa <- lapply(x20a,smrz)
x20aa <- do.call(rbind,x20aa)
colnames(x20aa) <- c("PREC","REC","F1")
x50a <- lapply(x50,PRECREC)
x50aa <- lapply(x50a,smrz)
x50aa <- do.call(rbind,x50aa)
colnames(x50aa) <- c("PREC","REC","F1")
x100a <- lapply(x100,PRECREC)
x100aa <- lapply(x100a,smrz)
x100aa <- do.call(rbind,x100aa)
colnames(x100aa) <- c("PREC","REC","F1")

x20aa %>% kbl(caption="sim 20 res") %>% kable_paper("hover", full_width = F)
x50aa %>% kbl(caption="sim 50 res") %>% kable_paper("hover", full_width = F)
x100aa %>% kbl(caption="sim 100 res") %>% kable_paper("hover", full_width = F)

barplot(t(x20aa),beside=TRUE,ylim=c(0,1),ylab="index",
  col=cols, cex.names=0.8, cex.axis=0.8, main="20 genes per set")
abline(h=seq(0.2,1,0.2),lty=2,col="gray")
legend("right", inset=.01, c("Precision","Recall","F1"), fill=cols, bg="white", cex=0.8)
barplot(t(x50aa),beside=TRUE,ylim=c(0,1),ylab="index",
  col=cols, cex.names=0.8, cex.axis=0.8, main="50 genes per set")
abline(h=seq(0.2,1,0.2),lty=2,col="gray")
legend("right", inset=.01, c("Precision","Recall","F1"), fill=cols, bg="white", cex=0.8)
barplot(t(x100aa),beside=TRUE,ylim=c(0,1),ylab="index",
  col=cols, cex.names=0.8, cex.axis=0.8, main="100 genes per set")
abline(h=seq(0.2,1,0.2),lty=2,col="gray")
legend("bottomright", inset=.01, c("Precision","Recall","F1"), fill=cols, bg="white", cex=0.8)
dev.off()

pdf("fig1_slurm.pdf",height=6,width=7)
par(mfrow=c(2,2))
par(mar = c(2.5, 4, 3, 1))
barplot(t(all),beside=TRUE,ylim=c(0,1),ylab="index",
  col=cols, cex.names=0.65, cex.axis=0.8, main="Overall simulation results")
abline(h=seq(0.2,1,0.2),lty=2,col="gray")
legend("right", inset=0, c("Precision","Recall","F1"), fill=cols, bg="white", cex=0.8)
mtext("A", 2, adj=2.5, las=1,padj=-4,cex=2)
barplot(t(x20aa),beside=TRUE,ylim=c(0,1),ylab="index",
  col=cols, cex.names=0.65, cex.axis=0.8, main="20 genes per set")
abline(h=seq(0.2,1,0.2),lty=2,col="gray")
legend("right", inset=0, c("Precision","Recall","F1"), fill=cols, bg="white", cex=0.8)
mtext("B", 2, adj=2.5, las=1,padj=-4,cex=2)
barplot(t(x50aa),beside=TRUE,ylim=c(0,1),ylab="index",
  col=cols, cex.names=0.65, cex.axis=0.8, main="50 genes per set")
abline(h=seq(0.2,1,0.2),lty=2,col="gray")
legend("right", inset=0, c("Precision","Recall","F1"), fill=cols, bg="white", cex=0.8)
mtext("C", 2, adj=2.5, las=1,padj=-4,cex=2)
barplot(t(x100aa),beside=TRUE,ylim=c(0,1),ylab="index",
  col=cols, cex.names=0.65, cex.axis=0.8, main="100 genes per set")
abline(h=seq(0.2,1,0.2),lty=2,col="gray")
legend("bottomright", inset=0, c("Precision","Recall","F1"), fill=cols, bg="white", cex=0.8)
mtext("D", 2, adj=2.5, las=1,padj=-4,cex=2)
dev.off()

```

## In depth analysis of GSA vs LAM

As GSA appears to be the best-performing existing method for enrichment analysis, it is a good idea
to compare it to LAM which is the best one out of the new methods evaluated here.
The following charts show the performance of GSA and LAM including precision and recall over the
range of group sizes and delta values when the sets contain 20, 50 and 100 genes each.

First with 20 genes per set.

```{r,020_gsa_vs_lam}

par(mfrow=c(2,3))
par(mar = c(4.1, 4.1, 3.5, 0.9))

gsa <- lapply(deltas,function(d) { subset(x20a$GSA,delta==d) } )
lam <- lapply(deltas,function(d) { subset(x20a$LAM,delta==d) } )

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"PREC"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"PREC"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"PREC"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"PREC"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"PREC"] ,type="b",col="darkorange")
grid()
legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2", "0.1"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen", "darkorange"), cex=1, bg="white")

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"REC"],type="b",las=1,ylab="recall",xlab="group size",
  main="GSA",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"REC"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"REC"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"REC"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"REC"] ,type="b",col="darkorange")
grid()

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"F1"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"F1"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"F1"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"F1"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"F1"] ,type="b",col="darkorange")
grid()

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"PREC"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"PREC"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"PREC"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"PREC"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"PREC"] ,type="b",col="darkorange")
grid()
legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2", "0.1"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen", "darkorange"), cex=1, bg="white")

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"REC"],type="b",las=1,ylab="recall",xlab="group size",
  main="LAM",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"REC"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"REC"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"REC"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"REC"] ,type="b",col="darkorange")
grid()

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"F1"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"F1"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"F1"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"F1"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"F1"] ,type="b",col="darkorange")
grid()

```

Now with 50 genes per set.

```{r,050_gsa_vs_lam}

par(mfrow=c(2,3))
par(mar = c(4.1, 4.1, 3.5, 0.9))

gsa <- lapply(deltas,function(d) { subset(x50a$GSA,delta==d) } )
lam <- lapply(deltas,function(d) { subset(x50a$LAM,delta==d) } )

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"PREC"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"PREC"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"PREC"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"PREC"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"PREC"] ,type="b",col="darkorange")
grid()
legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2", "0.1"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen", "darkorange"), cex=1, bg="white")

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"REC"],type="b",las=1,ylab="recall",xlab="group size",
  main="GSA",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"REC"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"REC"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"REC"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"REC"] ,type="b",col="darkorange")
grid()

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"F1"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"F1"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"F1"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"F1"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"F1"] ,type="b",col="darkorange")
grid()

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"PREC"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"PREC"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"PREC"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"PREC"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"PREC"] ,type="b",col="darkorange")
grid()
legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2", "0.1"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen", "darkorange"), cex=1, bg="white")

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"REC"],type="b",las=1,ylab="recall",xlab="group size",
  main="LAM",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"REC"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"REC"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"REC"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"REC"] ,type="b",col="darkorange")
grid()

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"F1"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"F1"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"F1"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"F1"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"F1"] ,type="b",col="darkorange")
grid()

```

Now 100 genes per set

```{r,100_gsa_vs_lam}

par(mfrow=c(2,3))
par(mar = c(4.1, 4.1, 3.5, 0.9))

gsa <- lapply(deltas,function(d) { subset(x100a$GSA,delta==d) } )
lam <- lapply(deltas,function(d) { subset(x100a$LAM,delta==d) } )

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"PREC"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"PREC"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"PREC"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"PREC"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"PREC"] ,type="b",col="darkorange")
grid()
legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2", "0.1"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen", "darkorange"), cex=1, bg="white")

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"REC"],type="b",las=1,ylab="recall",xlab="group size",
  main="GSA",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"REC"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"REC"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"REC"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"REC"] ,type="b",col="darkorange")
grid()

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"F1"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"F1"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"F1"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"F1"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"F1"] ,type="b",col="darkorange")
grid()

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"PREC"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"PREC"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"PREC"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"PREC"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"PREC"] ,type="b",col="darkorange")
grid()
legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2", "0.1"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen", "darkorange"), cex=1, bg="white")

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"REC"],type="b",las=1,ylab="recall",xlab="group size",
  main="LAM",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"REC"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"REC"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"REC"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"REC"] ,type="b",col="darkorange")
grid()

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"F1"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"F1"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"F1"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"F1"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"F1"] ,type="b",col="darkorange")
grid()

## PDF output

pdf("fig2_slurm.pdf",height=6,width=7)
par(mfrow=c(2,3))
par(mar = c(4.1, 4.1, 3.5, 0.9))

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"PREC"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"PREC"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"PREC"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"PREC"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"PREC"] ,type="b",col="darkorange")
grid()
legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2", "0.1"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen", "darkorange"), cex=1, bg="white")

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"REC"],type="b",las=1,ylab="recall",xlab="group size",
  main="GSA",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"REC"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"REC"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"REC"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"REC"] ,type="b",col="darkorange")
grid()

plot(gsa[[length(gsa)]][,"groupsize"],  gsa[[length(gsa)]][,"F1"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(gsa[[length(gsa)-1]][,"groupsize"], gsa[[length(gsa)-1]][,"F1"] ,type="b",col="red")
lines(gsa[[length(gsa)-2]][,"groupsize"], gsa[[length(gsa)-2]][,"F1"] ,type="b",col="blue")
lines(gsa[[length(gsa)-3]][,"groupsize"], gsa[[length(gsa)-3]][,"F1"] ,type="b",col="darkgreen")
lines(gsa[[length(gsa)-4]][,"groupsize"], gsa[[length(gsa)-4]][,"F1"] ,type="b",col="darkorange")
grid()

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"PREC"],type="b",las=1,ylab="precision",xlab="",
  main="",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"PREC"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"PREC"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"PREC"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"PREC"] ,type="b",col="darkorange")
grid()
legend("bottomright", legend=c("0.5", "0.4", "0.3", "0.2", "0.1"),title="delta",
  lty=1, lwd=2, col=c("black", "red", "blue", "darkgreen", "darkorange"), cex=1, bg="white")

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"REC"],type="b",las=1,ylab="recall",xlab="group size",
  main="LAM",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"REC"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"REC"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"REC"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"REC"] ,type="b",col="darkorange")
grid()

plot(lam[[length(lam)]][,"groupsize"],  lam[[length(lam)]][,"F1"],type="b",las=1,ylab="F1",xlab="",
  main="",ylim=c(0,1))
lines(lam[[length(lam)-1]][,"groupsize"], lam[[length(lam)-1]][,"F1"] ,type="b",col="red")
lines(lam[[length(lam)-2]][,"groupsize"], lam[[length(lam)-2]][,"F1"] ,type="b",col="blue")
lines(lam[[length(lam)-3]][,"groupsize"], lam[[length(lam)-3]][,"F1"] ,type="b",col="darkgreen")
lines(lam[[length(lam)-4]][,"groupsize"], lam[[length(lam)-4]][,"F1"] ,type="b",col="darkorange")
grid()

dev.off()

```

## Session information

```{r,save}

sessionInfo()

```