### 2024-01-15 example_workflow/example_workflow.Rmd
example workflow

---
title: "Example of using mitch package for Infinium methylation analysis"
author: "The GMEA team"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    fig_width: 7
    fig_height: 7
theme: cosmo
---

## Introduction

Here we are looking at putting together the easiest possible workflow for the package.

Source code: https://github.com/markziemann/gmea/blob/main/example_workflow/example_workflow.Rmd

## Requirements

Load packages.

```{r,packages}

suppressPackageStartupMessages({
  library("limma")
  library("eulerr")
  library("IlluminaHumanMethylation450kanno.ilmn12.hg19")
  library("IlluminaHumanMethylationEPICanno.ilm10b4.hg19")
  library("HGNChelper")
  library("tictoc")
  library("mitch")
  library("kableExtra")
  library("beeswarm")
  library("missMethyl")
  library("gridExtra")
  library("png")
})

CORES=2

```

## Load methylation data

The limma data table is read in.
This dataset is available from GEO (GSE49667).
The limma data was obtained using the "A cross-package Bioconductor workflow for analysing methylation array data" workflow.

```{r,loaddata}

dm <- read.csv("DMPs.csv.gz",header=TRUE)

head(dm)

```
## Load pathways

Reactome pathways were downloaded on the 14th Sept 2023 from MsigDB.

```{r,loadpathways}

gene_sets <- gmt_import("c5.go.v2023.2.Hs.symbols.gmt")

```

## Curate the annotation

Use all probes on the HM450K chip.

Updating gene names using the HGNC package because a lot of them have been updated over time.

An example of making a gene table for EPIC data is included down below.

```{r,anno1}

tic()
anno <- getAnnotation(IlluminaHumanMethylation450kanno.ilmn12.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt1 <- stack(gp2)
colnames(gt1) <- c("gene","probe")
gt1$probe <- as.character(gt1$probe)
dim(gt1)
str(gt1)
length(unique(gt1$gene))
toc() #6.0s

tic()
#new.hgnc.table <- getCurrentHumanMap()
new.hgnc.table <- readRDS("../figs/new.hgnc.table.rds")
fix <- checkGeneSymbols(gt1$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gt1$gene <- fix$Suggested.Symbol
toc() #  36 sec (2788 gene names updated)

head(gt1)

```

## Mitch pipeline

The first part is to import the data into mitch.

This only works properly when mitch's mapGeneIds function uses `mean` to aggregate (line 69 of mitch.R).

```{r,mg1}

# mitch aggregate function needs to be changes to `mean`
tic()
m <- mitch_import(x=dm,DEtype="limma",geneTable=gt1,geneIDcol="Name")
toc() #2.5 sec

head(m) %>%
  kbl(caption = "Example of differential gene methylation scores used for pathway analysis") %>%
  kable_paper("hover", full_width = F)

```

Now run the enrichment analysis.

Note that the results are not sorted by p-value, rather S.distance, an enrichment score.

```{r,mitch1}

tic()
mres <- mitch_calc(x=m,genesets=gene_sets,minsetsize=5,cores=4, priority="effect")
toc() #7 sec on 4 threads

mtable <- mres$enrichment_result
up <- subset(mtable,s.dist>0 & p.adjustANOVA<0.05)
dn <- subset(mtable,s.dist<0 & p.adjustANOVA<0.05)
nrow(up)
nrow(dn)

head(up,10) %>%
  kbl(caption = "Top significant pathways with higher methylation") %>%
  kable_paper("hover", full_width = F)

head(dn,10) %>%
  kbl(caption = "Top significant pathways with lower methylation") %>%
  kable_paper("hover", full_width = F)

top <- rbind(up[1:10,],dn[1:10,])
top <- top[order(top$s.dist),]

barnames <- gsub("_"," ",gsub("REACTOME_","",top$set))
cols <- as.character(((sign(top$s.dist)+1)/2)+1)
cols <- gsub("1","blue",gsub("2","red",cols))

par(mar = c(5.1, 29.1, 4.1, 2.1))
barplot(abs(top$s.dist),horiz=TRUE,las=1,names.arg=barnames, cex.names=0.8,
  cex.axis=0.8,col=cols, xlab="Enrichment score",main="Reactomes")
mtext("whole gene")
grid()

par( mar = c(5.1, 4.1, 4.1, 2.1) )

```

Make a html report and some charts.

```{r,report1}

mitch_report(res=mres,outfile="example_mitchreport.html",overwrite=TRUE)
mitch_plots(res=mres,outfile="example_mitchcharts.pdf")

```

## Make gene table for EPIC array data

```{r,gt2}

tic()
anno <- getAnnotation(IlluminaHumanMethylationEPICanno.ilm10b4.hg19)
myann <- data.frame(anno[,c("UCSC_RefGene_Name","UCSC_RefGene_Group","Islands_Name","Relation_to_Island")])
gp <- myann[,"UCSC_RefGene_Name",drop=FALSE]
gp2 <- strsplit(gp$UCSC_RefGene_Name,";")
names(gp2) <- rownames(gp)
gp2 <- lapply(gp2,unique)
gt <- stack(gp2)
colnames(gt) <- c("gene","probe")
gt$probe <- as.character(gt$probe)
dim(gt)
str(gt)
toc() #9.0s

tic()
#new.hgnc.table <- getCurrentHumanMap()
new.hgnc.table <- readRDS("../figsnew.hgnc.table.rds")
fix <- checkGeneSymbols(gt$gene,map=new.hgnc.table)
fix2 <- fix[which(fix$x != fix$Suggested.Symbol),]
length(unique(fix2$x))
gt$gene <- fix$Suggested.Symbol
toc()

```

## Session information

```{r,save}

sessionInfo()

```