### 2024-01-15 manuscript/manuscript.Rmd
update ms

---
title: "Direction-aware functional class scoring enrichment analysis of Infinium DNA methylation data"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
bibliography: references.bib
csl: epigenetics.csl
---

<strong>
Mark Ziemann<sup>1,2*</sup>,
Mandhri Abeysooriya<sup>2,3</sup>,
Anusuiya Bora<sup>1,2</sup>,
Severin Lamon<sup>3</sup>,
Mary Sravya Kasu<sup>1</sup>,
Mitch W. Norris<sup>1</sup>,
Yen Ting Wong<sup>4</sup>,
Jeffrey Craig<sup>4,5</sup>
</strong>

1. Burnet Institute, Melbourne, Australia

2. Deakin University, School of Life and Environmental Sciences, Geelong, Australia.

3. Deakin University, School of Exercise and Nutrition, Burwood, Australia.

4. Deakin University, School of Medicine, Geelong, Australia.

5. Murdoch Children's Research Institute, Melbourne, Australia.

(*) Correspondence: mark.ziemann@burnet.edu.au

| Author | ORCID |
| --- | --- |
| Mark Ziemann | 0000-0002-7688-6974 |
| Mandhri Abeysooriya | 0000-0003-2163-6203 |
| Anusuiya Bora | 0009-0006-2908-1352 |
| Severine Lamon | 0000-0002-3271-6551 |
| Mary Sravya Kasu | 0000-0002-7891-836X |

Source: https://github.com/markziemann/gmea

<br>

## Abstract

Infinium methylation beadarrays remain one of the most popular platforms for studying genomic
DNA methylation, but tools for downstream pathways analysis have their limitations.
Functional class scoring (FCS) is a group of pathway enrichment techniques that involve the ranking of
genes and evaluation of their collective regulation in biological systems, but the implementations described for
Infinium methylation data do not retain direction information which is important for mechanistic understanding
of genomic regulation.
Here, we evaluate several candidate FCS methods that retain directional information.
According to simulation results, the best performing method involves the mean aggregation of probe limma
t-statistics by gene followed by a rank-ANOVA enrichment test using the mitch package.
This method, which we call "LAM", outperformed an existing over-representation analysis method in
simulations, and showed higher sensitivity and robustness in an analysis of real lung tumor-normal paired datasets.
Using matched RNA-seq data we examine the relationship of methylation differences at promoters and
gene bodies with RNA expression at the level of pathways in lung cancer.
To further demonstrate the utility of our approach, we apply it to two other contexts; firstly an examination of
methylation differences associated with chronological age; and secondly to examine methylation differences in
infants conceived with in vitro fertilisation.
These results indicate that the LAM method is more sensitive and precise as compared to existing methods.
LAM has the potential to illuminate disease mechanisms and identify novel diagnostic signatures.
A vignette is provided to illustrate how to implement this method.

**Keywords:** Pathway analysis, functional enrichment analysis, Infinium Array, DNA methylation,
epigenetics.

<br>

## Introduction

DNA methylation is the most widely studied epigenetic mark, for its role in
development and disease [@Robertson2005-tu].
Many hundreds of epigenome-wide association studies (EWASs) are conducted each year to understand DNA
methylation patterns in disease and other contexts [@Li2019-og].
Infinium arrays remain the preferred platform for EWASs due to low cost and analytical simplicity
as compared to high throughput methylation sequencing [@Wei2021-qj].
Infinium arrays typically have multiple probes per gene in different locations including promoters,
CpG islands, gene bodies and enhancers [@Pidsley2016-ix], which complicates downstream functional enrichment analysis,
otherwise known as pathway analysis or gene set enrichment analysis.
Enrichment analysis comes in two popular types; over-representation analysis (ORA) and functional
class scoring (FCS) [@Khatri2012-qy].
A third type of enrichment analysis called pathway topology improves upon other methods with more sophisticated
modelling of gene network activations, but these are yet to be adopted widely [@Xie2021-py].
In ORA, genes with probes that meet an arbitrary significance threshold are selected and compared
to a background list consisting of genes measured in the assay.
The test seeks to identify sets of genes (eg: ontologies, pathways) that are over-represented in the
gene list of interest relative to the background [@Tavazoie1999-ng].
FCS takes a different approach by ranking all detected genes by a differential
regulation score (eg: fold change, confident effect size) followed by a test to assess whether each
set of genes has a distribution of scores that is different to the null [@Subramanian2005-ar].
The `gsameth()` function of the `missMethyl` package is the state-of-the-art method for ORA of infinium methylation array
data as it addresses issues related to probes belonging to more than one gene and the fact that genes can have multiple
probes [@Maksimovic2021-ur].
As gsameth is an ORA method, results dependent strongly on the significance threshold used [@Ren2019-mr], and also
on the proportion of probes that meet this threshold.
FCS methods `ebGSEA` and `methylGSA` have been developed for EWAS data and are suggested to have better sensitivity to
determine subtle associations between pathways and differential methylation [@Dong2019-np;@Ren2019-mr].
The ebGSEA tool uses an empirical Bayes modeling approach followed by a one-tailed nonparametric test for gene set
enrichment [@Dong2019-np].
The `methylGSA` package implements two different approaches.
Firstly, the `methylRRA` function which uses robust rank aggregation for detecting differentially methylated genes taking
into consideration the variable number of probes followed by an FCS test based on z-scores [@Ren2019-mr].
Secondly, `methylglm` uses a logistic regression approach to model differential methylation for genes inside and outside each gene set
[@Ren2019-mr].
These methods are indeed more sensitive than ORA approaches, but as they are based on one-tailed tests irrespective of
direction of methylation change, it limits the utility for downstream interpretation of the importance of methylation
changes on genomic regulation.
Moreover, as pathways tend to be either upregulated or downregulated in omics assays, combining the two tends to dilute
their signal and reduce sensitivity [@Hong2014-ab].
Neither GSAmeth nor methylGSA report computed enrichment scores, which hinders downstream interpretation, as the
enrichment score is a useful surrogate for effect size.
Without an enrichment score, users are limited to using statistical significance values to prioritise results, which
biases towards large gene sets that have non-specific biological functions.

Here we aim to develop and evaluate methods for two-tailed FCS of Infinium methylation array data that addresses these
limitations.
Due to the fact that genes have a variable number of probes, there are many concevable ways that FCS could be applied
to Infinium methylation array data.
The first set of methods use differential probe methylation data from `limma` [@Ritchie2015-zn] as an input.
The differential methylation of probes beloning to each gene are scored and these gene scores
undergo an enrichment test.
The second set of methods first aggregate probe methylation values for each gene before differential
methylation analysis with limma.
Then, differential gene methylation results are used for a downstream enrichment test.
With the above methods there are several methodological choices, such as the type of aggregation (eg: mean, median), and the type of
downstream test (eg: parametric or nonparametric).
To determine the best approach to applying FCS to Infinium methylation data, in this study we use
simulated data and compare results to existing ORA methods.
We then examine the sensitivity of selected methods on real cancer data, investigate the association with gene expression and end
with additional examples related to human aging and assisted reproductive technologies.

## Methods

### Implementation overview

Functional enrichment analysis is a process of data summarisation from genes to gene sets (pathways).
This is made more complicated for Infinium methylation array data due to the presence of multiple probes per gene,
meaning the data needs to be summarised from probes to genes and then to gene sets.
Here, we outline eight potential approaches for FCS of Infinium methylation data:

1. Limma Average t-test (LAT).
Differential methylation analysis is conducted at the probe level with limma, and the limma
t-statistics for each gene are summarised (mean).
The mean t-statistics are used in a downstream two-sample two-way t-test of gene set enrichment.

2. Limma Top t-test (LTT).
As above, except instead of calculating the mean t-statistic, the probe with the largest magnitude is selected
to represent the gene.

3. Limma Average Wilcox (LAW).
Similar to LAT, except instead of the two-sided t-test of gene set enrichment, a non-parametric alternative,
the Wilcoxon signed-rank test is used.

4. Limma Average Mitch (LAM).
Similar to LAW, however instead of the Wilcoxon test, the mitch package[@Kaspi2020-yx] is used to execute a two-way,
two-sample ANOVA-on-ranks test (Kruskal-Wallis).

5. Limma Rank Mitch (LRM).
Probes are ranked by t-statistic. The mean rank values for probes beloning to each gene are computed and used
as input for a two-way, two-sample ANOVA-on-ranks test.

6. Aggregate Limma t-test (ALT).
In this method, all probe measurements beloning to each gene are averaged before conducting differential
methylation andlysis with limma.
The limma t-statistics are then used downstream for enrichment analysis with the two-sample two-way t-test.

7. Aggregate Limma Wilcox (ALW).
Similar to ALT, however it uses the Wilcoxon non-parametric test for gene set enrichment.

8. Aggregate Limma Mitch (ALM).
Similar to ALW, however uses the Mitch package to execute a two-way, two-sample ANOVA-on-ranks test.

The above were compared to a standard ORA-based approach (GSA), which involves limma on probes, selection of
statistically significant probes (FDR<0.05) for ORA with the `GSAmeth()` function of missMethyl which conducts a
modified hypergeometric test that accounts for multiple probe biases.
In the case that fewer than 250 significant probes are identified, then 250 probes with the smallest p-values
were selected.
We conducted separate tests for increased and decreased probes, and we specified the background as all probes that
passed quality control filtering.

### Method validation using simulations

To assess these methods, we adopted a simulated data approach based on the selected modification of real methylation data.
We downloaded raw intensity EPIC IDAT files from NCBI Gene Expression Omnibus (GEO) for study GSE158422 which consists of
lung tumour and normal adjacent tissues from 37 patients [@Cho2022-yd].
For the simulations, only non-cancerous datasets were used.
Probe annotations were obtained from the "IlluminaHumanMethylationEPICanno.ilm10b4.hg19" Bioconductor
package.
We randomly sampled datasets to serve as control and cases (no replacement), with group sizes varying between
3 and 12.
One thousand random gene sets were created with sizes varying between 20 and 100 members, with member genes drawn from the
annotation set.
Random gene sets were used to avoid problems caused by the large overlap among real gene sets.
Throughout our evaluations, 50 gene sets were selected to be differentially methylated, 25 with increased and 25
with decreased methylation.
From these gene sets, half of the member genes were selected.
For those selected genes, half of the annotated probes were selected.
We adjusted the M-values in the case group by a specified amount, which we call the "delta", which we varied
between 0.1 and 0.5.
Following the incorporation of methylation changes to selected probes in the case group, the data underwent
limma differential methylation analysis, followed by enrichment analysis using `GSAmeth` approach, or one of
the eight FCS methods outlined above.
Enrichment results were then filtered for FDR<0.05 and correct direction of methylation change.
Differentially methylated gene sets observed were compared to gene sets selected to be the ground truth, so that
the number of true positives (TP), false positives (FP) and false negatives (FN) could be determined.
At each setting for group size, delta and gene set size, 100 replications were conducted each with a set seed
to ensure reproducibility.
Simulations were conducted on a SLURM cluster [@Yoo2003-mc].
Mean TP, FP and FN values were determined, and then precision, recall and F1 scores were calculated.

### Sensitivity analysis

The full GSE158422 dataset including normal and cancer samples of all 37 patients underwent differential analysis with limma
correcting for patient-of-origin effects.
Downstream pathway enrichment analysis was conducted using the GSAmeth and LAM methods
with Reactome pathways downloaded from MSigDB website (version 2023.1)[@Milacic2024-js].
LAM relies on probe-gene associations from the "UCSC_RefGene_Name" column of the "IlluminaHumanMethylationEPICanno.ilm10b4.hg19"
annotation.
As gene annotations are several years old, we used HGNChelper to update them (database current as of 21-Dec-2023) [@Oh2022-ts]
(3253 gene symbols were updated).
Euler diagrams were created with the `eulerr` R package.
To test sensitivity, a random subset of n patients was selected followed by pathway enrichment analysis with the
respective method, with n varied between 2 and 30.
Significantly enriched pathways were defined as those with FDR values < 0.05.
The subset significant results were compared to the full group (n=37) results to test the sensitivity of these enrichment
methods, and whether findings at a smaller sample size were consistent with the full group.
This process was repeated 50 times at each sample size.

### Association of methylation with gene expression

RNA-seq based gene counts for the same set of lung cancer patients were downloaded from GEO (accession: GSE158420).
HGNChelper was used to repair transmogrified gene names, followed by DESeq2 based differential analysis.
This underwent filtering to remove genes with expression below 10 reads per sample on average across the dataset.
Then DESeq2[@Love2014-wt] was used to compare normal and cancerous tissue gene expression taking into consideration sample pairing.
The differential expression results underwent enrichment analysis using the mitch package [@Kaspi2020-yx] with default settings for
DESeq2 data tables.
CpG sites annotated as promoters were considered separately from those located at gene bodies in mitch analysis.
For comparison, GSAmeth was used to separately examine pathways using separate analyses of promoter and gene body CpGs.
For this analysis, Reactome gene sets were used.
Subsequently, to generate the pathway heatmap, we used mitch in multivariate mode by analysing differential RNA expression,
promoter and gene body methylation together, prioritising the results by s-distance, a surrogate measure of effect size.

### Pathway-level differential methylation in aging

In order to deminstrate the utility of the LAM method for large EWAS studies, we examined
methylation profile associations with age of two independent cohorts involving a total of 7036
participants [@McCartney2020-ob].
Limma summary statistics for chronological age including discovery and replication groups were read into R and enrichment analysis was conducted with the LAM method.
Briefly, gene-probe associations were first obtained from the "IlluminaHumanMethylationEPICanno.ilm10b4.hg19"
dataset, followed by update of gene symbols with HGNChelper.
Discovery and replication data were imported separately with mitch and multivariate enrichment analysis was conducted with gene sets from Reactome.

### Pathway-level differential methylation in assisted reproductive technology

In order to demonstrate that methylation patterns across different array platforms can be compared using the LAM method,
we examined methylation differences in infants associated with assisted reproduction as described by two
independent studies.
The Estill study was published in 2016 and is based on the HM450K array and the data is available from GEO Accession GSE79257
[@Estill2016-jh].
The Novakovic study in was published 2019, and used the EPIC methylation array and the data is available from GEO under accession
GSE131433[@Novakovic2019-jt].
Various assisted reproductive technology conception groups are described in these studies, but we examined
just fresh in vitro fertilisation (IVF) conceived infants compared to naturally conceived infants.
Raw intensity files (IDAT format) were read into R with minfi [@Aryee2014-su], normalistion was conducted with the SWAN method
[@Maksimovic2012-hy], and probe filtering was conducted to remove proves with a detection p-value > 0.01 in addition to probes
located on X or Y chromosomes.
M values were computed and underwent differential methylation analysis using limma [@Ritchie2015-zn].
The Estill study included 43 naturally conceived infants and 38 with fresh embryo transfer.
The Novakovic study included 58 naturally conceived infants and 75 with fresh embryo transfer.
Limma differential analysis was conducted accounting for sex, then methylation tables underwent LAM analysis with Reactome gene sets.
Probe gene associations for HM450K array were established using the "IlluminaHumanMethylation450kanno.ilmn12.hg19" annotation set
and the gene symbols were updated with HGNChelper.
As the Novakovic study involved EPIC array data, gene-probe associations were obtained from the
"IlluminaHumanMethylationEPICanno.ilm10b4.hg19" annotation set, and gene symbols were updated with
HGNChelper.

## Results

### Evaluation of FCS methods for Infinium methylation array data

Simulations were conducted using the eight FCS methods (LAT, LTT, LAW, LAM, LRM, ALT, ALW, ALM) and GSAmeth,
with group sizes of 3-12 and delta values between 0.1 and 0.5, and random gene sets with sizes 20, 50
and 100.
At each parameter setting, the simulations were repeated 100 times with a different seed value.
True positives, false positives and false negatives were used to calculate overall precision and recall
at these three gene set sizes (**Figure 1A**).
The LAM and ALM methods recorded an overall precision of 0.94, while GSA scored lowest with 0.848.
Recall was highest for LAM with 0.22, while ALM scored 0.21 and GSA scored 0.15
F1 scores were highest for LAM with 0.36 followed by ALM with 0.35, while GSA scored 0.26.

We noticed that the size of gene sets used in the simulations strongly influenced the results, with overall recall increasing with
gene set size (**Figure 1B-D**).
The GSA method performed better than FCS methods with gene sets of 20 (**Figure 1B**), while at 50 genes per set GSA performance
was on par with LAM (**Figure 1C**).
At 100 genes per set however, LAM showed superior precision and recall compared to GSA (**Figure 1D**).

![**Figure 1. Precision, recall and F1 for enrichment tests with simulated data.**
(A) Overall results including 20, 50 and 100 genes per set.
(B-D) Results for simulations using 20, 50 and 100 genes per set respectively.
](img/fig1.svg)

Focusing on results from 100 genes per set, LAM precision was more consistent across the parameter ranges as compared to GSA which
showed lower precision when group size and delta were lower (**Figure 2**).
Recall was strongly dependent on group size and delta parameters.
LAM recorded relatively higher recall when group size and delta were lower.
F1 performance scores at 100 genes per set were better for LAM as compared to GSA.

![**Figure 2. Precision, recall and F1 for GSA and LAM enrichment tests with simulated data over a range of group sizes and delta
values.**
One hundred genes per set.](img/fig2.svg)

### Sensitivity of GSA and LAM methods with real cancer data

Next, GSA and LAM methods were applied to compare paired normal and cancer methylation data from 37 lung squamous cell carcinoma patients.
From the 839,473 probes passing quality control, 473,572 were considered significant at 5% FDR.
Of these, 119,785 probes were significantly hypermethylated and 353,787 were significantly hypomethylated.
These significantly hypermethylated probes mapped to 16,520 unique genes, while the hypomethylated probes mapped to 24,114 genes.
These numbers are large, considering that there are 26,219 genes represented on this array.
LAM resulted in 406 significant Reactome pathways at FDR<0.05.
Of these, 370 involved higher methylation, while 36 involved lower methylation.
LAM execution took 51 seconds using eight parallel threads.
With GSA, there were 75 Reactome pathways with higher methylation and 32 with lower methylation.
12 pathways were significant in both directions.
There were 30 pathways that were common between LAM and GSA methods.
This includes 26 and 4 with higher and lower methylation respectively (**Figure 3A**).
GSA execution for both directions in series took 22 seconds on one thread.

As the simulation results indicated LAM had better recall with larger gene sets, we were curious about whether there were differences
in the sizes of gene sets found by LAM and GSA.
GSA significant pathways had a median size of 135, while for LAM the median was 54.5, indicating pathways identified as significant
with LAM were collectively smaller than GSA (**Figure 3B**).
This may seem counterintuitive, but may be due to LAM's better sensitivity in dectecting pathway enrichment when effect sizes are
lower.

Although algorithmic accuracy cannot be inferred from the types of pathways identified in real data [@Timmons2015-kr],
top ranked results from LAM appeared to be more related to cell differentiation, identity and development (**Figure 3C**),
as compared to GSA (**Figure 3D**).

![**Figure 3. Pathway enrichment analysis comparing normal and cancer samples using LAM and GSA methods.**
(A) Euler diagrams showing the overlap in statistically significant pathways (FDR<0.05) from LAM and GSA results.
(B) Sizes of significant gene sets found with LAM and GSA methods.
(C) Significant LAM pathways with largest S distance, a proxy for effect size.
(D) Top ranked significant GSA pathways ranked by fold enrichment.](img/fig3.svg)

To compare the sensitivity of these methods with real data, sample sizes were randomly downsampled, analysed
as above and the significant pathways were compared to the results from the full
group of 37 patients.
This process was repeated 50 times at each sample size and the results shown in (**Figure 4**).
Pathways that were identified as statistically significant (FDR<0.05) in the smaller and full group were
termed "consistent", while pathways that were identified in the smaller but not in the full group were
termed "inconsistent".
At a sample size of 10, LAM was able to detect 354 out of 406 (87%) consistent pathways,
while GSA detected just 13 of 107 at this sample size (12%).
LAM identified more inconsistent pathways than GSA, but the proportion of
inconsistent findings was lower across all sample sizes.
At a sample size of 10, the proportion of inconsistent pathways was 10% for LAM and 28% for GSA.
This result suggests that LAM has superior sensitivity to detect differentially methylated
pathways in real cancer data.

![**Figure 4. Sensitivity of LAM and GSA methods with real cancer array data.**
Upper panels show the number of significant pathways identified in the smaller sample that were consistent
with the full group for LAM (left) and GSA (right).
Middle panels show the number of significant pathways identified in the smaller sample that were inconsistent
with the full group.
Lower panels show the proportion of significant pathways identified that were inconsistent with the full
group.
Left panels correspond to LAM, and right panels to GSA methods.](img/fig4.svg)

## Integrating methylation and RNA expression pathways

Associations of epigenetic marks with gene expression are of great interest for understanding disease processes.
Matching tumour-normal RNA-seq datasets were analysed and we conducted enrichment analysis of gene expression together with
DNA methylation.
From 18,704 detected genes, There were 12,380 differentially expressed genes (FDR<0.05), with 7,415 and 4,965 up- and down-regulated
in the tumour group respectively.
At the pathway level there were 304 and 135 up- and down-regulated Reactome pathways with altered gene expression (FDR<0.05).
As the context of gene methylation is important in influencing gene expression, promoter and gene body methylation
were considered separately.
In promoters, GSA identified 15 and 7 pathways with higher and lower methylation respectively, while at gene bodies, there were were
18 and 75 pathways with higher and lower methylation respectively (FDR<0.05).
Using LAM at promoters, there were 190 and 20 pathways identified with higher and lower methylation respectively.
LAM at gene bodies identified 251 and 31 pathways with higher and lower methylation respectively (FDR<0.05).
The relative overlap between significant promoter and gene body pathways was larger for LAM pathway sets (22%) as compared to GSA
(3%) (**Figure 5A,B**).
There was no observed overlap between GSA methylation-based pathways and gene expression pathways (**Figure 5A**), but there were some
overlaps between pathways identified with LAM based promoter and gene body methylation and gene expression (**Figure 5B**).

Using LAM, 136/304 upregulated RNA expression pathways were associated with increasing gene body methylation.
Likewise, 21 pathways with downregulated RNA expression were associated with increasing promoter methylation.
Some of these pathways with large enrichment scores across the three contrasts are depicted in heatmap form (**Figure 5C**).
There were a small number of pathways that exhibited an inverse relationship between promoter methylation and RNA expression such as
"unwinding of DNA", "mucopolysaccharidoses","interaction with cumulus cells and the zona pellucida", "glucocorticoid biosynthesis" and
"olfactory signaling".
These were however less common than pathways that showed a positive relationship between RNA expression and promoter methylation.
These results demonstrate that LAM can illuminate the complicated relationship between DNA methylation and gene expression at the
pathway level in cancer.

![**Figure 5. Integration of gene expression data with methylation pathways.**
(A,B) Euler diagrams showing the overlap in statistically significant pathways (FDR<0.05) from gene expression and GSA results (A)
and LAM (B) results.
(C) Multi-dimensional enrichment analysis of promoter and gene body methylation with gene expression using mitch.
Top 20 gene sets shown with largest s distance after FDR filtering at 0.05.](img/fig5.svg)

## Exploration of methylation pathways in aging

We sought to explore the utility of LAM for exploring large EWAS datasets.
One of the best studied associations with DNA methylation is with chronological aging, although
pathway level differential methylation is not well defined.
Using limma summary statistics from independent discovery and replication studies using the EPIC array platform described
by McCartney et al (2020) [@McCartney2020-ob], we applied the LAM method to explore the pathway level
differential methylation.
In the discovery cohort we observed 304 differential pathways (FDR<0.05), while in the replication group we observed 107.
There was a high degree of agreement, with 43 pathways significant in both groups.
This is not surprising, as the gene-level rank differential methylation scores show a strong positive association between discovery and
replication studies (**Figure 6A**).
Using the rank-MANOVA test of the mitch package, we identified 390 pathways with altered methylation.
Visualised as a scatterplot, the overall pattern is concordant, although a cluster of points in the lower right of the chart
indicates 268 pathways with higher methylation in the discovery sample and lower in methylation in the replication
sample (**Figure 6B**).
When prioritising the results by S distance, a surrogate measure of effect size, 29 of the top 30 pathways were concordant in their
direction of regulation (**Figure 6C**).
Some of the observed methylation changes make sense with reports in the literature.
For example the complement pathway is associated with aging and in this analysis we observe a reduction in methylation of genes
beloning to this set including *C1R*, *MBL2*, *C1S* and *C1QC* (**Figure 6D**).

![**Figure 6. Pathway level DNA methylation alterations with chronological age.**
(A) Contour heatmap showing the similarity in gene methylation score ranks in the discovery and replication studies.
(B) Mitch pathway enrichment scores (S distances) in discovery and replication studies.
Pathways with MANOVA FDR<0.05 are shown in red while others are shaded gray.
(C) Heatmap of 30 pathways with most extreme S distances (a surrogate of effect size).
Red indicates increasing methylation and blue shows lower methylation.
(D) An example of a pathway identified with this method, "creation of C4 and C2 activators" shows lower methylation of member genes
in both discovery and replication studies.](img/fig6.svg)

## Exploration of methylation pathway differences in infants conceived with in vitro fertilisation

Another application for this method is the joint pathway enrichment analysis of studies conducted with different
array systems.
Epigenetic differences between infants conceived naturally and with assisted reproductive technologies have been the subject of
studies since 2009 [@Katari2009-ya].
Two of the highest powered studies with publicly available data are Estill 2016 which used the 450K array and Novakovic 2019 study
which used the EPIC array [@Estill2016-jh;@Novakovic2019-jt].
Joint enrichment analysis can uncover pathways that are common or different between these studies and illuminate
the biological differences between conception groups.
For this example we focused on comparing naturally conceived infants to those conceived with in vitro fertilisation (IVF) with fresh
embryo transfer.
The Estill dataset on the 450k platform had 418,833 probes that met the quality criteria (Natural n=43; IVF n=38).
Of these, 149,589 probes showed statistical significant differential methylation (FDR<0.05), of which 28,767 and 120,822 exhibited
higher or lower methylation in the IVF group respectively.
The Novakovic dataset on EPIC array had 793,844 probes that met the filtering criteria (n=133).
Of these, 5,562 were statistically significant (FDR<0.05) with 1,348 and 4,214 exhibited higher or lower methylation in the IVF group
respectively.
After summarisation to the gene level, the Estill dataset on 450k array describes 19,240 genes, while the Novakovic dataset on EPIC
array describes 22,588 genes.
There were 19,234 genes common to both platforms.
The rank-rank plot of differential gene methylation scores indicates a high degree of similarity overall, and interestingly shows a
trend of more genes having lower methylation level in the IVF group (**Figure 7A**).
A scatterplot of pathway enrichment scores indicates a moderate degree of agreement between these studies
(*r*=0.35,*p*=2.2e-16) despite these studies being conducted on independent cohorts years apart and analysed with different array
systems (**Figure 7B**).
A heatmap shows a high degree of agreement of pathways with higher methylation in the IVF groups,
while pathways with overall lower methylation in the IVF group showed some variability in enrichment scores
between studies (**Figure 7C**).
We were interested in which pathways had consistently lower methylation in the IVF group and identified the Reactome pathway
"Adrenoreceptors" which had enrichment scores of -0.42 in the Estill study and -0.56 in the Novakovic study, recording a MANOVA FDR
value of 0.04 (**Figure 7D**).
Top ranked genes from this pathway include *ADRA2C*, *ADRB2*, *ADRA2B* and *ADRA1B*.
A loss of methylation of these genes could contribute to reported elevated blood pressure in people conceived by IVF [@Meister2018-cp].
Taken together, this analysis shows that LAM method can be used to compare enrichment of independent studies
conducted with 450k and EPIC arrays.

![**Figure 7. Pathway level DNA methylation differences in natural and IVF conceived infants.**
(A) Contour heatmap showing the similarity in gene methylation score ranks in the Estill (450k) and Novakovic (EPIC) studies.
(B) Mitch pathway enrichment scores (S distances) in Estil and Novakovic studies.
Pathways with MANOVA FDR<0.05 are shown in red while others are shaded gray.
(C) Heatmap of 30 pathways with most extreme S distances (a surrogate of effect size).
Red indicates higher methylation and blue shows lower methylation.
(D) An example of a pathway identified with this joint enrichment analysis method, "Adrenoreceptors" shows lower methylation of
member genes in both Estill and Novakovic studies.](img/fig7.svg)

## Discussion

* Overall, we found that the FCS methods performed better across all simulation conditions.
Of the different FCS methods tested, LAM was superior in terms of precision and recall.
LAM also fits relatively easily into existing workflows, as most EWAS studies seem to use
limma in their analyses of differential probe methylation.
These analyses also consistently showed that non-parametric methods showed better recall with only a small decrease in precision, in accordance with some previous reports.
Summarisation to gene level only takes a few seconds with the efficient base R aggregation command,
and once aggregated, signatures can be analysed with existing packages like mitch.
Mitch does not require a large amount of system memory and due to its parallel architecture can
make use of multi-threaded processors, so it is not much slower than conducting ORA with existing methods.

* But it was interesting that ORA performed better when gene sets are smaller.
Given that KEGG and Reactome pathways consist of both large (>100) and small (<20) gene sets,
a hybrid FCS and ORA approach might increase sensitivity and recall.

* At 100 genes per set, LAM had superior sensitivity compared to GSA, which agrees with mitch's performance with simulated RNA-seq gene
expression data[@Kaspi2020-yx].
The improved sensitivity of this method provides a new opportunity for researchers to reanalyse previously conducted EWASs with contemporary pathway databases to better understand
the subtle signatures, as we have demonstrated with the chronological age and assisted reproduction examples.
The analyses conducted here used Reactome pathways, but this method is general purpose, so different gene sets can be used to examine various hypotheses.
Transcription factor target gene sets can be used to identify transcription factors associated with changes in pathway methylation.
MicroRNA target gene sets can identify potential associations between microRNA targets and DNA methylation patterns.
The MSigDB resource contains these and several other types of gene sets for exploration [@Liberzon2015-um].

* The differences in pathway results obtained with LAM and GSA from lung cancer data were striking, not only in the numbers of pathways identified, but also their relevance to the disease.
Interestingly the relationship between promoter methylation and gene expression was not always inverse as we had anticipated.
Joint enrichment analysis showed some genes with increasing promoter methylation together with increasing RNA expression (eg: "Negative regulation of AP2 transcription factors").
Likewise a few pathways showed decreases in methylation with decreases in RNA expression such as "Dectin family", "TLR regulation" and "glycan biosynthesis termination".
These results indicate that there is a complicated relationship between DNA methylation and gene expression in lung cancer and further work is required to appreciate the mechanisms involved.

* We have demonstrated that this method works with both HM450K and EPIC array data and can even facilitate the comparison of EWAS using different platforms.
We have provided an example workflow that researchers can use as a template.
It shows how to generate a probe-gene table from Bioconductor annotation datasets, how to update the defunct gene names and how to run the enrichment analysis.
Using this example, generating similar gene tables for the new EPIC 2.0 array and any future arrays will be straightforward.
Updating defunct gene symbols in important as 2788 and 3253 gene symbols were updated on the HM450K and EPIC array annotations, which is approximately 10-14% of genes on these platforms.
Omitting this step could lead to a loss of these genes from the pathway enrichment and a loss of sensitivity.

## Code availability

Code to generate the results shown here are included in a GitHub repository (https://github.com/markziemann/gmea/).
To enable reproduction of the findings, a docker image is made available at DockerHub
(https://hub.docker.com/repository/docker/mziemann/gmea/general), and an example workflow can be found
in the GitHub repository (https://github.com/markziemann/gmea/tree/main/example_workflow).

## Bibliography